---
title: "Test Facies Simulations Indicators"
output: html_document
date: "2023-07-07"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

This file is meant to demonstrate the use of gstlearn by loading a numpy array containing a facies indicator variable simulated by flumy and perform some calculations based on the grid organization (transition probabilities or connectivity)

```{r install_flumy, include=TRUE}
install.packages("flumy",repos="https://soft.mines-paristech.fr/cran")
```

```{r init}
# Import package
library(gstlearn)
library(flumy)
library(ggplot2)
```

# Generate a Flumy simulation
Use the default Fluvial scenario

```{r init}
nx      = 250   # Number of grid nodes along Ox
ny      = 200   # Number of grid nodes along Oy
mesh    = 10    # Horizontal grid mesh size: 10m
hmax    = 3     # Maximum channel depth: 3m
ng      = 50    # Required Net-to-Gross: 50%
isbx    = 80    # Required sand bodies extension (medium extension = few meander cutoffs)
verbose = TRUE  # Verbose mode

res = 30                # Vertical resolution (increase 'res' to get higher resolution)
dz  = hmax / res        # Vertical discretization step (0.1m)
zul = 3 * hmax          # Fill a reservoir of 3 x hmax height (9m)
nz  = as.integer(zul / dz) # Number of vertical nodes of the resulted block of sediments
```

```{r launch}
# Launch the simulation
seed = 123456 # Simulation seed
# Create the simulator
flsim = Flumy(nx, ny, mesh, verbose)
# Launch up to zul
success = flsim$launch(seed, hmax, isbx, ng, zul)
if (!success) {
    cat("Error while running Flumy\n")
}
```

# Extract one section from the facies array

```{r display_slice}
# Retrieve facies identifiers in a 3D regular grid
fac = flsim$getDomain()$getRegularFacies(dz,0,nz)
fac[fac==255] = NA
fac3 = array(fac,dim=c(nz,ny,nx))
cat("dim(fac3)=",dim(fac3),"\n")
ufacies = sort(unique(fac))
cat("Unique facies are ", ufacies, "\n")
nfacies = length(ufacies) 

# Display the cross-section #20 (cross-flow) (col.flumy and fac.flumy are defined in the package)
nsect = 20
df = expand.grid(x=1:nx, y=1:nz)
df$val = factor(as.vector(matrix(fac3[,nsect,], nx, nz, byrow=T)))
values = col.flumy[as.integer(levels(df$val))] # Discard colors for missing facies
labels = fac.flumy[as.integer(levels(df$val))] # Discard labels for missing facies
ggplot(df, aes(x, y, fill=val)) + geom_raster() + coord_fixed() + scale_fill_manual(values=values, labels=labels)
```

# Proportions
Calculate facies global proportions

```{r proportions}
cat("Number of Facies:", nfacies)
cat("Facies Proportions:\n")
round(100*table(df$val)/(nx*ny),2)
```

# Grouping the facies
```{r group_slice}
# Grouping facies (see Flumy user guide appendix)
# https://stackoverflow.com/questions/3418128/how-to-convert-a-factor-to-integer-numeric-without-loss-of-information
arr_flat = as.numeric(levels(df$val))[df$val]
# arr_flat[arr_flat==1] = 1 # CL
# arr_flat[arr_flat==2] = 2 # PB
arr_flat[arr_flat==3]   = 2
arr_flat[arr_flat==4]   = 2
arr_flat[arr_flat==5]   = 3 # LV
arr_flat[arr_flat==6]   = 3
arr_flat[arr_flat==7]   = 3
arr_flat[arr_flat==8]   = 4 # OB
arr_flat[arr_flat==9]   = 4
arr_flat[arr_flat==10]  = 4
arr_flat[arr_flat==11]  = 4
arr_flat[arr_flat==12]  = 4
ufacies = sort(unique(arr_flat))
cat("Unique facies are ", ufacies, "\n")
nfacies = length(ufacies) 
```

# Proportions
Calculate facies global proportions

```{r proportions2}
cat("Number of Facies:", nfacies)
cat("Facies Proportions:\n")
round(100*table(arr_flat)/(nx*ny),2)
```

# Import this array in gstlearn
The Grid file is created first (defining the origin, mesh size and count). Then the facies variable is added giving the name and the locator. 

```{r create_db}
ndim = 2
gnx = nx
gny = nz
gdx = mesh
gdy = dz
defineDefaultSpace(ESpaceType_RN(), ndim)
grid = DbGrid_create(c(gnx,gny),c(gdx,gdy),c(0,0))
grid$addColumns(arr_flat, "fac", ELoc_Z())
att_simu = grid$getLastUID()
if (verbose)
{
    dbfmt = DbStringFormat()
    dbfmt$setFlags(flag_extend=TRUE, flag_stats=TRUE)
    grid$display(dbfmt)
}
ggplot() + plot.grid(grid,nameRaster="fac",flagLegendRaster=TRUE)
```

# Transform categorical variable into indicators

```{r transition_proba}
limits = Limits(nfacies)
if (verbose) {
    limits$display()
}
err = limits$toIndicator(grid, "fac")
grid$display()
ggplot() + plot.grid(grid,nameRaster="Indicator.fac.Class.4",flagLegendRaster=TRUE)
```

# Variogram of indicators

```{r create_db}
variop = VarioParam()
dir1 = DirParam(40,gdx)
dir1$setGrincr(c(1,0))
dir2 = DirParam(20,gdy)
dir2$setGrincr(c(0,1))
variop$addDir(dir1)
variop$addDir(dir2)
vario = Vario(variop, grid)
err = vario$compute(ECalcVario_TRANS1())
vario$display()
```

# Transition probabilities

```{r transition_proba}
ndir = vario$getDirectionNumber()
for (idir in 0:(ndir-1)) {
    for (ifac in 0:(nfacies-1)) {
        for (jfac in 0:ifac) {
            if (ifac != jfac) {
                hh = vario$getHhVec(idir,ifac,jfac)
                gg = vario$getGgVec(idir,ifac,jfac)
                plot(hh,gg, type='l', main=paste0("Transition probabilities in dir #",idir+1," leaving F",ifac+1," / entering F",jfac+1))
            }
        }
    }
}
```

# Connected Components
Calculate the connected components for each facies.

```{r connected_components}
eps = 0.5
compnum = grid$getColumnByUID(att_simu)
for (fac in ufacies) {
    vmin = fac - eps
    vmax = fac + eps
    image   = morpho_double2image(c(gnx,gny),compnum,vmin,vmax);
    compnew = morpho_labelling(0,0,image,-1)
    sizes   = morpho_labelsize(0,image)
    df$compnew = compnew
    
    print(ggplot(df, aes(x, y, fill=compnew)) +
          geom_raster() + coord_fixed() +
          ggtitle(paste0("Connected components for F",fac)))
    
    hist(sizes, breaks=50, main=paste0("Size of connected component for F",fac))
}
```

