---
title: "Hermite Formulae"
author: "D. Renard, N. Desassis, X. Freulon"
date: "May 17th 2022"
output:
  pdf_document: default
  html_document: default
editor_options: 
  chunk_output_type: console
---

<!-- SUMMARY: Perform the elementary calculation based on Hermite polynomials  -->

<!-- CATEGORY: Basic_Calculations -->

```{r setup, include=FALSE, echo=FALSE}
knitr::opts_chunk$set(echo = TRUE)
rm(list=ls())
library(gstlearn)
library(ggplot2)
```

# General Introduction

The objective of this document is to provide some guides for manipulation of Hermite polynomials.
In particular, it will give some hints for calculating of complex expressions
(e.g. metal above cutoff) and be double checked using a Monte Carlo approach.

It will also demonstrate the quality of the polynomial expansion for a given known distribution of the data (i.e. lognormal).
Finally it will serve in demonstrating the ability of representing any function through a Hermite polynomial expansion.

# Part I

```{r child="../../references/Hermite_Polynomials.md"}
```

# Evaluation on the lognormal case

Processing parameters

```{r Initial-setup, echo=TRUE, eval=TRUE}
nbpoly = 100
nbsimu = 1000
params = list(nbpoly=nbpoly, nbsimu=nbsimu, yc=1.5)
```

We consider the Lognormal case (where the anamorphosis is known exactly).

```{r lognormal, echo=TRUE, eval=TRUE}
m     <- 2.5
sigma <- 0.5
hn     = hermiteLognormal(m, sigma, params$nbpoly)
```

We calculate the various functions (expectation, ore and metal quantity, for expectation and standard deviation) for different values of $y$ varying from -3 to 3 and different values of $r$ varying from 0 to 1. The results are stored in a Db (organized as a matrix) whose parameters are given next.

```{r Grid-Definition}
y0  = -3
dy  = 0.1
dr  = 0.1
y   = seq(-3,3,by=dy)
ny  = length(y)
r   = seq(1,0,by=-dr)
nr  = length(r)
```

# Comparisons

We evaluate the variable, the ore and metal quantity above a cutoff, in conditional expectation or in conditional standard deviation.

```{r Internal-Functions, echo = FALSE}
# Function to generate the title
get.title <- function(type,calcul,flag.est,params)
{
  if (calcul == 1)
  {
    if (flag.est)
      title = "E(Z)"
    else
      title = "St.dev.(Z)"
  }
  else if (calcul == 2)
  {
    if (flag.est)
      title = paste("Ore above Cutoff (",params$yc,")")
    else
      title = paste("Ore St.Dev. above Cutoff (",params$yc,")")
  }
  else
  {
    if (flag.est)
      title = paste("Metal above Cutoff (",params$yc,")")
    else
      title = paste("Metal St.Dev. above Cutoff (",params$yc,")")
  }
  
  if (type == 1)
    title = paste(title," using Hermite (",params$nbpoly,")")
  else if (type == 2)
    title = paste(title," using Monte Carlo (",params$nbsimu,")")

  title
}

# Function to calculate a quantity (Cond. Exp., Tonnage, Metal) by Hermite or Monte Carlo
# for different parameters of the parameters
# The results are provided on a regular Db
calculate.matrix <- function(type,calcul,flag.est,params)
{
  # Setting the result grid
  
  db = DbGrid_create(nx=c(ny,nr),x0=c(y0,0),dx=c(dy,dr))
  
  if (type == 1)
  {
    # Calculation using Hermite expansion
    
    if (calcul == 1)
    {
      if (flag.est)
        psi = hermiteCondExp(db["x1"],db["x2"],hn)
      else
        psi = hermiteCondStd(db["x1"],db["x2"],hn)
    }
    else if (calcul == 2)
    {
      if (flag.est)
        psi = hermiteIndicator(params$yc,db["x1"],db["x2"])
      else
        psi = hermiteIndicatorStd(params$yc,db["x1"],db["x2"])
    }
    else if (calcul == 3)
    {
      if (flag.est)
        psi = hermiteMetal(params$yc,db["x1"],db["x2"],hn)
      else
        psi = hermiteMetalStd(params$yc,db["x1"],db["x2"],hn)
    }
    else
      stop("Wrong value for 'calcul'")
  }
  else if (type == 2)
  {
    # Calculation using Monte Carlo
    
    if (calcul == 1)
    {
      if (flag.est)
        psi = MCCondExp(db["x1"],db["x2"],hn,params$nbsimu)
      else
        psi = MCCondStd(db["x1"],db["x2"],hn,params$nbsimu)
    }
    else if (calcul == 2)
    {
      if (flag.est)
        psi = MCIndicator(params$yc,db["x1"],db["x2"],params$nbsimu)
      else
        psi = MCIndicatorStd(params$yc,db["x1"],db["x2"],params$nbsimu)
    }
    else if (calcul == 3)
    {
      if (flag.est)
        psi = MCMetal(params$yc,db["x1"],db["x2"],hn,params$nbsimu)
      else
        psi = MCMetalStd(params$yc,db["x1"],db["x2"],hn,params$nbsimu)
    }
    else
      stop("Wrong value for 'calcul'")
  }
  else
    stop("Wrong value for 'type'")
  
  uid = db$addColumns(as.numeric(psi),"psi")
  db
}

# Representation of the grid
draw.matrix <- function(type,calcul,flag.est,params,flag.cut=TRUE)
{
  title = get.title(type,calcul,flag.est,params)
  db = calculate.matrix(type,calcul,flag.est,params)
  p = ggplot()
  p = p + plot.grid(db,nameRaster="psi")
  p = p + plot.decoration(title=title)
  if (flag.cut) p <- p + geom_vline(xintercept = params$yc)
  ggPrint(p)
}

# Representation of a scatter plot
draw.correlation <- function(calcul,flag.est,params,flag.iso=TRUE)
{
  title = get.title(0,calcul,flag.est,params)
  db1 = calculate.matrix(type=1,calcul,flag.est,params)
  db2 = calculate.matrix(type=2,calcul,flag.est,params)
  p = ggplot()
  p = p + plot.correlation(db1,"psi","psi",db2, flagDiag=TRUE, bins=100)
  p = p + plot.decoration(title=title,xlab="Hermite",ylab="Monte Carlo")
  ggPrint(p)
}
```

Calculating all elements using either Hermite expansion or Monte Carlo Simulations

```{r Grids-for-Estimation-and-Variances, eval=FALSE}
draw.matrix(type=1,calcul=1,flag.est=TRUE,params,flag.cut=FALSE)
draw.matrix(type=2,calcul=1,flag.est=TRUE,params,flag.cut=FALSE)
draw.correlation(calcul=1,flag.est=TRUE,params)

draw.matrix(type=1,calcul=2,flag.est=TRUE,params)
draw.matrix(type=2,calcul=2,flag.est=TRUE,params)
draw.correlation(calcul=2,flag.est=TRUE,params)

draw.matrix(type=1,calcul=3,flag.est=TRUE,params)
draw.matrix(type=2,calcul=3,flag.est=TRUE,params)
draw.correlation(calcul=3,flag.est=TRUE,params)

draw.matrix(type=1,calcul=1,flag.est=FALSE,params,flag.cut=FALSE)
draw.matrix(type=2,calcul=1,flag.est=FALSE,params,flag.cut=FALSE)
draw.correlation(calcul=1,flag.est=FALSE,params)

draw.matrix(type=1,calcul=2,flag.est=FALSE,params)
draw.matrix(type=2,calcul=2,flag.est=FALSE,params)
draw.correlation(calcul=2,flag.est=FALSE,params)

draw.matrix(type=1,calcul=3,flag.est=FALSE,params)
draw.matrix(type=2,calcul=3,flag.est=FALSE,params)
draw.correlation(calcul=3,flag.est=FALSE,params)
```

# Part II

```{r child="../../references/Hermite_Lognormal.md"}
```

The next lines define some functions which can be explitely written in the lognormal case

```{r hermite_test, echo=TRUE, eval=TRUE}
GM_hermite <- function(y, nh){
  h <- matrix(NaN, nrow = length(y), ncol = 1+nh)
  h[,1] <- 1.0
  h[,2] <- -y
  for (i in 2:nh)
  {
    h[,1+i] <- -y*h[,i]/sqrt(i) - sqrt((i-1)/i)*h[,i-1]
  }
  h
}

# fonct 
GM_LN_psi <- function(lambda, nh){
  (-lambda)^{0:nh} / sqrt(factorial(0:nh))
}
GM_LN_Cov_theo <- function(lambda){
  function(rho) {exp(lambda^2 * rho) - 1}
}
GM_LN_Cov_nh <- function(lambda, nh)
{
  psi <- GM_LN_psi(lambda = lambda, nh = nh)
  foo <- function(rho) {outer(X = rho, Y = 1:nh, FUN = function(r,n){r^n}) %*% psi[-1]^2}
}
```

The next paragraph initiates a data set and runs the previous functions. The output will serve as reference for comparison with rgstlearn functions.

```{r}
np <- 10000
nh <- 100
lambda <- 1.0
set.seed(123)
# anamorphosis in the lognormal case (theoretical)
yy <- qnorm(p = (1:np)/(np+1), mean = 0, sd = 1.0)
aymin = range(yy)[1]
aymax = range(yy)[2]
zz_theo <- exp(lambda*yy - lambda^2/2)
```

## Perfoming the tests with the functions programmed internally

Anamorphosis in the lognormal case (Hermite pol.)

```{r ana_test, echo=TRUE, eval=TRUE}
zz_hm   <- GM_hermite(y = yy, nh = nh) %*% GM_LN_psi(lambda = lambda, nh = nh)

p = ggplot()
p = p + plot.XY(yy, zz_theo, color = "red", flagLine=TRUE)
p = p + plot.XY(yy, zz_hm, color = "blue", flagLine=TRUE, linetype="dashed")
p = p + plot.decoration(xlab = "Y", ylab = "Z", title = "Lognormal transform")
ggPrint(p)
```


```{r}
p = ggplot()
p = p + plot.XY(zz_theo, zz_hm, color = "blue", flagLine=TRUE)
p = p + plot.decoration(xlab = "theoretical", ylab = "Hermite pol.", 
                        title = "Lognormal transform")
ggPrint(p)
```

Covariance

```{r}
r <- seq(from = 0, to =1, by = 0.01)
C_theo <- GM_LN_Cov_theo(lambda = lambda)
C_nh   <- GM_LN_Cov_nh(lambda = lambda, nh = nh)
p = ggplot()
p = p + plot.XY(r, C_theo(r), color = "red", flagLine=TRUE)
p = p + plot.XY(r,C_nh(r), color = "blue", linetype="dashed", flagLine=TRUE)
p = p + plot.decoration(xlab = "Covariance of Y", ylab = "Covarariance of Z", 
                        title = "Lognormal transform")
ggPrint(p)
```

## Performing similar tests with gstlearn

```{r rgeostats_test, echo=TRUE, eval=TRUE}
db <- Db_createFromSamples(np,ELoadBy_SAMPLE(),zz_theo)
db$setNameByUID(db$getLastUID(),"zz_theo")

ana = AnamHermite_create(nh)
err = ana$fit(db, name="zz_theo")

# Patch the Hermite coefficients to have exact numerical values
localPsi <- GM_LN_psi(lambda = lambda, nh = nh-2)
err = ana$setPsiHns(localPsi)
ana$display()

p = ggplot()
p = p + plot.anam(ana, aymin=aymin, aymax=aymax)
p = p + plot.XY(yy, zz_theo, color="red", linetype="dashed", flagLine=TRUE)
p = p + plot.decoration(title="Lognormal Transform")
ggPrint(p)

#legend("topleft", legend = c("theoretical", "gstlearn"), col = c("red", "black"), lty = c(1,1))
```


```{r,eval=FALSE}
zz_rg = ana$gaussianToRawVector(yy)
p = plot.XY(zz_theo, zz_rg, flagLine=TRUE, color="blue")
p = plot.decoration(xlab="theoretical", ylab="Hermite pol.", title="Lognormal Transform")
ggPrint(p)
#legend("topleft", legend = c("transform", "Y=X"), col = c("blue", "red"), lty = c(1,2))
```

Covariance

```{r}
mod_y <- Model(nvar=1, ndim=2)
mod_y$addCovFromParam(ECov_SPHERICAL(),range=1,sill=1)
h     <- seq(from = 0, to =1, by = 0.01)
r = mod_y$evalIvarNpas(h)
C_theo<- GM_LN_Cov_theo(lambda = lambda)
cov_theo<- C_theo(rho = r) 

# Creation of the lognormal model (il faudrait pouvoir créer un modèle gaussien à partir de la covariance de la gaussienne mod_Y et de l'anamorphose gaussienne ana)
# mod_Z <- model.create.GM(model = mod_Y, anam = ana)
# cov_rg<- model.eval(model = mod_Z, h = h)

model.eval.GM <- function(model, anam, h){
  r = model$evalIvarNpas(h)
  psi <- anam$getPsiHns()
  nh <- length(psi) - 1
  outer(X = r, Y = 1:nh, FUN = function(rho,n){rho^n}) %*% (psi[-1])^2
}
cov_rg <- model.eval.GM(model = mod_y, anam = ana, h = h)

p = ggplot()
p = p + plot.XY(r, cov_theo, color = "red", flagLine=TRUE) 
p = p + plot.XY(r, cov_rg, color = "blue", linetype = "dashed", flagLine=TRUE)
p = p + plot.decoration(xlab = "Covariance of Y", ylab = "Covariance of Z", 
                        title = "Lognormal transform")
ggPrint(p)

#legend("topleft", legend = c("theoretical", "gstlearn"), col = c("red", "blue"), lty = c(1,2))
```

The empirical anamorphosis does not reproduce the lognormal transform and the Hermite coefficient need to be altered.

# Part III

```{r, child="../../references/Hermite_Truncated.md"}
```

```{r}
np <- 1000
nh <- 1000
yc <- qnorm(0.25)
yy <- qnorm(p = (1:np)/(np+1))
hm <- GM_hermite(y = yy, nh = nh-1)
```

Floor variable Z = max(yc, Y)

```{r}
psi = hermiteCoefLower(yc, nh)
zz <- hm %*% psi
ylim = c(-3.5, 3.5)
p = ggplot()
p = p + plot.XY(yy, zz, color = "red", flagLine=TRUE)
p = p + plot.XY(yy, pmax(yc, yy), color = "blue", linetype = "dashed", flagLine=TRUE)
p = p + plot.decoration(xlab = "Y", ylab = "Z", 
                        title = paste0("cut = ",round(yc,2), " nh = ", nh))
p = p + plot.geometry(ylim=ylim)
ggPrint(p)

#legend("topleft", legend = c("theoretical", "Hermit. pol."), col = c("blue", "red"), lty = #c(1,2))
```

QC Indicator

```{r}
psi = hermiteIndicatorLower(yc, nh)
zz <- hm %*% psi

p = ggplot()
p = p + plot.XY(yy, zz, color = "red", flagLine=TRUE)
p = p + plot.decoration(xlab = "Y", ylab = "Indicator", 
                        title = paste0("cut = ",round(yc,2), " nh = ", nh)) 
p = p + plot.geometry(ylim = range(-0.5, 1.5))
p = p + plot.XY(yy, as.numeric(yy >= yc), color = "blue", linetype="dashed", flagLine=TRUE)
ggPrint(p)

#legend("bottomright", legend = c("theoretical", "Hermit. pol."), col = c("blue", "red"), lty = #c(1,2))
```
