---
title: "Starting gstlearn with R"
author: "RGeostats Team"
date: "September 2019"
output: pdf_document
---


```{r setup,echo=FALSE}
library(knitr)
```
# Introduction

This document gives the template preliminary statements that need to be
specified in order to launch gstlearn under R.
This is a temporary solution ... unitl rgstlearn becomes a package.

The current version consists in loading the loader and the contents
You must adapt the following lines$display before launching it.

```{r}
gstlearn_path = file.path("/home","drenard","project_gstlearn","gstlearn")
source(file.path(gstlearn_path,"doc","dev","rgstlearn_loader.R"))
load_gstlearn(file.path(gstlearn_path,"build","r","Release","gstlearn"))
```

# Demonstration

This script:
- constructs a regular grid
- load values generated randomly
    
```{r}
# Grid size
nx = 60
ny = 30
mygrid = DbGrid_create(c(nx,ny))

# Add a uniform random field
var = ut_vector_simulate_uniform(nx * ny)
uid = mygrid$addColumns(var, "var1")

# Display the current contents of the Data Base
mygrid$display()
```

The grid is plotted (this uses a function provided in plot.r)

```{r}
# Display the field
plot.grid(mygrid, "var1")
```
