---
  title: "Kriging linear combinations"
  author: "gstlearn core team"
  date: "January 31, 2024"
  output:
    pdf_document: default
    html_document: default
    editor_options: 
      chunk_output_type: console
    chunk_output_type: console
---

# Introduction

This notebook presents examples for the co-kriging of linear combinations. 
Starting from a set of variables $\mathbf{Z}(\alpha) = (Z_i(\alpha))_{i \in 1:n}$, the task is to compute the co-kriging of a linear combination of the variable at location $\mathbf{o}$:
$$
(\beta^T \mathbf{Z}(\mathbf{o}))^{K} = \sum_{i \in 1:n} \sum_{\alpha \in A_i} \lambda_i^{\alpha} Z_i(\alpha)
$$
Kriging is linear, hence $(\beta^T \cdot \mathbf{Y}(\mathbf{o}))^{K} = \sum_{i \in 1:p} \beta_i \times Y_i(\mathbf{o})^{K}$ but
the variance of the kriging error is not easily computable. However, the *matLC* option of the **kriging** function offers the possibility to compute directly the estimate and its kriging variance of any linear combinations.
The method is evaluated on a simulated data set.

The following cases are considered:

* Ordinary cokriging using a IRF-0 model (Intrinsic Random Function of order 0 with unknown means)
* Simple cokriging using a SOS model (Second Order Stationary random function with known means)
* Homotopic data set (both variables are defined for each location)
* Heterotopic data set (variables may be defined or not at each location)
* Direct kriging of the linear combination or linear combination of the estimates

In all cases, the results seem consistent with the theory.

Kriging systems when kriging a linear combination are given below:

## Simple kriging with a second order stationary model (SK/SOS)

The expression of the estimator is:
$$ 
(\beta^T \, \mathbf{Z_{(o)}})^{SK} = \beta^T \mathbf{m} +\sum_{i \in 1:n} \sum_{\alpha \in A_i} \lambda^{\alpha} (Z_i(\alpha) - m_i)
$$

There is no bias as:
$$
\mathbb{E}\{(\beta^T \, \mathbf{Z_{(o)}})^{SK}\} = \beta^T \mathbf{m} = \mathbb{E}\{\beta^T \, \mathbf{Z_{(o)}}\}.
$$

The variance of the kriging error is:
$$
\begin{array}{cl} 
\text{Var}\{\beta^T \, \mathbf{Z_{(o)}} - (\beta^T \, \mathbf{Z_{(o)}})^{SK}\} = &
\beta^T \, \text{Var}\{\mathbf{Z_{(o)}}\} \, \beta \\
 & - 2 \sum_{k,i \in 1:n} \sum_{\alpha \in A_i} \beta_k \, \lambda^{\alpha} \text{Cov}(Z_k(o),Z_i(\alpha)) \\
 & + \sum_{i,j \in 1:n} \sum_{\alpha \in A_i, \alpha' \in A_j} \beta_k \, 
\lambda^{\alpha} \lambda^{\alpha'} \text{Cov}(Z_i(\alpha),Z_j(\alpha')) 
\end{array}
$$

Minimizing this variance, we gets the kriging system of for the estimation of the linear combination:
$$
\left(
\text{Cov}(Z_{i}(\alpha),Z_{j}(\alpha')) 
\right) \Lambda =  \sum_{k\in 1:n} \beta_k \text{Cov}(Z_k(o), Z_i(\alpha))
$$

The variance of the kriging error boils down to:
$$
\text{Var}\{\beta^T \, \mathbf{Z_{(o)}} - (\beta^T \, \mathbf{Z_{(o)}})^{SK}\} = 
\beta^T \, \text{Var}\{\mathbf{Z_{(o)}}\} \, \beta - 
\sum_{k,i \in 1:n} \sum_{\alpha \in A_i} \beta_k \text{Cov}(Z_k(o),Z_i(\alpha)) \lambda^{\alpha}
$$


## ordinary kriging with a Intrinsic Random Fonction of order 0 (OK/IRF0)

The expression of the estimator is:
$$ 
(\beta^T \, \mathbf{Z_{(o)}})^{OK} = \sum_{i \in 1:n} \sum_{\alpha \in A_i} \lambda^{\alpha} Z_i(\alpha)
$$

The no-bias conditions are:
$$
\sum_{\alpha \in A_i} \lambda^{\alpha} = \beta_i, \text{ for } \forall i \in 1:n.
$$

The variance of the kriging error is:
$$
\begin{array}{cl}
\text{Var}\{\beta^T \, \mathbf{Z_{(o)}} - (\beta^T \, \mathbf{Z_{(o)}})^{OK}\} = &
\beta^T \, \text{Var}\{\mathbf{Z_{(o)}}\} \, \beta \\
 &  - 2 \sum_{k\in 1:n} \sum_{i, \alpha \in A_i} \beta_k \text{Cov}(Z_k(o),Z_i(\alpha)) \, \lambda^{\alpha} \\
 &   +
\sum_{i, \alpha \in A_i} \sum_{j, \alpha' \in A_j} \lambda^{\alpha}  \text{Cov}(Z_i(\alpha),Z_j(\alpha')) \lambda^{\alpha'}
\end{array}
$$

Minimizing the Lagrangian, 
$$
\mathcal{L}(\Lambda, \mu) = \text{Var}\{\beta^T \, \mathbf{Z_{(o)}} - (\beta^T \, \mathbf{Z_{(o)}})^{OK}\} + \sum_{k \in 1:n} \mu_k (\sum_{\alpha \in A_k} \lambda^{\alpha} - \beta_k)
$$

we gets the kriging system of for the estimation of the linear combination:
$$
\left[
\begin{array}{cc}
\text{Cov}(Z_i(\alpha), Z_j(\alpha')) & \mathbf{1}_{\alpha \in A_j} \\
\mathbf{1}_{\alpha' \in A_i} & \mathbf{0}
\end{array}
\right]
\left[
\begin{array}{c}
\lambda \\
\mu
\end{array}
\right] =
\left[
\begin{array}{c}
\sum_{k \in 1:n} \beta_k\text{Cov}(Z_k(o), Z_i(\alpha)) \\
\beta_i
\end{array}
\right]
$$


The variance of the kriging error boils down to:
$$
\text{Var}\{\beta^T \, \mathbf{Z_{(o)}} - (\beta^T \, \mathbf{Z_{(o)}})^{OK}\} = 
\beta^T \, \text{Var}\{\mathbf{Z_{(o)}}\} \, \beta - 
\sum_{k \in 1:n} \sum_{i,\alpha \in A_i} \beta_k \text{Cov}(Z_k(o),Z_i(\alpha))\, \lambda^{\alpha} - \sum_{k \in 1:n} \beta_k \, \mu_k
$$

## Initialisation

```{r setup, include=FALSE, echo=TRUE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(align="center")
rm(list=ls())
library(gstlearn)
library(ggplot2)
library(ggpubr)
set.seed(1235)
OptDbg_setReference(0)
# Defining a set of statistics
opers = EStatOption_fromKeys(c("NUM","MINI","MAXI","MEAN","STDV"))
flag.verbose = FALSE
ssp = 0.005 # sub sampling proportion
```

## Simulation of a reference data set

We create a reference data set using the the Turning Bands method (*simtub*):

* the distribution is log normal,

* the stationary covariance function is exponential.

5\% of the two dimensional grid is sampled and considered as the reference data set.

```{r data_set, echo=TRUE, eval=TRUE}
# parameters for the simulation
m   = c(1, 2)
sig = c(0.25, 0.75)
rho = 0.7
CO = matrix(c(1, rho, rho, 1), 2, 2)
# initialization of the grid
grd = DbGrid_create(x0=c(0.0,0.0), dx=c(0.01,0.01), nx=c(100,100))

# simulation from a model
model = Model_createFromParam(type = ECov_EXPONENTIAL(), range=0.2, sills = CO)
err = simtub(dbin = NULL, dbout = grd, model = model, nbsimu = 1)
err = grd$setName("Simu.1", "Y1")
err = grd$setName("Simu.2", "Y2")
grd["Z1"] = m[1] * exp(sig[1] * grd["Y1"] - sig[1]**2 / 2)
grd["Z2"] = m[2] * exp(sig[2] * grd["Y2"] - sig[2]**2 / 2)

# Data set (ssp proportion of the grid)
data = Db_createFillRandom(ndat = ssp*grd$getSampleNumber(), nvar = 0)
err  = data$setName("x-1", "x1")
err  = data$setName("x-2", "x2")
err  = migrate(dbin = grd, name = "Z1", dbout = data, dmax = c(0.1, 0.1), 
               namconv = NamingConvention(""))
err  = migrate(dbin = grd, name = "Z2", dbout = data, dmax = c(0.1, 0.1), 
               namconv = NamingConvention(""))

# plots
p1 = ggDefaultGeographic() + 
  plot.grid(grd, nameRaster = "Z1", palette = "Spectral") +
  plot.point(data, colour = "red") +
  plot.decoration(title="Z1")
p2 = ggDefault() +
  plot.hist(grd, name='Z1',  bins = 25, bg = "orange", color= "gray") +
  plot.decoration(xlab = "Raw variable", title="Histogram of the initial variable Z1")
p3 = ggDefaultGeographic() + 
  plot.grid(grd, nameRaster = "Z2", palette = "Spectral") +
  plot.point(data, colour = "red") +
  plot.decoration(title="Z2")
p4 = ggDefault() +
  plot.hist(grd, name='Z2',  bins = 25, bg = "orange", color= "gray") +
  plot.decoration(xlab = "Raw variable", title="Histogram of the initial variable Z2")
ggarrange(p1, p2, p3, p4, ncol = 2, nrow = 2)

# defining the linear combination
beta = c(1, -3) # coefficients and computing the linear combination
nm_var = c("Z1", "Z2")
nm_LC  = c("LC", "Z1")
matLC  = matrix(c(beta, 1, 0), nrow = 2, ncol = 2, byrow = TRUE)
colnames(matLC) <- nm_var
rownames(matLC) <- nm_LC

# computing the linear computation
data["LC"] = matrix(data$getColumns(colnames(matLC)), nrow = data$getSampleNumber(), 
                    ncol = dim(matLC)[2]) %*% t(matLC)[,1]
grd["LC"]  = matrix(grd$getColumns(colnames(matLC)),  nrow = grd$getSampleNumber(),  
                    ncol = dim(matLC)[2]) %*% t(matLC)[,1]

# statistics
tab = rbind(
  dbStatisticsMono(grd, names = c("Z*", "LC"), opers = opers)$toTL(),
  dbStatisticsMono(data, names = c("Z*", "LC"), opers = opers)$toTL())
rownames(tab) <- c(
  "grid-Z1", "grid-Z2", "grid-LC", 
  "samples-Z1", "samples-Z2", "samples-LC")
knitr::kable(tab, caption = "Statistics on the reference data set", digits = 4)
```

## Global parameters for spatial modelling 

```{r vario_param, echo=TRUE, eval=TRUE}

# Variogram parameters (omnidirectional)
varioParam = VarioParam_createOmniDirection(npas=10, dpas=0.05)
# Variogram basic structures
types = ECov_fromKeys(c("NUGGET", "EXPONENTIAL", "EXPONENTIAL"))
# types = ECov_fromKeys(c("EXPONENTIAL"))
# weighted proportional to the number of pairs and inverse proportional to the distance
opt   = Option_AutoFit()
err   = opt$setWmode(2) 

#define neighborhood
neigh = NeighMoving_create(nmaxi=30, radius=0.5)

# Display the kriging system for a fixed target when using kriging
OptDbg_setReference(0)

# function to compute the model of the linear combination (using a square matrix) 
# TODO: generalize to rectangular matrix?
model_of_LC <- function(model, A) {
  stopifnot(dim(A)[1] == dim(A)[2])
  nv = model$getVariableNumber()
  stopifnot(nv == dim(A)[1])
  ns = model$getCovaNumber()
  model_LC = model$clone()
  for (k in 1:ns) {
    CC = A %*% model$getSillValues(k-1)$toTL() %*% t(A)
    for (i in 1:nv) {
      for (j in 1:nv) {
        err = model_LC$setSill(icov = k-1, ivar=i-1, jvar=j-1, value = CC[i, j])
        }
      }
    }
  model_LC
}
```

# Ordinary kriging (OK)

## Intrinsic Random Function of order 0

```{r OK_modeling, echo=TRUE, eval=TRUE}
# Locate Z
err = data$setLocators("Z*", locatorType = ELoc_Z(), cleanSameLocator = TRUE)
# Compute variogram
var_Z = Vario_computeFromDb(varioParam, data)
# fit model
mod_Z_IRF0 = Model()
err   = mod_Z_IRF0$fit(var_Z, types = types, mauto = opt)

# setting the drift "unknown mean" for Ordinary Kriging
err = mod_Z_IRF0$setDriftIRF(order = 0)

# plot
multi.varmod(var_Z, mod_Z_IRF0, flagLegend=TRUE)

if(flag.verbose) {
  mod_Z_IRF0
}
```

## Ordinary kriging of the raw variables

```{r OK_estimate, echo=TRUE, eval=TRUE}
err = grd$deleteColumns(paste("OK", nm_var, "*", sep = "."))
err = grd$deleteColumns(paste("LC", "OK", "*", sep = "."))
if(flag.verbose) {
  print(paste0(">>> Ordinary kriging of the raw variables"))
}
err = data$setLocators(c("Z1", "Z2"), ELoc_Z(), cleanSameLocator=TRUE)
err = kriging(data, grd, mod_Z_IRF0, neigh, 
              flag_est = TRUE, flag_std = TRUE,
              namconv = NamingConvention("OK"))

# computing the LC of kriging
grd["LC.OK.estim"]  = matrix(grd$getColumns(names = paste("OK", colnames(matLC), "estim", sep = ".")),
                    nrow = grd$getSampleNumber(),  ncol = dim(matLC)[2]) %*% t(matLC)[,1]

# statistics
vn  = c("Z1", paste("OK", "Z1", c("estim", "stdev"), sep = "."),
        "Z2", paste("OK", "Z2", c("estim", "stdev"), sep = "."))
tab = dbStatisticsMono(grd, vn, opers = opers)$toTL()
knitr::kable(tab, caption = "Statistics on the grid for Ordinary Kriging with the initial model", digits = 4)

# plot of estimated values
p1 = ggDefaultGeographic() + 
    plot.grid(grd, nameRaster =  paste("OK", "Z1", "estim", sep = "."),
            palette = "Spectral", flagLegendRaster = TRUE, legendNameRaster = "estim") +
    plot.decoration(xlab = "", ylab = "", title = paste0("Ordinary coK for Z1"))
p2 = ggDefaultGeographic() + 
    plot.grid(grd, nameRaster =  paste("OK", "Z1", "stdev", sep = "."),
            palette = "Spectral", flagLegendRaster = TRUE, legendNameRaster = "stdev") +
    plot.decoration(xlab = "", ylab = "", title = paste0("Std. of Ordinary coK for Z1"))
p3 = ggDefaultGeographic() + 
    plot.grid(grd, nameRaster =  paste("OK", "Z2", "estim", sep = "."),
            palette = "Spectral", flagLegendRaster = TRUE, legendNameRaster = "estim") +
    plot.decoration(xlab = "", ylab = "", title = paste0("Ordinary coK for Z2"))
p4 = ggDefaultGeographic() + 
    plot.grid(grd, nameRaster =  paste("OK", "Z2", "stdev", sep = "."),
            palette = "Spectral", flagLegendRaster = TRUE, legendNameRaster = "stdev") +
    plot.decoration(xlab = "", ylab = "", title = paste0("Std. of Ordinary coK for Z2"))
ggarrange(p1, p2, p3, p4, ncol = 2, nrow = 2)

# plot Z vs. Z*
p1 = ggDefault() + 
  plot.correlation(grd, namex = paste("OK", "Z1", "estim", sep = "."), namey = "Z1", asPoint = TRUE, 
                   flagDiag = TRUE, diagColor = "red", flagSameAxes = TRUE) +
  plot.decoration(xlab = "Estimated values", ylab = "Actual values", title = "Ordinary coK of Z1")
p2 = ggDefault() + 
  plot.correlation(grd, namex = paste("OK", "Z2", "estim", sep = "."), namey = "Z2", asPoint = TRUE, 
                   flagDiag = TRUE, diagColor = "red", flagSameAxes = TRUE) +
  plot.decoration(xlab = "Estimated values", ylab = "Actual values", title = "Ordinary coK of Z2")
ggarrange(p1, p2, ncol = 2, nrow = 1)
```

## Ordinary kriging of a linear combination of the raw variables

```{r OK_LC_estimate, echo=TRUE, eval=TRUE}
err = grd$deleteColumns("OK-LC*")
if(flag.verbose) {
  print(paste0(">>> Ordinary kriging of a linear combination of the variables"))
}

mod_LC_IRF0 = model_of_LC(mod_Z_IRF0, A = matLC)

# modelling the linear combination
# Locate Z
err = data$setLocators(nm_LC, locatorType = ELoc_Z(), cleanSameLocator = TRUE)
# computing the model of the LC from the initial model
var_LC = Vario_computeFromDb(varioParam, data)
# setting the drift "unknown mean" for Ordinary Kriging
err = mod_LC_IRF0$setDriftIRF(order = 0)

# plot
multi.varmod(var_LC, mod_LC_IRF0, flagLegend=TRUE)
if(flag.verbose) {
  mod_LC_IRF0
}

# kriging the LC variables
err = data$setLocators(nm_LC, ELoc_Z(), cleanSameLocator=TRUE)
err = kriging(data, grd, mod_LC_IRF0, neigh, 
              flag_est = TRUE, flag_std = TRUE,
              namconv = NamingConvention("OK-LC"))

# statistics
vn = c(
  paste(c("OK.Z1", "OK-LC.Z1", "OK-LC.LC", "LC.OK"), "estim", sep = "."),
  paste(c("OK.Z1", "OK-LC.Z1", "OK-LC.LC"), "stdev", sep = ".")
)
tab = dbStatisticsMono(grd, names = vn, opers = opers)$toTL()
knitr::kable(tab, caption = "Statistics on the grid for Ordinary Kriging with the LC model", 
             digits = 4)

# plot Z vs. Z*
p1 = ggDefault() + 
  plot.correlation(grd, 
                   namex = paste("OK", "Z1", "estim", sep = "."), 
                   namey = paste("OK-LC", "Z1", "estim", sep = "."), 
                   asPoint = TRUE, flagDiag = TRUE, diagColor = "red", flagSameAxes = TRUE) +
  plot.decoration(xlab = "Initial estimate", ylab = "Estimate with the LC model", 
                  title = "Ordinary coK of Z1")
p2 = ggDefault() + 
  plot.correlation(grd, 
                   namey = paste("OK-LC", "LC", "estim", sep = "."), 
                   namex = paste("LC", "OK", "estim", sep = "."), 
                   asPoint = TRUE, flagDiag = TRUE, diagColor = "red", flagSameAxes = TRUE) +
  plot.decoration(xlab = "LC of the estimates", ylab = "Estimate with the LC model", 
                  title = "Ordinary coK of CL")
ggarrange(p1, p2, ncol = 2, nrow = 1)
```

## Direct computation of linear combination kriging

```{r dOK_LC_estimate, echo=TRUE, eval=TRUE}
err = grd$deleteColumns("dOK*")
if(flag.verbose) {
  print(paste0(">>> Direct computation of linear combination kriging"))
}

# the R matrix matLC has to be converted into a gstlearn MatrixRectangular:
# MatrixRectangular_createFromVD(beta, nrow=1, ncol=length(beta))
# OR
# beta |> matrix(nrow=1, ncol=length(beta)) |> fromTL()
# OR
# fromTL(matrix(beta, nrow=1, ncol=length(beta)))

# kriging LC variables
err = data$setLocators(nm_var, ELoc_Z(), cleanSameLocator=TRUE)
err = kriging(data, grd, mod_Z_IRF0, neigh, 
              matLC = fromTL(matrix(beta, nrow=1, ncol=length(beta))),
              flag_est = TRUE, flag_std = TRUE,
              namconv = NamingConvention("dOK"))

# statistics
vn = c(
  paste(c("OK-LC.LC", "dOK.LC", "LC.OK"), "estim", sep = "."),
  paste(c("OK-LC.LC", "dOK.LC"), "stdev", sep = ".")
)
tab = dbStatisticsMono(grd, names = vn, opers = opers, flagIso = FALSE)$toTL()
knitr::kable(tab, caption = "Statistics on the grid for Ordinary Kriging of LC", digits = 4)

# plot direct LC estimate vs. estimate using the LC model
p1 = ggDefault() + 
  plot.correlation(grd, 
                   namex = paste("OK-LC", "LC", "estim", sep = "."), 
                   namey = paste("dOK", "LC", "estim", sep = "."), 
                   asPoint = TRUE, flagDiag = TRUE, diagColor = "red", flagSameAxes = TRUE) +
  plot.decoration(xlab = "Estimate using LC model", ylab = "Direct estimate of the LC", 
                  title = "Ordinary coK: estimated value")
p2 = ggDefault() + 
  plot.correlation(grd, 
                   namex = paste("OK-LC", "LC", "stdev", sep = "."), 
                   namey = paste("dOK", "LC", "stdev", sep = "."), 
                   asPoint = TRUE, flagDiag = TRUE, diagColor = "red", flagSameAxes = TRUE) +
  plot.decoration(xlab = "Estimate using LC model", ylab = "Direct estimate of the LC", 
                  title = "Ordinary coK: Std. of kriging error")
ggarrange(p1, p2, ncol = 2, nrow = 1)

```

# Simple kriging (SK)

## Second Order Stationary model of the raw variable Z 

```{r SK_modeling, echo=TRUE, eval=TRUE}
# copying the IRF0 model into SOS model
mod_Z_SOS = mod_Z_IRF0$clone()
err = mod_Z_SOS$setDriftIRF(-1)
err = mod_Z_SOS$setMeans(m)
if(flag.verbose) {
  err = mod_Z_SOS$display()
}

# defining the model of the LC
mod_LC_SOS = mod_LC_IRF0$clone()
err = mod_LC_SOS$setDriftIRF(-1)
err = mod_LC_SOS$setMeans(as.numeric(matLC %*% m))

if(flag.verbose) {
  err = mod_LC_SOS$display()
}
```

## Simple kriging of the raw variables

```{r SK_estimate, echo=TRUE, eval=TRUE}
err = grd$deleteColumns(paste("SK", nm_var, "*", sep = "."))
err = grd$deleteColumns(paste("LC", "SK", "*", sep = "."))
if(flag.verbose) {
  print(paste0(">>> Simple kriging of the raw variables"))
}
err = data$setLocators(c("Z1", "Z2"), ELoc_Z(), cleanSameLocator=TRUE)
err = kriging(data, grd, mod_Z_SOS, neigh, 
              flag_est = TRUE, flag_std = TRUE,
              namconv = NamingConvention("SK"))

# computing the LC of kriging
grd["LC.SK.estim"]  = matrix(grd$getColumns(names = paste("SK", colnames(matLC), "estim", sep = ".")),
                    nrow = grd$getSampleNumber(),  ncol = dim(matLC)[2]) %*% t(matLC)[,1]

# statistics
vn  = c("Z1", paste("SK", "Z1", c("estim", "stdev"), sep = "."),
        "Z2", paste("SK", "Z2", c("estim", "stdev"), sep = "."))
tab = dbStatisticsMono(grd, vn, opers = opers)$toTL()
knitr::kable(tab, caption = "Statistics on the grid for Simple Kriging with the initial model", 
             digits = 4)

# plot of estimated values
p1 = ggDefaultGeographic() + 
    plot.grid(grd, nameRaster =  paste("SK", "Z1", "estim", sep = "."),
            palette = "Spectral", flagLegendRaster = TRUE, legendNameRaster = "estim") +
    plot.decoration(xlab = "", ylab = "", title = paste0("Simple coK for Z1"))
p2 = ggDefaultGeographic() + 
    plot.grid(grd, nameRaster =  paste("SK", "Z1", "stdev", sep = "."),
            palette = "Spectral", flagLegendRaster = TRUE, legendNameRaster = "stdev") +
    plot.decoration(xlab = "", ylab = "", title = paste0("Std. of Simple coK for Z1"))
p3 = ggDefaultGeographic() + 
    plot.grid(grd, nameRaster =  paste("SK", "Z2", "estim", sep = "."),
            palette = "Spectral", flagLegendRaster = TRUE, legendNameRaster = "estim") +
    plot.decoration(xlab = "", ylab = "", title = paste0("Simple coK for Z2"))
p4 = ggDefaultGeographic() + 
    plot.grid(grd, nameRaster =  paste("SK", "Z2", "stdev", sep = "."),
            palette = "Spectral", flagLegendRaster = TRUE, legendNameRaster = "stdev") +
    plot.decoration(xlab = "", ylab = "", title = paste0("Std. of Simple coK for Z2"))
ggarrange(p1, p2, p3, p4, ncol = 2, nrow = 2)

# plot Z vs. Z*
p1 = ggDefault() + 
  plot.correlation(grd, namex = paste("SK", "Z1", "estim", sep = "."), namey = "Z1", asPoint = TRUE, 
                   flagDiag = TRUE, diagColor = "red", flagSameAxes = TRUE) +
  plot.decoration(xlab = "Estimated values", ylab = "Actual values", title = "Simple coK of Z1")
p2 = ggDefault() + 
  plot.correlation(grd, namex = paste("SK", "Z2", "estim", sep = "."), namey = "Z2", asPoint = TRUE, 
                   flagDiag = TRUE, diagColor = "red", flagSameAxes = TRUE) +
  plot.decoration(xlab = "Estimated values", ylab = "Actual values", title = "Simple coK of Z2")
ggarrange(p1, p2, ncol = 2, nrow = 1)
```

## Simple kriging of a linear combination of the raw variables

```{r SK_LC_estimate, echo=TRUE, eval=TRUE}
err = grd$deleteColumns("SK-LC*")
if(flag.verbose) {
  print(paste0(">>> Simple kriging of a linear combination of the variables"))
}

# modelling the linear combination
# Locate Z
err = data$setLocators(nm_LC, locatorType = ELoc_Z(), cleanSameLocator = TRUE)
# computing the model of the LC from the initial model
var_LC = Vario_computeFromDb(varioParam, data)

# plot
multi.varmod(var_LC, mod_LC_SOS, flagLegend=TRUE)
if(flag.verbose) {
  mod_LC_SOS
}

# kriging the LC variables
err = data$setLocators(nm_LC, ELoc_Z(), cleanSameLocator=TRUE)
err = kriging(data, grd, mod_LC_SOS, neigh, 
              flag_est = TRUE, flag_std = TRUE,
              namconv = NamingConvention("SK-LC"))

# statistics
vn = c(
  paste(c("SK.Z1", "SK-LC.Z1", "SK-LC.LC", "LC.SK"), "estim", sep = "."),
  paste(c("SK.Z1", "SK-LC.Z1", "SK-LC.LC"), "stdev", sep = ".")
)
tab = dbStatisticsMono(grd, names = vn, opers = opers)$toTL()
knitr::kable(tab, caption = "Statistics on the grid for Simple Kriging with the LC model", digits = 4)

# plot Z vs. Z*
p1 = ggDefault() + 
  plot.correlation(grd, 
                   namex = paste("SK", "Z1", "estim", sep = "."), 
                   namey = paste("SK-LC", "Z1", "estim", sep = "."), 
                   asPoint = TRUE, flagDiag = TRUE, diagColor = "red", flagSameAxes = TRUE) +
  plot.decoration(xlab = "Initial estimate", ylab = "Estimate with the LC model", 
                  title = "Simple coK of Z1")
p2 = ggDefault() + 
  plot.correlation(grd, 
                   namey = paste("SK-LC", "LC", "estim", sep = "."), 
                   namex = paste("LC", "SK", "estim", sep = "."), 
                   asPoint = TRUE, flagDiag = TRUE, diagColor = "red", flagSameAxes = TRUE) +
  plot.decoration(xlab = "LC of the estimates", ylab = "Estimate with the LC model", 
                  title = "Simple coK of CL")
ggarrange(p1, p2, ncol = 2, nrow = 1)
```

## Direct computation of linear combination kriging

```{r dSK_LC_estimate, echo=TRUE, eval=TRUE}
err = grd$deleteColumns("dSK*")
if(flag.verbose) {
  print(paste0(">>> Direct computation of linear combination kriging"))
}

# the R matrix matLC has to be converted into a gstlearn MatrixRectangular:
# MatrixRectangular_createFromVD(beta, nrow=1, ncol=length(beta))
# OR
# beta |> matrix(nrow=1, ncol=length(beta)) |> fromTL()
# OR
# fromTL(matrix(beta, nrow=1, ncol=length(beta)))

# kriging LC variables
err = data$setLocators(nm_var, ELoc_Z(), cleanSameLocator=TRUE)
err = kriging(data, grd, mod_Z_SOS, neigh, 
              matLC = fromTL(matrix(beta, nrow=1, ncol=length(beta))),
              flag_est = TRUE, flag_std = TRUE,
              namconv = NamingConvention("dSK"))

# statistics
vn = c(
  paste(c("SK-LC.LC", "dSK.LC", "LC.SK"), "estim", sep = "."),
  paste(c("SK-LC.LC", "dSK.LC"), "stdev", sep = ".")
)
tab = dbStatisticsMono(grd, names = vn, opers = opers, flagIso = FALSE)$toTL()
knitr::kable(tab, caption = "Statistics on the grid for Simple Kriging of LC", digits = 4)

# plot direct LC estimate vs. estimate using the LC model
p1 = ggDefault() + 
  plot.correlation(grd, 
                   namex = paste("SK-LC", "LC", "estim", sep = "."), 
                   namey = paste("dSK", "LC", "estim", sep = "."), 
                   asPoint = TRUE, flagDiag = TRUE, diagColor = "red", flagSameAxes = TRUE) +
  plot.decoration(xlab = "Estimate using LC model", ylab = "Direct estimate of the LC", 
                  title = "Simple coK: estimated value")
p2 = ggDefault() + 
  plot.correlation(grd, 
                   namex = paste("SK-LC", "LC", "stdev", sep = "."), 
                   namey = paste("dSK", "LC", "stdev", sep = "."), 
                   asPoint = TRUE, flagDiag = TRUE, diagColor = "red", flagSameAxes = TRUE) +
  plot.decoration(xlab = "Estimate using LC model", ylab = "Direct estimate of the LC", 
                  title = "Simple coK: Std. of kriging error")
ggarrange(p1, p2, ncol = 2, nrow = 1)
```

# Kriging with heterotopic data

## Creating heterotopic variables

```{r heterotopic_data, echo=TRUE, eval=TRUE}
z1 = data["Z1"]
z1[sample(c(TRUE, FALSE), size = length(z1), prob = c(0.2, 0.8), replace = TRUE)] <- NaN
data["Z1_miss"] <- z1
z2 = data["Z2"]
z2[sample(c(TRUE, FALSE), size = length(z2), prob = c(0.5, 0.5), replace = TRUE)] <- NaN
data["Z2_miss"] <- z2

vn = c("Z1", "Z1_miss", "Z2", "Z2_miss")
tab = dbStatisticsMono(data, names = vn, opers = opers, flagIso = FALSE)$toTL()
rownames(tab) <- vn
knitr::kable(tab, caption = "Statistics on the heterotopic data", digits = 4)
```

## Ordinary Kriging of a linear combination with heterotopic data

```{r heterotopic_dOK, echo=TRUE, eval=TRUE}
err = grd$deleteColumns("dOK_He*")
if(flag.verbose) {
  print(paste0(">>> Direct computation of linear combination kriging with heterotopic data"))
}

# kriging LC variables
err = data$setLocators(paste(nm_var, "miss", sep = "_"), ELoc_Z(), cleanSameLocator=TRUE)
err = kriging(data, grd, mod_Z_IRF0, neigh, 
              matLC = fromTL(matrix(beta, nrow=1, ncol=length(beta))),
              flag_est = TRUE, flag_std = TRUE,
              namconv = NamingConvention("dOK_He"))

# statistics
vn = c(
  paste(c("dOK_He.LC", "dOK.LC", "LC.OK"), "estim", sep = "."),
  paste(c("dOK_He.LC", "dOK.LC"), "stdev", sep = ".")
)
tab = dbStatisticsMono(grd, names = vn, opers = opers, flagIso = FALSE)$toTL()
knitr::kable(tab, caption = "Statistics on the grid for Ordinary Kriging with heterotopic data",
             digits = 4)

# plot direct LC estimate: Heterotopic data vs. Homotopic data
p1 = ggDefault() + 
  plot.correlation(grd, 
                   namex = paste("dOK", "LC", "estim", sep = "."), 
                   namey = paste("dOK_He", "LC", "estim", sep = "."), 
                   asPoint = TRUE, flagDiag = TRUE, diagColor = "red", flagSameAxes = TRUE) +
  plot.decoration(xlab = "Homotopic data", ylab = "Heterotopic data", 
                  title = "Ordinary coK of LC: estimated value")
p2 = ggDefault() + 
  plot.correlation(grd, 
                   namex = paste("dOK", "LC", "stdev", sep = "."), 
                   namey = paste("dOK_He", "LC", "stdev", sep = "."), 
                   asPoint = TRUE, flagDiag = TRUE, diagColor = "red", flagSameAxes = TRUE) +
  plot.decoration(xlab = "Homotopic data", ylab = "Heterotopic data", 
                  title = "Ordinary coK of LC: Std. of kriging error")
ggarrange(p1, p2, ncol = 2, nrow = 1)

# estimate maps: Heterotopic data vs. Homotopic data
p1 = ggDefaultGeographic() + 
  plot.grid(grd, nameRaster =  paste("dOK", "LC", "estim", sep = "."), 
            flagLegendRaster = TRUE, legendNameRaster = "Z*", palette = "Spectral", 
            limits = c(-25,0)) +
  plot.decoration(title = "Estimate: homotopic data")
p2 = ggDefaultGeographic() + 
  plot.grid(grd, nameRaster =  paste("dOK_He", "LC", "estim", sep = "."),
            flagLegendRaster = TRUE, legendNameRaster = "Z*", palette = "Spectral", 
            limits = c(-25, 0)) +
  plot.decoration(title = "Estimate: heterotopic data")
ggarrange(p1, p2, ncol = 2, nrow = 1)

# Std. maps: Heterotopic data vs. Homotopic data
lim = range(c(grd[paste("dOK", "LC", "stdev", sep = ".")], grd[paste("dOK_He", "LC", "stdev", 
                                                                     sep = ".")]))
p1 = ggDefaultGeographic() + 
  plot.grid(grd, nameRaster =  paste("dOK", "LC", "stdev", sep = "."), 
            flagLegendRaster = TRUE, legendNameRaster = "Std.", palette = "Spectral", limits = lim) +
  plot.decoration(title = "IRF-0: homotopic data")
p2 = ggDefaultGeographic() + 
  plot.grid(grd, nameRaster =  paste("dOK_He", "LC", "stdev", sep = "."),
            flagLegendRaster = TRUE, legendNameRaster = "Std.", palette = "Spectral", limits = lim) +
  plot.decoration(title = "IRF-0: heterotopic data")
ggarrange(p1, p2, ncol = 2, nrow = 1)
```

## Simple Kriging of a linear combination with heterotopic data

```{r heterotopic_dSK, echo=TRUE, eval=TRUE}
err = grd$deleteColumns("dSK_He*")
if(flag.verbose) {
  print(paste0(">>> Direct computation of linear combination simple cokriging with heterotopic data"))
}

# kriging LC variables
err = data$setLocators(paste(nm_var, "miss", sep = "_"), ELoc_Z(), cleanSameLocator=TRUE)
err = kriging(data, grd, mod_Z_SOS, neigh, 
              matLC = fromTL(matrix(beta, nrow=1, ncol=length(beta))),
              flag_est = TRUE, flag_std = TRUE,
              namconv = NamingConvention("dSK_He"))

# statistics
vn = c(
  paste(c("dSK_He.LC", "dSK.LC", "LC.SK"), "estim", sep = "."),
  paste(c("dSK_He.LC", "dSK.LC"), "stdev", sep = ".")
)
tab = dbStatisticsMono(grd, names = vn, opers = opers, flagIso = FALSE)$toTL()
knitr::kable(tab, caption = "Statistics on the grid for simple coKriging with heterotopic data",
             digits = 4)

# plot direct LC estimate: Heterotopic data vs. Homotopic data
p1 = ggDefault() + 
  plot.correlation(grd, 
                   namex = paste("dSK", "LC", "estim", sep = "."), 
                   namey = paste("dSK_He", "LC", "estim", sep = "."), 
                   asPoint = TRUE, flagDiag = TRUE, diagColor = "red", flagSameAxes = TRUE) +
  plot.decoration(xlab = "Homotopic data", ylab = "Heterotopic data", 
                  title = "Simple coK of LC: estimated value")
p2 = ggDefault() + 
  plot.correlation(grd, 
                   namex = paste("dSK", "LC", "stdev", sep = "."), 
                   namey = paste("dSK_He", "LC", "stdev", sep = "."), 
                   asPoint = TRUE, flagDiag = TRUE, diagColor = "red", flagSameAxes = TRUE) +
  plot.decoration(xlab = "Homotopic data", ylab = "Heterotopic data", 
                  title = "Simple coK of LC: Std. of kriging error")
ggarrange(p1, p2, ncol = 2, nrow = 1)

# estimate maps: Heterotopic data vs. Homotopic data
p1 = ggDefaultGeographic() + 
  plot.grid(grd, nameRaster =  paste("dSK", "LC", "estim", sep = "."), 
            flagLegendRaster = TRUE, legendNameRaster = "Z*", palette = "Spectral", 
            limits = c(-25,0)) +
  plot.decoration(title = "SOS estimate: homotopic data")
p2 = ggDefaultGeographic() + 
  plot.grid(grd, nameRaster =  paste("dSK_He", "LC", "estim", sep = "."),
            flagLegendRaster = TRUE, legendNameRaster = "Z*", palette = "Spectral", 
            limits = c(-25, 0)) +
  plot.decoration(title = "SOS estimate: heterotopic data")
ggarrange(p1, p2, ncol = 2, nrow = 1)

# Std. maps: Heterotopic data vs. Homotopic data
lim = range(c(grd[paste("dSK", "LC", "stdev", sep = ".")], grd[paste("dSK_He", "LC", "stdev", 
                                                                     sep = ".")]))
p1 = ggDefaultGeographic() + 
  plot.grid(grd, nameRaster =  paste("dSK", "LC", "stdev", sep = "."), 
            flagLegendRaster = TRUE, legendNameRaster = "Std.", palette = "Spectral", limits = lim) +
  plot.decoration(title = "SOS: homotopic data")
p2 = ggDefaultGeographic() + 
  plot.grid(grd, nameRaster =  paste("dSK_He", "LC", "stdev", sep = "."),
            flagLegendRaster = TRUE, legendNameRaster = "Std.", palette = "Spectral", limits = lim) +
  plot.decoration(title = "SOS: heterotopic data")
ggarrange(p1, p2, ncol = 2, nrow = 1)
```

# QC OK vs. SK

```{r qc_ok_vs_sk, echo=TRUE, eval=TRUE}
v = "Z1"

# cross plot: SOS vs. IRF0
p1 = ggDefault() + 
  plot.correlation(grd, 
                   namex = paste("SK", v, "estim", sep = "."), 
                   namey = paste("OK", v, "estim", sep = "."), 
                   asPoint = TRUE, flagDiag = TRUE, diagColor = "red", flagSameAxes = TRUE) +
  plot.decoration(xlab = "SOS model (SK)", ylab = "IRF-0 model (OK)", 
                  title = paste0(v, " Estimation"))

p2 = ggDefault() + 
  plot.correlation(grd, 
                   namex = paste("SK", v, "stdev", sep = "."), 
                   namey = paste("OK", v, "stdev", sep = "."), 
                   asPoint = TRUE, flagDiag = TRUE, diagColor = "red", flagSameAxes = TRUE) +
  plot.decoration(xlab = "SOS model (SK)", ylab = "IRF-0 model (OK)", 
                  title = "Std. of estimation error")
ggarrange(p1, p2, ncol = 2, nrow = 1)

v = "Z2"
# cross plot: SOS vs. IRF0
p1 = ggDefault() + 
  plot.correlation(grd, 
                   namex = paste("SK", v, "estim", sep = "."), 
                   namey = paste("OK", v, "estim", sep = "."), 
                   asPoint = TRUE, flagDiag = TRUE, diagColor = "red", flagSameAxes = TRUE) +
  plot.decoration(xlab = "SOS model (SK)", ylab = "IRF-0 model (OK)", 
                  title = paste0(v, " Estimation"))

p2 = ggDefault() + 
  plot.correlation(grd, 
                   namex = paste("SK", v, "stdev", sep = "."), 
                   namey = paste("OK", v, "stdev", sep = "."), 
                   asPoint = TRUE, flagDiag = TRUE, diagColor = "red", flagSameAxes = TRUE) +
  plot.decoration(xlab = "SOS model (SK)", ylab = "IRF-0 model (OK)", 
                  title = "Std. of estimation error")
ggarrange(p1, p2, ncol = 2, nrow = 1)

# direct kriging of LC
v = "LC"
# cross plot: SOS vs. IRF0
p1 = ggDefault() + 
  plot.correlation(grd, 
                   namex = paste("dSK", v, "estim", sep = "."), 
                   namey = paste("dOK", v, "estim", sep = "."), 
                   asPoint = TRUE, flagDiag = TRUE, diagColor = "red", flagSameAxes = TRUE) +
  plot.decoration(xlab = "SOS model (SK)", ylab = "IRF-0 model (OK)", 
                  title = "LC estimation (direct)")

p2 = ggDefault() + 
  plot.correlation(grd, 
                   namex = paste("dSK", v, "stdev", sep = "."), 
                   namey = paste("dOK", v, "stdev", sep = "."), 
                   asPoint = TRUE, flagDiag = TRUE, diagColor = "red", flagSameAxes = TRUE) +
  plot.decoration(xlab = "SOS model (SK)", ylab = "IRF-0 model (OK)", 
                  title = "Std. of LC estimation error (direct)")
ggarrange(p1, p2, ncol = 2, nrow = 1)
```
