---
title: "Multivariate Covariance and Variogram"
author: "D. Renard"
date: "5 avril 2022"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
rm(list=ls())
library(gstlearn)
library(ggplot2)
library(ggpubr)
library(ggnewscale)
```

## Introduction

This slideshow is meant to demonstrate the experimental multivariate covariance vs. variogram. It is demonstrated in 1-D space.

```{r}
err = defineDefaultSpace(ESpaceType_RN(), 1)
```


Building the data sets

```{r}
grid = DbGrid_create(nx=500)
```

Creating three models

```{r}
m1 = Model_createFromParam(type=ECov_GAUSSIAN(),range=100)
m2 = Model_createFromParam(type=ECov_SPHERICAL(),range=30)
m3 = Model_createFromParam(type=ECov_NUGGET())
```

Simulating independent (normalized) components

```{r}
err = simtub(NULL,grid,m1,nbtuba=1000)
err = grid$setName("Simu","Long")
err = simtub(NULL,grid,m2,nbtuba=1000)
err = grid$setName("Simu","Short")
err = simtub(NULL,grid,m3,nbtuba=1000)
err = grid$setName("Simu","Nugget")
```

Graphic representation

```{r}
p1 = ggDefault() + plot.curve(grid["Nugget"]) + plot.decoration(xlab="Nugget")
p2 = ggDefault() + plot.curve(grid["Short"]) + plot.decoration(xlab="Short Range")
p3 = ggDefault() + plot.curve(grid["Long"]) + plot.decoration(xlab="Long Range")
ggarrange(p1, p2, p3, ncol = 3, nrow = 1)
```

Building the final variables

```{r}
grid["Z1"] =  2*grid["Short"] + 3*grid["Long"] + grid["Nugget"] - 5
grid["Z2"] = -3*grid["Short"] + 4*grid["Long"]
```

Plotting the variables

```{r}
p = ggDefault() + 
  plot.curve(grid["Z1"], col="black") + 
  plot.curve(grid["Z2"], col="blue") + 
  plot.decoration(title="Z1 and Z2")
ggPrint(p)
```

Computing Variograms (Simple and Cross)

```{r}
err = grid$setLocators("Z*",ELoc_Z())
varioparam = VarioParam_createMultipleFromGrid(grid, nlag=100)
vario = Vario_computeFromDb(varioparam, grid)

p = ggDefault() + 
  plot(vario, ivar=0, jvar=0, col="black", cex=0, varioLinetype="solid") + 
  plot(vario, ivar=1, jvar=1, col="blue", cex=0, varioLinetype="solid") +
  plot(vario, ivar=0, jvar=1, col="red", cex=0, varioLinetype="solid") +
  plot.decoration(title="Simple and Cross-Variograms")
ggPrint(p)
```

Computing Covariances (Simple and Cross)

```{r}
err = grid$setLocators("Z*",ELoc_Z())
covar = Vario_computeFromDb(varioparam, grid, calcul=ECalcVario_COVARIANCE())

p = ggDefault() + 
  plot(covar, ivar=0, jvar=0, col="black", cex=0, varioLinetype="solid") + 
  plot(covar, ivar=1, jvar=1, col="blue", cex=0, varioLinetype="solid") +
  plot(covar, ivar=0, jvar=1, col="red", cex=0, varioLinetype="solid") +
  plot.decoration(title="Simple and Cross-Covariances")
ggPrint(p)
```

## Covariance with delay

```{r}
shift = 30
v2 = grid["Z2"]
grid["Z3"] = c(v2[-{1:shift}],rep(NA,shift))
```

Plotting the variables

```{r}
p = ggDefault() + 
  plot.curve(grid["Z2"], col="black") + 
  plot.curve(grid["Z3"], col="blue") + 
  plot.decoration(title="Z2 and Z3")
ggPrint(p)
```

Calculate the cross-variogram

```{r}
grid$setLocators(c("Z2","Z3"),ELoc_Z(),cleanSameLocator=TRUE)
covarh = Vario_computeFromDb(varioparam, grid, calcul=ECalcVario_VARIOGRAM())

p = ggDefault() + 
  plot(covarh, ivar=0, jvar=0, col="black", cex=0, varioLinetype="solid") + 
  plot(covarh, ivar=1, jvar=1, col="blue", cex=0, varioLinetype="solid") +
  plot(covarh, ivar=0, jvar=1, col="red", cex=0, varioLinetype="solid") +
  plot.decoration(title="Simple and Cross-Variograms (shift)")
ggPrint(p)
```

Calculate the cross-covariance

```{r}
covarh = Vario_computeFromDb(varioparam, grid, calcul=ECalcVario_COVARIANCE())

p = ggDefault() + 
  plot(covarh, ivar=0, jvar=0, col="black", cex=0, varioLinetype="solid") + 
  plot(covarh, ivar=1, jvar=1, col="blue", cex=0, varioLinetype="solid") +
  plot(covarh, ivar=0, jvar=1, col="red", cex=0, varioLinetype="solid") +
  plot.decoration(title="Simple and Cross-Covariances (shift)")
ggPrint(p)
```
