---
title: "Graphics in 3D"
author: "gstlearn Team"
date: "2023-09-06"
output:
  html_document:
    df_print: paged
editor_options:
  chunk_output_type: console
---

<!-- SUMMARY: Perform Graphic representations in 3D  -->

<!-- CATEGORY: Graphics -->

# Introduction

In this tutorial, we show how to use the graphics in 3D. It is essentially focused on the representation of a variable on the skin of a sphere.

Initialization of *gstlearn*

```{r intialization, include=FALSE}
library(gstlearn)
library(ggplot2)
library(ggpubr)
library(rgl) # used for 3D vizualisation
rm(list = ls())
```

Initialization of the sphere

```{r create_S2, echo=TRUE, eval = TRUE}
nx = c(360, 180)
dx = nx/(nx-1)*pi/180
x0 = c(0,0)
grd = DbGrid_create(nx = nx, x0 = x0, dx = dx)
err = grd$setName("x1", "phi")
err = grd$setName("x2", "theta")
grd["x"] = sin(grd["theta"]) * cos(grd["phi"])
grd["y"] = sin(grd["theta"]) * sin(grd["phi"])
grd["z"] = cos(grd["theta"])
x = grd["x"]
y = grd["y"]
z = grd["z"]
```

Display of a variable on the skin of the 3D sphere. This variable is known as the Spherical Harmonic (a kind of primary function which stands as the traditional polynomials and which used in many simulation engines on the sphere).

```{r display, echo = TRUE, eval = TRUE}
# Calculate the (real part of) spherical harmonics
val = ut_sphericalHarmonicVec(n = 15, k = 5, theta = grd["theta"], phi = grd["phi"])

# Initialization of S2 for 3d viewer
viz_S2_in_3D = IniView_S2_in_3D(grd)

err = viz_S2_in_3D$display(val = val, ncol = 50, palette = "rainbow")
rglwidget()
```
