---
title: "Spectral simulation of the Gneiting model"
author: "X. Freulon, and the gstlearn Core Team"
date: 2025-11-12
output:
  pdf_document: default
  html_document:
    df_print: paged
editor_options:
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(gstlearn)
rm(list = ls())
h   = seq(from = 0, to = 10, length.out = 100)
```

# The Gneiting covariance

The expression of the space-time Gneiting's covariance on $\mathbb{R}^d \times \mathbb{R}$ is:

$$
C(h, u; \alpha, \beta, \nu) = (1 + \|u\|^{\alpha})^{-\frac{\beta}{2}d} \mathcal{K}_{\nu} (\frac{\|h\|}{(1 + \|u\|^{\alpha})^{\frac{\beta}{2}}})
$$

where $(h,u)$ are the increments, respectively in the space $\mathbb{R}^d$ and time $\mathbb{R}$, $\mathcal{K}$ is the Gauss, Matern, or Cauchy kernel, defining the spatial trace,

* the Gauss kernel is $k(h) = e^{-\|h\|^2}$ (no parameters)
* the Matern kernel is $k(h) = \frac{2^{-\nu}}{\Gamma(\nu)} \|h\|^\nu K_\nu(\|h\|)$ ($\nu > 0$)
* the Cauchy kernel is $k(h) = (1 + \|h\|^2)^{-\nu}$ ($\nu > 0$)

and, the parameters $\alpha \in (0,2]$ and $\beta \in (0,1]$ defines the temporal trance.

In addition a geometrical anisotropy can be defined in scaling coefficients $(\boldsymbol{a}, a_t) > 0$ and and a rotation matrix $R$: $\|h\| = \sqrt{(h^t \left(R^t D_{(1/\boldsymbol{a}^2)}R\right)h)}$ and $\|u\|= |u|/a_t$.

```{r plot_gneiting, echo=FALSE}
# parameters
type = 2 # in 0:2 (GNEITING-GAUSS, GNEITING-MATERN, GNEITING-CAUCHY)
type_label = c("Gneiting-Gauss", "Gneiting-Matern", "Gneiting-Cauchy")
d = 1 # in 1:3
scales = c(1, 2, 2.5, 3)[c(1:d,4)]
alpha  = 1.5
beta   = 1
nu = 3/4 # not used if type == 0

if (type == 0) {
  # Two parameters are required for the Gneiting-Gauss model
  params = c(alpha, beta*d/2) 
  nu = NULL
} else {
  # Three parameters are required for the Gneiting-Matern or Gneiting-Cauchy model
  params = c(alpha, beta*d/2, nu)
}
sep = 1.0
angles = list()
angles[[1+length(angles)]] = 0
angles[[1+length(angles)]] = c(30,0)
angles[[1+length(angles)]] = c(0,0,0)

# using the static *create* function
mg_in_gstlearn  = CorGneiting_create(
  ctxt = CovContext(nvar = 1, ndim = d), 
  type = switch(paste(type), "0" = ECov_GNEITING_G(), "1" = ECov_GNEITING_M(), "2" = ECov_GNEITING_C()),
  alpha = alpha,
  beta = beta, 
  timeRange = scales[d+1],
  params = nu, 
  kappas = 1.0,
  ranges = scales[1:d], # vector of length d
  angles = angles[[d]], # vector of length d
  separability = 1.0,   # scalar in [0,1] (not used)
  flagRange = FALSE
  )
# XF : Old style
  # CorGneiting_create(
  # ctxt = CovContext(nvar = 1, ndim = d+1), 
  # type = switch(paste(type), "0" = ECov_GNEITING_G(), "1" = ECov_GNEITING_M(), "2" = ECov_GNEITING_C()),
  # params = params,
  # ranges = scales,      # vector of length d+1
  # angles = angles[[d]], # vector of length d (no angles in time)
  # separability = 1.0,   # scalar in [0,1]
  # flagRange = FALSE
  # )

```


```{r plot_gstlearn, echo=FALSE}
# R^d \times Time
defineDefaultSpaceRnT(ndim = d)
displayDefaultSpace()

```

# Covariance function

```{r}
tit_main = paste0(type_label[1+type], "(", d, "d)")
if (type == 0) {
tit_sub  = bquote(paste(alpha == .(alpha), "; ", beta == .(beta)))
} else {
tit_sub  = bquote(paste(alpha == .(alpha), "; ", beta == .(beta), "; ", nu == .(nu)))
}
```

## Temporal trace

```{r plot_gneiting_temporal, echo=FALSE}
# Temporal covariance
plot(NULL,  NULL, xlim = range(h), ylim = c(0,1), 
     xlab = "Time distance", ylab = "Covariance", 
     main = paste0(tit_main, ": temporal trace"), 
     sub = tit_sub
     )
abline(h = c(0, 1), col = "gray", lty = 3)
abline(v = c(0), col = "gray", lty = 3)

# model in gstlearn
u0  = c(rep(0.0, d), 1.0); u0 = u0 / sqrt(sum(u0^2))
lines(h, mg_in_gstlearn$sample(h = h, codir = u0, ivar = 0, jvar = 0), col = "black", lty = 1)
legend("topright", 
       legend = c("model"),
       col = c("black"),
       lty = c(1), 
       pch = c(NaN)
       )

```

## Spatial trace
```{r plot_gneiting_spatial, echo=FALSE}
if (d == 1) {
dir_u   = matrix(c(1,0), nrow = 1, ncol = d+1, byrow = TRUE)
}
if (d == 2) {
dir_u   = matrix(c(1,0,0, 1,1,0, 0,1,0, -1,1,0), nrow = 4, ncol = d+1, byrow = TRUE)
}
if (d == 3) {
dir_u   = matrix(c(1,0,0,0, 0,1,0,0, 0,0,1,0), nrow = 3, ncol = d+1, byrow = TRUE)
}
dir_col = palette.colors(nrow(dir_u))

# Spatial covariance
plot(NULL,  NULL, xlim = range(h), ylim = c(0,1), 
     xlab = "Space distance", ylab = "Covariance", 
     main = paste0(tit_main, ": spatial trace"), 
     sub = tit_sub
     )
abline(h = c(0, 1), col = "gray", lty = 3)
abline(v = c(0), col = "gray", lty = 3)
for (idir in 1:nrow(dir_u)) {
  u0 = dir_u[idir,]; u0 = u0 / sqrt(sum(u0^2))
  lines(h, mg_in_gstlearn$sample(h = h, codir = u0, ivar = 0, jvar = 0), 
        col = dir_col[idir], lty = 1)
}
legend("topright", 
       legend = c(paste0("Dir #", 1:nrow(dir_u))),
       col = c(dir_col),
       lty = c(rep(1, nrow(dir_u))), 
       pch = c(rep(NaN, nrow(dir_u)))
       )

```


## Space-Time covariance
```{r plot_gneiting_ST, echo=FALSE}
ndir = 4
dir_col = palette.colors(nrow(dir_u))

# random directions in space-time
dir_u = matrix(rnorm(ndir*(d+1)), nrow = ndir, ncol = d+1)
dir_u = dir_u / apply(X = dir_u, MARGIN = 1, FUN = function(u){sqrt(sum(u^2))})

# space-time covariance
plot(NULL,  NULL, xlim = range(h), ylim = c(0,1), 
     xlab = "Space-time distance", ylab = "Covariance", 
     main = paste0(tit_main, ": space-time covariance"), 
     sub = tit_sub
     )
abline(h = c(0, 1), col = "gray", lty = 3)
abline(v = c(0), col = "gray", lty = 3)
# dir Ox
for (idir in 1:nrow(dir_u)) {
  lines(h, mg_in_gstlearn$sample(h = h, codir = dir_u[idir,], ivar = 0, jvar = 0), 
        col = dir_col[idir], lty = 1)
}
dir_lab = NULL
for (i in 1:ndir)  {
  if (d == 1) {lb = sprintf("[%4.2f, %4.2f]", dir_u[i,1], dir_u[i,2])}
  else if (d == 2) {lb = sprintf("[%4.2f, %4.2f, %4.2f]", dir_u[i,1], dir_u[i,2], dir_u[i,3])}
  else if (d == 3) {lb = sprintf("[%4.2f, %4.2f, %4.2f, %4.2f]", dir_u[i,1], dir_u[i,2], dir_u[i,3], dir_u[i,4])}
dir_lab = c(dir_lab, lb)
}
legend("topright", 
       legend = c(dir_lab),
       col = c(dir_col),
       lty = c(rep(1, nrow(dir_u))), 
       pch = c(rep(NaN, nrow(dir_u)))
       )

```

# Simulations

```{r simulation_parameters, echo=TRUE, eval=TRUE}
d = mg_in_gstlearn$getNDim() - 1
stopifnot(d <= 3)
ns = 5000 #  number of spectral components
prefix = paste("GM", d, "d", sep = "_")

# grid
nx = c(rep(100, 3), 50)[c(1:d, 4)] 
dx = c(rep(0.15, 3), 0.3)[c(1:d, 4)] 
x0 = c(rep(0, 3), 0)[c(1:d, 4)]
grd = DbGrid_create(nx = nx, dx = dx, x0 = x0)

# variogram parameters
nlag = 20
incr = diag(d+1)
if (d == 1) {
incr = cbind(incr, c(1,1))
} else if (d == 2) {
incr = cbind(incr, c(1,1,0))
incr = cbind(incr, c(1,0,1))
incr = cbind(incr, c(0,1,1))
}
ndir = ncol(incr)
col_dir = palette.colors(ndir)
lab_dir = NULL
varioPar = VarioParam()
for (idir in (1:ndir)) {
  dir_param = DirParam_createFromGrid(grd, nlag = nlag, grincr = incr[,idir])
  err = varioPar$addDir(dir_param)
  key = "["
  for (k in 1:(d+1)) {
    key = paste0(key, incr[k,idir])
    if (k <= d) {key = paste0(key, ",")}
    }
  key = paste0(key, "]")
  lab_dir = c(lab_dir, key)
}

```

Performing a single simulation using the *SimuSpectralRN* interface:

```{r one_simulation, echo=TRUE, eval=TRUE}
modelGeneric = ModelGeneric()
err = modelGeneric$setCov(mg_in_gstlearn)

# simulation
gssimu = SimuSpectralRN(1, ns, 100, 123, NULL, TRUE)
err = gssimu$setModelGeneric(modelGeneric)
err = gssimu$simulate()

# computation
err = grd$deleteColumns(names = paste0(prefix, "*"))
err = gssimu$compute(dbout = grd) 

v_name  = paste(prefix, "V1", "simu", sep = ".")
err = grd$setName("Simu", v_name)
err = grd$setLocator(v_name,ELoc_Z())
```


```{r simulate_QC, echo=FALSE, eval=TRUE}
opers   = EStatOption_fromKeys(c("NUM", "MINI", "MAXI", "MEAN", "STDV"))

# -----------------------------------------------------
# statistics
# -----------------------------------------------------
tab = dbStatisticsMono(grd, names = v_name, opers = opers)$toTL()
knitr::kable(tab, caption = paste0("Statistics for simulations of ", tit_main))

# -----------------------------------------------------
# histogram
# -----------------------------------------------------
p.hist = plot.init()
p.hist = p.hist + plot.hist(grd, name = v_name, bins = 20)
p.hist = p.hist + geom_vline(xintercept = mean(mean(grd[v_name])), color= "orange",linetype="solid",alpha=0.5,linewidth=1)
p.hist = p.hist + xlim(c(-3, 3))
p.hist = p.hist + xlab("S1") + ylab("Frequency")
p.hist = p.hist + ggtitle(paste0("Distribution of the values"))
p.hist = p.hist + theme_bw()
p.hist

# -----------------------------------------------------
# maps
# -----------------------------------------------------
lab.coor = c("x", "y", "z", "time")[c(1:d, 4)] 
if (d == 1) {
  idx = matrix(c(1, 2), nrow = 2, ncol = 1)
} else if (d == 2) {
  idx = matrix(c(1, 2, 1, 3, 2, 3), nrow = 2, ncol = 3)
} else if (d == 3) {
  idx = matrix(c(1, 2, 1, 3, 2, 3, 1, 4, 2, 4, 3, 4), nrow = 2, ncol = 6)
}
plot.map <- function(i) {
  plot.init(asp = 1) + 
  plot.raster(grd, name = v_name, posX = idx[1,i]-1, posY = idx[2,i]-1) +
  plot.decoration(title = tit_main, xlab = lab.coor[idx[1,i]], ylab = lab.coor[idx[2,i]])
}

plot.end(plot.map(1))
if (d > 1) {
plot.end(plot.map(2))
plot.end(plot.map(3))
} 
if (d > 2) {
  plot.end(plot.map(4))
  plot.end(plot.map(5))
  plot.end(plot.map(6))
}

# -----------------------------------------------------
# variogram on the simulation
# -----------------------------------------------------
# computing the experimental variogram in various directions
vario = Vario(varioPar)
err   = vario$compute(grd)

# TODO: plot.model with a space-time model does not work here!
# p.var = plot.init()
# p.var = p.var + plot.vario(vario)
# p.var = p.var + plot.model(mg_in_gstlearn)
# p.var

gg  = matrix(NaN, nrow = nlag, ncol = ndir)
hh  = matrix(NaN, nrow = nlag, ncol = ndir)
for (idir in 1:ndir){
  gg[,idir] = vario$getAllGg(idir = idir-1)
  hh[,idir] = vario$getAllHh(idir = idir-1)
}

# plotting variograms
dd     = seq(from = 0, to = max(hh, na.rm = TRUE), length.out = 100)
plot(NULL, NULL, xlim = range(dd), ylim = 1.1*c(0.0, max(gg, na.rm = TRUE)),
     xlab = "Distance", ylab = "Variogram", 
     main = paste0(tit_main, ": one simulation"),
     sub  = tit_sub)
abline(h = 0, col = "black")
abline(h = 1, col = "grey", lty = 3)
abline(v = 0, col = "black")
for (idir in 1:ndir) {
  # empirical variogram
  lines(hh[,idir], gg[,idir], col = col_dir[idir], lty = 3)
  points(hh[,idir], gg[,idir], col = col_dir[idir], pch = 19, cex = 0.5)
  # variogram model in gstlearn
  u0  = incr[, idir]*dx/sqrt(sum((incr[, idir]*dx)^2))
  val = mg_in_gstlearn$sample(h = 0, codir = u0, ivar = 0, jvar = 0) - 
    mg_in_gstlearn$sample(h = dd, codir = u0, ivar = 0, jvar = 0)
  lines(dd, val, col  = col_dir[idir], lw = 2)
}
legend("topright", legend = c(lab_dir, "model"), col = c(col_dir, "gray"), 
       lty = c(rep(3, ndir),1), lw = c(rep(1, ndir),2), pch = c(rep(19, ndir), NA))

```

# Multiple simulations 

Performing multiple simulations using the *simuSpectral* interface:

```{r multiple_simulations, echo = TRUE, eval = TRUE}
# simulations
nsim = 20
err = grd$deleteColumns(names = paste0(prefix, "*"))
modelGeneric = ModelGeneric(CovContext(nvar = 1, ndim = d))
err = modelGeneric$setCov(mg_in_gstlearn)
err = simuSpectral(dbin = NULL, dbout = grd, model = modelGeneric, seed = 0,
                   nbsimu = nsim, ns = ns, namconv = NamingConvention(prefix), verbose = FALSE)
```


```{r multiple_simulations_vario, echo = FALSE, eval = TRUE}
# variograms
gg  = array(NaN, dim = c(nlag, ndir, nsim))
hh  = array(NaN, dim = c(nlag, ndir, nsim))
for (is in 1:nsim) {
  err = grd$setLocator(
    name = paste(prefix, "V1", paste0("S",is), sep = "."), 
    locatorType = ELoc_Z(), 
    cleanSameLocator = TRUE)
  vario = Vario(varioPar)
  err   = vario$compute(grd)
  for (idir in 1:ndir){
    gg[,idir, is] = vario$getAllGg(idir = idir-1)
    hh[,idir, is] = vario$getAllHh(idir = idir-1)
  }
}
# Average variogram
 gg_moy = apply(X = gg, MARGIN = c(1,2), FUN = mean)
 hh_moy = apply(X = hh, MARGIN = c(1,2), FUN = mean)
 
# plotting variograms
dd = seq(from = 0, to = max(hh_moy, na.rm = TRUE), length.out = 100)
plot(NULL, NULL, xlim = range(dd), ylim = 1.1*c(0.0, max(gg_moy, na.rm = TRUE)),
     xlab = "Distance", ylab = "Variogram", 
     main = paste0(tit_main, ": ", nsim, " simulations"),
     sub  = tit_sub)
abline(h = 0, col = "black")
abline(h = 1, col = "grey", lty = 3)
abline(v = 0, col = "black")

for (idir in 1:ndir) {
  # empirical variogram
  lines(hh_moy[,idir], gg_moy[,idir], col = col_dir[idir], lty = 3)
  points(hh_moy[,idir], gg_moy[,idir], col = col_dir[idir], pch = 19, cex = 0.5)
  # variogram model in gstlearn
  u0  = incr[, idir]*dx/sqrt(sum((incr[, idir]*dx)^2))
  val = mg_in_gstlearn$sample(h = 0, codir = u0, ivar = 0, jvar = 0) - 
    mg_in_gstlearn$sample(h = dd, codir = u0, ivar = 0, jvar = 0)
  lines(dd, val, col  = col_dir[idir], lw = 2)

}
legend("topright", legend = c(lab_dir, "model"), col = c(col_dir, "gray"), 
       lty = c(rep(3, ndir),1), lw = c(rep(1, ndir),2), pch = c(rep(19, ndir), NA))

```

# References

* Gneiting, T. (2013). Strictly and non-strictly positive definite functions on spheres.

* Allard, D., Emery, X., Lacaux, C., & Lantuéjoul, C. (2020). Simulating space-time random fields with nonseparable Gneiting-type covariance functions. Statistics and Computing, 30(5), 1479-1495.

* Allard, D., Benoit, L., & Obakrim, S. (2025). Modeling and simulating spatio-temporal, multivariate and nonstationary Gaussian Random Fields: a Gaussian mixtures perspective.
