---
title: "Various Krigings"
author: "gstlearn core team"
date: "`r Sys.Date()`"
output:
  pdf_document: default
  html_document:
    df_print: paged
editor_options:
  chunk_output_type: console
---

# Introduction

This script is meant to compare the different methods to compute the kriging estimate:

* kriging with a unique neighbourhood
* kriging with a moving neighbourhood
* kriging with the Vecchia approximation
* kriging with the SPDE approach

Simple and ordinary kriging are used.

The spatial univariate covariance is a Matern's covariance function with a nugget effect.

The kriging with a unique neighbourhood is considered as the reference.

Notes:
1) on the Vecchia implementation

* it computes only the estimate but not the std. of the kriging error,
* it uses the naming convention differently (the name of the variable is not retrieved from the db),
* it does not use the actual mean for the Simple Kriging (m = 0 is used instead)

2) on the SPDE implementation

* the std. values differ from the reference. The default mesh should adapted?

We will use the Geostatistics library **gstlearn**.

```{r setup, include=FALSE, echo=TRUE}
knitr::opts_chunk$set(echo = TRUE)
rm(list=ls())
library(gstlearn)
OptCst_defineByKey("ASP", 0)
library(ggplot2)
library(ggpubr)

ndim = 2
err = defineDefaultSpace(ESpaceType_RN(), ndim)
```

# Model and data

A Model is created which:

* allows to simulate values at the conditioning data points

* is used for subsequent Kriging operations

```{r def_model, echo=TRUE, eval=TRUE}
type = "OK"
# known mean if SK
m   = 5.0
# Drift order if !SK
order = 0
# variance of the nugget effect
c0  = 0.0
# Matern's covariance
sill = 2.0
range = 0.3
param = 1

model = Model(CovContext(nvar = 1, space = SpaceRN_create(ndim = ndim)))
if (c0 > 0) {
  err = model$addCovFromParam(ECov_NUGGET(), sill = c0)
  }
err = model$addCovFromParam(ECov_MATERN(), range = range, sill = sill, param = param)
if (type == "OK") {
err = model$setDriftIRF(order = order)
} else if (type == "SK") {
err = model$setMean(m)
}
model
```

The set of conditioning data points is generated at random into the 1 by 1 square. Simulation at the conditioning data points is performed using the Turning Bands method.

```{r def_data, echo=TRUE, eval=TRUE}
ndat = 100
data = Db_createFillRandom(ndat = ndat, ndim = ndim, nvar = 0)
err  = simtub(dbin = NULL, dbout = data, model = model,
                   nbsimu = 1, nbtuba = 1000, seed = 1234,
                   namconv = NamingConvention("Z"))
```

A Grid is created which covers the area of the conditioning data set.

```{r def_grid, echo=TRUE, eval=TRUE}
nmesh = 100
grid = DbGrid_create(nx = c(nmesh, nmesh), dx = c(1.0 / nmesh, 1.0 / nmesh))
```

## Standard Kriging using a Unique neighbourhood

The standard Kriging requires the definition of a Neighbourhood. Due to the reasonably small number of data, a Unique neighbourhood has been chosen.

```{r standard_kriging_unique, echo=TRUE, eval=TRUE}
neighU = NeighUnique_create()
err = kriging(dbin = data, dbout = grid, model = model,
              neigh = neighU,
              flag_est = TRUE, flag_std = TRUE,
              namconv = NamingConvention(paste(sep = ".", type, "Unique")))
```

# Standard Kriging using a Moving neighbourhood

```{r kriging_moving, echo=TRUE, eval=TRUE}
neighM = NeighMoving_create(nmaxi = 20, nmini = 20)
err = kriging(dbin = data, dbout = grid, model = model,
              neigh = neighM,
              flag_est = TRUE, flag_std = TRUE,
              namconv = NamingConvention(paste(sep = ".", type, "Moving")))
```

# Kriging using the Vecchia approximation

```{r kriging_vecchia, echo=TRUE, eval=TRUE}
nb_vecchia = 50
err = krigingVecchia(dbin = data, dbout = grid, model = model,
              nb_vecchia = nb_vecchia,
              # flag_est = TRUE, flag_std = TRUE,
              namconv = NamingConvention(paste(sep = ".", type, "Vecchia")))
```

# Kriging using the SPDE approach

```{r kriging_spde, echo=TRUE, eval=TRUE}
err = krigingSPDE(dbin = data, dbout = grid, model = model,
              flag_est = TRUE, flag_std = TRUE,
              namconv = NamingConvention(paste(sep = ".", type, "SPDE")))
```

# Base map for the reference estimate

```{r ref_base_map, echo=FALSE, eval=TRUE}
p  = plot.init(asp = 1.0) +
  plot.raster(grid, name=paste(sep = ".", type, "Unique", "Z",  "estim"),
              palette = "Spectral") +
  plot.symbol(data, nameSize="Z", color="black") +
  plot.decoration(title = "Estimated value")

q  = plot.init(asp = 1.0) +
  plot.raster(grid, name=paste(sep = ".", type, "Unique", "Z",  "stdev"),
              palette = "Spectral") +
  plot.decoration(title = "Std. of kriging error")
ggarrange(p, q, nrow = 1, ncol = 2)
```

# Correlation plot for the estimates

```{r correlation_estimate}
for (met in c("Moving", "Vecchia", "SPDE")) {
  p  = plot.init(asp = 1.0) +
    plot.correlation(grid,
           namex = paste(sep = ".", type, met, "Z",  "estim"),
           namey = paste(sep = ".", type, "Unique", "Z",  "estim"),
           flagDiag = TRUE, diagColor = "red", asPoint = TRUE) +
    plot.decoration(title = paste0(type, " estimate"), xlab = met, ylab = "Unique")
  plot.end(p)
}
```

# Correlation plot for the Std. of kriging error

```{r correlation_stdev}
for (met in c("Moving", "SPDE")) {
  p  = plot.init(asp = 1.0) +
    plot.correlation(grid,
           namex = paste(sep = ".", type, met, "Z",  "stdev"),
           namey = paste(sep = ".", type, "Unique", "Z",  "stdev"),
           flagDiag = TRUE, diagColor = "red", asPoint = TRUE) +
    plot.decoration(title = paste0("Std. of ", type, " error"), xlab = met, ylab = "Unique")
  plot.end(p)
}
```
