---
title: "SPDE Spiral"
author: "Didier Renard, Nicolas Desassis, Xavier Freulon"
date: "2023-07-10"
output: pdf_document
editor_options: 
  chunk_output_type: console
---
<!-- SUMMARY: Perform Estimation with non-stationary covariance (anisotropy following a spiral) in the framework of SPDE  -->

<!-- CATEGORY: SPDE -->

In this tutorial, we show how the use of SPDE for Varying Anisotropy when this Anisotropy must follow a Spiral shape (defined as an external function)

```{r intialization, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
rm(list=ls())
library(gstlearn)
library(ggplot2)
library(ggpubr)
library(ggnewscale)
```

Defining the Model as a single Bessel structure. This function is defined as anisotropic: we clearly specify the extension of the ranges in the two main directions. The angle does not have to be defined here: it will be overwritten later as the non-stationary parameter. Note that it is essential to define the short range of the anisotropy ellipsoid first (for the definition of angle as defined in the Spiral function used as a function)... otherwise future results will represent the shape otabined as the orthogonal of the spiral.

```{r Model}
model = Model_createFromParam(ECov_MATERN(), 1., 1., 1., c(4.,45.))
```

A Spiral function is defined and attached to the Model: this is a manner to update the Model by transforming the anisotropy angle as the unique non-stationary parameter.

```{r Spiral}
spirale = FunctionalSpirale(0., -1.4, 1., 1., 50., 50.)
nostat = NoStatFunctionalCov(spirale)
err = model$getCova(0)$addNoStat(nostat)
model$display()
```

Creating a output grid

```{r Grid}
grid = DbGrid_create(c(101,101), c(1.,1.)) 
```

Perform several non-conditional simulations on the grid, using the Model and the non-stationarity.

```{r Non-conditional_simulations}
nbsimu = 4
iuid = simulateSPDE(NULL,grid,model,nbsimu)
grid
```

We represent the non-conditional simulations

```{r Graphic_Non-conditional_simulations}
p1 = ggDefaultGeographic() + plot.grid(grid, nameRaster="SimuSPDE.1")
p2 = ggDefaultGeographic() + plot.grid(grid, nameRaster="SimuSPDE.2")
p3 = ggDefaultGeographic() + plot.grid(grid, nameRaster="SimuSPDE.3")
p4 = ggDefaultGeographic() + plot.grid(grid, nameRaster="SimuSPDE.4")
ggarrange(p1, p2, p3, p4, ncol = 2, nrow = 2)
```

Extracting a set of nodes randomly located in order to create a data file which will serve as conditioning. The data is extracted from the first non-conditional simulation.

```{r Data_extraction}
data = Db_createSamplingDb(grid, number=100, names=c("x1", "x2", "SimuSPDE.1"))
err = data$setName("*SimuSPDE.1", "data")
data
```

```{r Display_data}
ggDefaultGeographic() + plot.point(data, nameColor="data")
```

Use the previous data set (and the non-stationary Model) in order to perform an estimation

```{r Kriging}
iuid = krigingSPDE(data,grid,model)
grid
```

Representing the Estimation obtained on the Grid

```{r Display_Kriging}
ggDefaultGeographic() + plot.grid(grid, nameRaster="KrigingSPDE.data.estim")
```

Performing several conditional simulation

```{r Conditional_simulations}
nbsimu = 4
iuid = simulateSPDE(data,grid,model,nbsimu)
grid
```

Representing the conditional simulations

```{r Display_Conditional_Representations}
p1 = ggDefaultGeographic() + plot.grid(grid, nameRaster="SimuSPDE.data.1")
p2 = ggDefaultGeographic() + plot.grid(grid, nameRaster="SimuSPDE.data.2")
p3 = ggDefaultGeographic() + plot.grid(grid, nameRaster="SimuSPDE.data.3")
p4 = ggDefaultGeographic() + plot.grid(grid, nameRaster="SimuSPDE.data.4")
ggarrange(p1, p2, p3, p4, ncol = 2, nrow = 2)
```

