---
title: "Test of the correlations functions using the spectral approach"
author: "The gstlearn Team"
date: "`r Sys.Date()`"
output: pdf_document
editor_options:
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Introduction

This script evaluates the following class of *gstlearn*:

* The class **CorAniso** implements a univariate stationary covariance on $\mathbb{R}^d$,

* The class **CorMatern** implements the multivariate stationary Matern's stationary covariance on $\mathbb{R}^d$,

* The class **CorGaussianMixture** implements the multivariate stationary Matern's or Cauchy covariance on $\mathbb{R}^d$,

* The class **CorGneiting** implements the multivariate stationary covariance on $\mathbb{R}^d \times \mathbb{R}$.


For all these covariances, the variance is equal to one and a geometrical anisotropy is defined on $\mathbb{R}^d$ by a rotation and $d$ scaling or range positive coefficients.

Four covariances functions are defined:

1) the **EXPONENTIAL** covariance: $C(\boldsymbol{h}) = \exp(-\|\boldsymbol{h}\|)$,

2) the **GAUSSIAN** covariance: $C(\boldsymbol{h}) = \exp(-\|\boldsymbol{h}\|^2)$,

3) the **MATERN** covariance: $C(\boldsymbol{h}; \nu) = \frac{2^{1-\nu}}{\Gamma(\nu)} \|\boldsymbol{h}\|^\nu K_\nu(\|\boldsymbol{h}\|) = \mathcal{M}_{\nu}(\|\boldsymbol{h}\|)$,

4) the **CAUCHY** covariance: $C(\boldsymbol{h}; \nu) = (1 + \|\boldsymbol{h}\|^2)^{-\nu} = \mathcal{C}_{\nu}(\|\boldsymbol{h}\|)$,

5) the **CAUCHY_GEN** covariance: $C(\boldsymbol{h}; \alpha, \nu) = (1 + \|\boldsymbol{h}\|^\alpha)^{-\nu}$,

6) the **GNEITING_G** covariance: $C((\boldsymbol{h}, u); \alpha, \beta, \kappa) = (1 + |\frac{u}{a_t}|^\alpha)^{-\beta\frac{d}{2}} \exp(-\frac{\|\kappa\boldsymbol{h}\|^2}{(1 + |\frac{u}{a_t}|^\alpha)^{\beta}})$,

7) the **GNEITING_M** covariance: $C_{ij}((\boldsymbol{h}, u); \alpha, \beta, \boldsymbol{\nu}, \boldsymbol{\kappa}) = \tau_{ij} \,  (1 + |\frac{u}{a_t}|^\alpha)^{-\beta\frac{d}{2}} \mathcal{M}_{\nu_{ij}}(\kappa_{ij}\, \frac{\|\boldsymbol{h}\|}{(1 + |\frac{u}{a_t}|^\alpha)^{\beta/2}})$, with

* $\nu_{ij} = (\nu_i + \nu_j)/2$,

* $\kappa_{ij} = \sqrt{(\kappa_i^2 + \kappa_j^2)/2}$,

* $\tau_{ij} = \frac{\Gamma(\nu_{ij})}{\sqrt{\Gamma(\nu_{i})\Gamma(\nu_{j})}}\, \frac{\kappa_{i}^{\nu_{i}}\kappa_{j}^{\nu_{j}}}{\kappa_{ij}^{2\nu_{ij}}}$

8) the **GNEITING_C** covariance: $C((\boldsymbol{h}, u); \alpha, \beta, \boldsymbol{\nu}, \boldsymbol{\kappa}) = \tau_{ij} \, (1 + |\frac{u}{a_t}|^\alpha)^{-\beta\frac{d}{2}} \mathcal{C}_{\nu_{ij}}(\kappa_{ij}\, \frac{\|\boldsymbol{h}\|}{(1 + |\frac{u}{a_t}|^\alpha)^{\beta/2}})$,

* $\nu_{ij} = (\nu_i + \nu_j)/2$,

* $\kappa_{ij} = 1/\sqrt{(1/\kappa_i^2 + 1/\kappa_j^2)/2}$,

* $\tau_{ij} = \frac{\Gamma(\nu_{ij})}{\sqrt{\Gamma(\nu_{i})\Gamma(\nu_{j})}}\, \frac{\kappa_{ij}^{2\nu_{ij}}}{\kappa_{i}^{\nu_{i}}\kappa_{j}^{\nu_{j}}}$

In all cases,

* $\boldsymbol{h} \in \mathbb{R}^d$ is the increment,
* $\|\boldsymbol{h}\|$ is the norm of the increments defined by the geometrical anisotropy,
* $\boldsymbol{\nu} > 0$ are the shape parameters defined for each variable, and
* $\boldsymbol{\kappa} > 0$ are the scale/range parameters defined for each variable.

For the Gneiting's models, the covariance trace in time is defined by

* $\alpha \in (0,2]$,
* $\beta \in (0,1]$, and
* $a_t > 0$ a scale/range parameter of the time increment.

```{r, echo=FALSE}
knitr::opts_chunk$set(echo = TRUE)
rm(list = ls())
library(gstlearn)
library(ggplot2)
library(ggpubr)

# global parameters
flag.verbose  = TRUE
flag.msg      = FALSE
flag.gstlearn = FALSE
opers = EStatOption_fromKeys(c("NUM", "MINI", "MAXI", "MEAN", "STDV"))
```

# R functions

### Auxiliary functions

```{r my_functions_aux, echo=FALSE, eval=TRUE}
# ------------------------------------------------------------------------------
# Simulation of stable distribution implemented in LawStable class
# ------------------------------------------------------------------------------

# vector function for cardinal sine function
sinc <- function(u) {
  val = rep(1.0, length(u))
  sel = (abs(u) > 0)
  val[sel] <- sin(u[sel])/u[sel]
  val
}

# Simulation of unilateral stable distribution S_{\alpha} with alpha \in (0,1)
A <- function(u, alpha) {
  ((sin(alpha*u)^alpha) * (sin((1-alpha)*u)^(1-alpha)) / sin(u))^(1/(1-alpha))
}

BB0 <- function(x, alpha) {
  stopifnot((alpha >= 0)&(alpha <= 1))
  sinc(x)/(sinc(alpha*x)^alpha)/(sinc((1-alpha)*x)^(1-alpha))
}

#' Simulation of unilateral stable distribution  $S_{\alpha}$
#' with $alpha \in (0,1)$

simulate_stable_unilateral <- function(n = 1, alpha) {
  b = (1-alpha)/alpha
  (A(pi*runif(n), alpha = alpha)/(-log(runif(n))))^b
}

# bilateral stable distribution:alpha in (1,2]

#' simulation of a Cauchy distribution
#' @param n number of values to be simulated
simulate_cauchy <- function(n) {rnorm(n)/rnorm(n)}

#' simulation of a bilateral stable distribution
#' @param n number of values to be simulated
#' @param alpha the tail index (1 < alpha =< 2)
simulate_stable_bilateral <- function(n, alpha) {
  stopifnot((alpha > 0)&(alpha <= 2))
  U = pi*(runif(n)-0.5) # uniform on [-pi/2, pi/2]
  if (alpha == 1.0) { # Cauchy distribution
    tan(U)
  } else {
    E = -log(runif(n))        # exponentially distributed
    sin(alpha*U)/(cos(U)^(1/alpha))*(cos((1-alpha)*U)/E)^((1-alpha)/alpha)
  }
}

#' Sinulation of exponentially tilted unilateral stable distribution $S_{\alpha, \lambda}$
#' with $alpha \in (0,1)$ and $\lambda > 0$

simulate_stable_unilateral_exptilted <- function(n = 1, alpha, lambda, flag.double = TRUE) {
  stopifnot((alpha > 0)&(alpha <= 1))
  stopifnot(lambda >= 0)
  X = rep(NaN, n)
  nt = 0
  for (i in 1:n) {
    if (flag.double) {
      res  = simulate_stable_unilateral_exptilted_scalar_opt(alpha = alpha, lambda = lambda)
    } else {
      res  = simulate_stable_unilateral_exptilted_scalar(alpha = alpha, lambda = lambda)
    }
    X[i] = res$X
    nt = nt + res$n
  }
  list(X = X, nt = nt/n)
}

# ------------------------------------------------------------------
# naive rejection
# ------------------------------------------------------------------
# simulation of a single value
simulate_stable_unilateral_exptilted_scalar <- function(alpha, lambda) {
  n = 0
  while (TRUE) {
    X = simulate_stable_unilateral(n = 1, alpha = alpha)
    n = n + 1
    if(runif(1) < exp(-lambda*X)) {break}
  }
  list(X = X, n = n)
}
# ------------------------------------------------------------------
# double rejection method
# ------------------------------------------------------------------
# simulation of a single value
simulate_stable_unilateral_exptilted_scalar_opt <- function(alpha, lambda) {
  # No tilting
  if (lambda == 0.0) {
    X = simulate_stable_unilateral(alpha = alpha)
    n = 1
  } else {
    # lambda > 0
    # ------------------------------------------------------------
    # set-up
    # ------------------------------------------------------------
    n   = 0
    b   = (1 - alpha)/alpha
    gam = lambda^alpha * alpha * (1 - alpha)
    xi  = ((2 + sqrt(pi/2))*sqrt(2*gam) + 1) / pi
    psi = exp(-1/8 * gam *pi^2) * (2 + sqrt(pi/2)) * sqrt(gam*pi) / pi
    w1  = xi * sqrt(pi/(2*gam))
    w2  = 2  * psi * sqrt(pi)
    w3  = xi * pi
    if (gam >= 1.0) {w3 = 0.0}
    else {w1 = 0.0}
    w_tot = w1 + w2 + w3

    # generate U with density proportional to g*/G*
    while(TRUE) {
      # generate a proposal according to g**
      n = n + 1
      W = runif(1)
      if (W < w1/w_tot) {
        U <- abs(rnorm(1))/sqrt(gam)
      } else if (W < (w1 + w2)/w_tot) {
        U <- pi *(1 - runif(1)^2)
      } else {
        U <- pi * runif(1)
      }
      if (U < pi) {
        zeta = sqrt(BB0(U, alpha = alpha))
        phi  = (sqrt(gam)+alpha*zeta)^(1/alpha)
        z    = phi/(phi - sqrt(gam)^(1/alpha))
        g1 = 1/pi*exp(lambda^alpha*(1-1/zeta^2))*((1+sqrt(pi/2))*sqrt(gam)/zeta +z)
        g2 = psi/sqrt(pi-U)
        if (gam >= 1.0) {g2 = g2 + xi*exp(-1/2*gam*U^2)}
        else {g2 = g2 + xi}
        rho = g2/g1
        Z = runif(1)*g2/g1
        if(Z <= 1.0) {break}
      }
    } # U has density g*/G* and Z is uniform on [0,1]

    # generate X with density proportional to g(x, U)
    a = A(u = U, alpha = alpha)
    m = (b*lambda/a)^alpha
    delta = sqrt(m*alpha/a)
    a1 = delta*sqrt(pi/2)
    a2 = delta
    a3 = z/a
    s  = a1 + a2 + a3
    while(TRUE) {
      n = n + 1
      V = runif(1)
      if(V < a1/s) { # X < m
        N_prime = rnorm(1)
        X = m - delta * abs(N_prime)
        val_test = - 1/2 * N_prime^2
      } else if (V < (a1+a2)/s) { # m < X < m + delta
        X = m + delta*runif(1)
        val_test = 0.0
      } else { # m + delta < X
        E_prime = -log(runif(1))
        X = m + delta + a3*E_prime
        val_test = - E_prime
      }
      val_test = val_test + a*(X-m)+lambda*(1/X^b - 1/m^b)
      if((X >=0)&(val_test <= -log(Z))) {break}
    }
    X = 1/X^b
  } # lambda > 0
  list(X = X, n = n)
}

# ------------------------------------------------------------------------------
# function for QC
# ------------------------------------------------------------------------------
reduce_model <- function(model, ivar = 1, jvar = 1, verbose = TRUE) {
  val  = NULL
  nvar = model$getNVar()
  stopifnot((ivar %in% 1:nvar)&(jvar %in% 1:nvar))
  type = model$type()
  if (nvar == 1) {
    val = model
  }
  if (nvar > 1) {
    nu    = model$getNus()[ivar, jvar]
    kappa = model$getKappas()[ivar, jvar]
    sigma = model$variance(ivar, jvar)
    ani   = model$getAnisotropy()
    if (type %in% MyType_[c(3,4)]) {
      val = MyCorGaussianMixture(type = type, nu = nu, kappa = kappa, sigma = sigma,
                        anisotropy = ani, verbose = verbose)
    }
    if (type %in% MyType_[c(6,7,8)]) {
      alpha = model$alpha()
      beta  = model$beta()
      timeRange   = model$timeRange()
      val = MyCorGneiting(type = type, nu = nu, kappa = kappa, sigma = sigma,
                        alpha = alpha, beta = beta, timeRange = timeRange,
                        anisotropy = ani, verbose = verbose)
    }
  }
  val
}

# ------------------------------------------------------------------------------
# Auxiliary functions for simulation and QC
# ------------------------------------------------------------------------------
expand_dbgrid <- function(grid) {
  ndim = grid$getNDim()
  x0 = grid$getX0s()
  dx = grid$getDXs()
  nx = grid$getNXs()
  xx = list()
  for (idim in 1:ndim) {
    xx[[1+length(xx)]] <- (x0[idim] + (0:(nx[idim]-1))*dx[idim])
  }
  X = as.matrix(expand.grid(xx))
  if (ndim == 1) {X = t(X)}
  X
}

# ------------------------------------------------------------------------------
# rotations 1D/2D/3D
# ------------------------------------------------------------------------------
rot1D <- function() {
  matrix(1,1,1)
}
rot2D <- function(angles = 0.0) {
  theta = angles[1] * pi / 180
  cs = cos(theta)
  sn = sin(theta)
  matrix(c(cs, sn, -sn, cs), 2, 2)
}
rot3D <- function(angles = rep(0.0, 3), flag.euler = TRUE) {
  # yaw
  alpha = angles[1] * pi / 180
  cs = cos(alpha); sn = sin(alpha)
  Rz = matrix(c(cs, sn, 0, -sn, cs, 0, 0, 0, 1), 3, 3)
  # pitch
  beta = angles[2] * pi / 180
  cs = cos(beta); sn = sin(beta)
  Ry = matrix(c(cs, 0, -sn, 0, 1, 0, sn, 0, cs), 3, 3)
  # roll
  gamma = angles[3] * pi / 180
  cs = cos(gamma); sn = sin(gamma)
  Rx = matrix(c(1, 0, 0, 0, cs, sn, 0, -sn, cs), 3, 3)
  if(flag.euler) {
    # extrinsic rotation whose (improper) Euler angles are (alpha, beta, gamma)
    Rx %*% Ry %*% Rz
  } else {
    # intrinsic rotation whose Tait-Bryan angles are (alpha, beta, gamma)
    Rz %*% Ry %*% Rx
  }
}

# rotations 1D/2D/3D
MyRotation = function(angles, flag.euler = FALSE) {
  stopifnot(length(angles) %in% 1:3)
  if (length(angles) == 1) {
    direct = rot1D()
  }
  if (length(angles) == 2) {
    direct = rot2D(angles = angles)
  }
  if (length(angles) == 3) {
    direct = rot3D(angles = angles, flag.euler = flag.euler)
    }
  fn_applyDirect = function(U) {
    stopifnot(ncol(U) == dim(direct)[1])
    U %*% t(direct)
  }
  fn_applyInverse = function(U) {
    stopifnot(ncol(U) == dim(direct)[1])
    U %*% direct
  }
  list(
    getNDim = function(){dim(direct)[1]},
    getAngles = function(){angles},
    getMatrixDirect = function(){direct},
    getMatrixInverse = function(){t(direct)},
    applyDirect  = function(u){direct %*% u},
    applyInverse = function(u){t(direct) %*% u},
    applyDirectVec = fn_applyDirect,
    applyInverseVec = fn_applyInverse
  )
}

# ------------------------------------------------------------------------------
# geometrical anisotropy defined by scale parameters and a rotation
# ------------------------------------------------------------------------------
MyAniso = function(scales = c(1, 1), rotation = MyRotation(angles = rep(0.0, length(scales)))) {
  stopifnot(rotation$getNDim() == length(scales))
  stopifnot(all(scales > 0))
  d = length(scales)
  if (d == 1) {
    diag_inv = matrix(1/scales, nrow = 1, ncol = 1)
    diag_dir = matrix(scales, nrow = 1, ncol = 1)
  } else {
    diag_inv = diag(1/scales)
    diag_dir = diag(scales)
  }
  L_h2u = diag_inv %*% rotation$getMatrixInverse()
  Q = t(L_h2u) %*% L_h2u
  L_u2h = rotation$getMatrixDirect() %*% diag_dir

  fn_description <- function() {
    desc   = NULL
    d      = rotation$getNDim()
    angles = rotation$getAngles()
    if (d == 2) {
      if (any(scales != 1)) {
        desc = bquote(paste(.(desc), .(sprintf("a = [%.1f, %.1f]", scales[1], scales[2]))))
      }
      angles = rotation$getAngles()
      if (any(angles != 0)) {
        desc = bquote(paste(.(desc), "; ", theta == .(sprintf("%.0f°", angles[1]))))
      }
    }
    if (d == 3) {
      if (any(scales != 1)) {
        desc = bquote(paste(.(desc), .(sprintf("a = [%.1f, %.1f, %.1f]", scales[1], scales[2], scales[3]))))
      }
      angles = anisotropy$getRotation()$getAngles()
      if (any(angles != 0)) {
        desc = bquote(paste(.(desc), "; ", theta == .(sprintf("[%.0f, %.0f, %.0f]", angles[1], angles[2], angles[3]))))
      }
    }
    desc
  }
  list(
    description = fn_description,
    getNDim = function(){rotation$getNDim()},
    getScales = function(){scales},
    getAngles = function(){rotation$getAngles()},
    getRotation = function(){rotation},
    getMatrixDirect = function(){L_h2u},
    getMatrixInverse = function(){L_u2h},
    getMatrix = function(){Q},
    jacobian = function(){prod(scales)},
    applyDirect = function(H){H %*% t(L_h2u)},
    applyInverse = function(U){U %*% t(L_u2h)},
    normH = function(H) {apply(X = H %*% t(L_h2u),
                              MARGIN = 1, FUN=function(u){sqrt(sum(u^2))})},
    normS = function(U) {apply(X = U %*% L_u2h,
                              MARGIN = 1, FUN=function(u){sqrt(sum(u^2))})}
  )
}

#---------------------------
# Exponential kernel
#---------------------------
# h and u as scalars...
cov_E <- function(h) {
    exp(- abs(h))
}
#---------------------------
# Gaussian kernel
#---------------------------
# h and u are scalar
cov_G <- function(h) {
    exp(-h^2)
}

spectral_density_G <- function(u, d) {
  exp(-1/4 * u^2)/(4*pi)^(d/2)
}

simulate_spectral_G <- function(n, d) {
  stopifnot(n >  0)
  stopifnot(d > 0)
  matrix(data = rnorm(n = n*d), nrow = n, ncol = d)*sqrt(2)
}

#---------------------------
# Matern kernel
#---------------------------
cov_M <- function(h, nu) {
  stopifnot(nu > 0)
  h = abs(h)
  sel = (h > 0)
  cov = rep(1.0, length(h))
  cov[sel] <- h[sel]^nu * besselK(x = h[sel], nu = nu) / 2^(nu-1) / gamma(nu)
  cov
}

spectral_density_M <- function(u, d, nu) {
    stopifnot(nu > 0)
    gamma(nu + d/2)/gamma(nu)/(pi)^(d/2)/(1 + u^2)^(nu + d/2)
}

simulate_spectral_M <- function(n, d, nu) {
    stopifnot(n >  0)
    stopifnot(d > 0)
    stopifnot(nu > 0)
    matrix(data = rnorm(n = n*d), nrow = n, ncol = d)/sqrt(2*rgamma(n = n, shape = nu))
}

#---------------------------
# Cauchy kernel
#---------------------------
cov_C <- function(h, nu) {
  stopifnot(nu > 0)
  1/(1+abs(h)^2)^nu
}

spectral_density_C <- function(u, d, nu) {
  nd = 1000 # a parameter to be set
  r  = qgamma(p = (1:nd)/(1+nd), shape = nu)
  X  = outer(X = u, Y = r, FUN = function(u, r){exp(-u^2/4/r)/sqrt(2*r)^2})
  apply(X = X, MARGIN = 1, FUN = mean)/(2*pi)^(d/2)
}

simulate_spectral_C <- function(n, d, nu) {
    stopifnot(n >  0)
    stopifnot(d > 0)
    stopifnot(nu > 0)
    matrix(data = rnorm(n = n*d), nrow = n, ncol = d)*sqrt(2*rgamma(n = n, shape = nu))
}

#---------------------------
# Generalized Cauchy kernel
#---------------------------
cov_GC <- function(h, nu, alpha) {
  stopifnot(nu > 0)
  stopifnot((alpha > 0)&(alpha <= 2))
  1/(1+abs(h)^alpha)^nu
}

spectral_density_GC <- function(u, d, nu, alpha) {
  print(paste0("Generalized CauchySpectral density not implented"))
  NULL
}

simulate_spectral_GC <- function(n, d, nu, alpha) {
  print(paste0("Generalized CauchySpectral density not implented"))
  NULL
}


# functions for the Gaussian mixture
density_IG <- function(x, alpha, beta) {
  (beta/x)^(1+alpha) * exp(- beta/x)/beta/gamma(alpha)
}
density_G <- function(x, nu, lambda) {
  dgamma(x, shape = nu, rate = lambda)
#  exp(- lambda*x) * (lambda*x)^(nu-1) * lambda / gamma(nu)
}
density_scale_Matern <- function(x, nu, kappa) {
  density_IG(x, alpha = nu, beta = (kappa/2)^2)
}
density_scale_Cauchy <- function(x, nu, kappa) {
  dgamma(x, shape = nu, rate = kappa^2, log = TRUE)
  # density_G(x, nu = nu, lambda = kappa^2)
}

kernel_power_exponential <- function(t, alpha = 2.0) {
  stopifnot((alpha > 0)&(alpha <= 2.0))
  exp(-abs(t)^alpha)
}

kernel_matern <- function(h, nu){
  val = rep(1.0, length(h))
  sel = (abs(h) > 0.0)
  val[sel] = abs(h[sel])^nu*besselK(abs(h[sel]), nu = nu)/gamma(nu)/2^(nu-1)
  val
}

kernel_gauss  <- function(h){exp(-h^2)}

kernel_generalized_cauchy <- function(u, alpha, beta) {(1+abs(u)^alpha)^{-beta}}

kernel_cauchy <- function(u, nu) {kernel_generalized_cauchy(u, alpha = 2, beta = nu)}

# -------------------------------------------------------------
MyKernel <- function(type = MyType_[1], d = 2, params = NULL) {
  stopifnot(type %in% MyType_[1:5])
  stopifnot(d %in% 1:3)
  fn_covariance = NULL
  fn_spectrum   = NULL
  fn_simulate   = NULL
  # exponential or Matern kernel
  if (type %in% MyType_[c(1,3)]) { # Exponential or Matern
    if (type %in% MyType_[c(1)]) {params = c(1/2)}
    fn_covariance = function(h) {cov_M(h = h, nu = params[1])}
    fn_spectrum = function(u) {spectral_density_M(u = u, d = d, nu = params[1])}
    fn_simulate = function(n) {simulate_spectral_M(n = n, d = d , nu = params[1])}
  }
  # Gaussian kernel
  if (type %in% MyType_[c(2)]) {
    fn_covariance = function(h) {cov_G(h = h)}
    fn_spectrum = function(u) {spectral_density_G(u = u, d = d)}
    fn_simulate = function(n) {simulate_spectral_G(n = n, d = d)}
  }
  # Cauchy kernel
  if (type %in% MyType_[c(4)]) {
    fn_covariance = function(h) {cov_C(h = h, nu = params[1])}
    fn_spectrum = function(u) {spectral_density_C(u = u, d = d, nu = params[1])}
    fn_simulate = function(n) {simulate_spectral_C(n = n, d = d, nu = params[1])}
  }
  # Generalized Cauchy kernel
  if (type %in% MyType_[c(5)]) {
    # NOTE: the order of the parameters is inverted...
    fn_covariance = function(h) {cov_GC(h = h, nu = params[2], alpha = params[1])}
    fn_spectrum = function(u) {spectral_density_GC(u = u, d = d, nu = params[2], alpha = params[1])}
    fn_simulate = function(n) {simulate_spectral_GC(n = n, d = d, nu = params[2], alpha = params[1])}
  }

  fn_description <- function() {
    desc = bquote(paste(.(type)))
    if (type %in% MyType_[c(3, 4, 5)])  { # "Matern", "Cauchy", "Generalized Cauchy"
    desc = bquote(paste(.(desc), "; ", nu == .(round(params[1],2))))
    }
    if (type %in% MyType_[c(5)]) { # "Generalized Cauchy"
    desc = bquote(paste(.(desc), "; ", alpha == .(round(params[2],2))))
    }
    desc
  }
  list(
    type = function(){type},
    description = fn_description,
    getNDim = function(){d},
    getNVar = function(){1},
    getParameters = function(){params},
    variance = function() {1.0},
    covariance = fn_covariance,
    spectrum = fn_spectrum,
    simulate = fn_simulate
  )
}


# ------------------------------------------------------------------------------
# computing a spectral model on R^d
# ------------------------------------------------------------------------------
# It works in mono and multi variate cases
#' @param sp a list containing spectral waves
#'  - sp$omega a [np, d] matrix of space-time random frequencies
#'  - sp$phi a [np] vector of the random phases
#'  - sp$gamma a [np, nvar] matrix of the weights for the random waves
#'  @param st a [n, d] matrix of the d+1 coordinates of the n target space-time points
#'  @verbose a boolean
#'  @value a [n, nvar] matrix of the simulated values

compute_simulation <- function(sp, X, verbose = FALSE) {
  ns = dim(sp$omega)[1]
  nd = dim(sp$omega)[2]
  nv = dim(sp$lambda)[2]
  stopifnot(length(sp$phi) == ns)
  if (!is.matrix(sp$lambda)) {
    sp$lambda = matrix(sp$lambda, nrow = length(sp$lambda), ncol = 1)
  }
  stopifnot(dim(sp$lambda)[1] == ns)
  stopifnot(nd == dim(X)[2])
  np = dim(X)[1]
  if(verbose) {
    print(paste0(">>> Computing using the spectral method:"))
    print(paste0(">>> - number of dimensions  = ", nd))
    print(paste0(">>> - number of variables   = ", nv))
    print(paste0(">>> - number of frequencies = ", ns))
    print(paste0(">>> - number of points      = ", np))
  }
  t(sp$lambda) %*% (cos(sp$omega %*% t(X) + outer(X = sp$phi, Y = rep(1.0, np))))
}

# ------------------------------------------------------------------------------
# computing a simple and cross variogram on a 2d grid
#' @param val
#' @param val2
#' @param nval
# ------------------------------------------------------------------------------
# computing a spectral model on R^d
#--------------------------------------------------------------------------------------------
compute_vario_on_grid <- function(val, val2 = NULL, nlag = 10, dir = c(1,0), dx = c(1,1)) {
  nx = dim(val)
  gg = rep(NaN, nlag)
  hh = rep(NaN, nlag)
  nc = rep(NaN, nlag)
  dir_norm = sqrt(sum((dir*dx)^2))
  for (lag in 1:nlag) {
      idx_deb = max(1,1-lag*dir[1]):min(nx[1], nx[1]-lag*dir[1])
      jdx_deb = max(1,1-lag*dir[2]):min(nx[2], nx[2]-lag*dir[2])
      idx_end = idx_deb + lag*dir[1]
      jdx_end = jdx_deb + lag*dir[2]
      if (is.null(val2)) {
        tmp = 1/2 * as.numeric((val[idx_deb, jdx_deb] - val[idx_end, jdx_end])^2)
      } else {
        tmp = 1/2 * as.numeric((val[idx_deb, jdx_deb] - val[idx_end, jdx_end])*
                               (val2[idx_deb, jdx_deb] - val2[idx_end, jdx_end])
                              )
      }
      nc[lag] = length(tmp)
      if (nc[lag] > 0) {
          hh[lag] = dir_norm * lag
          gg[lag] = sum(tmp)/nc[lag]
      }
    }
  list(
    nc = nc,
    hh = hh,
    gg = gg
  )
}

# ------------------------------------------------------------------------------
#' compute univariate statistics on a vector of values
# ------------------------------------------------------------------------------
compute_statistics <- function(val, na.rm = TRUE) {
  res = c(
    length(!is.na(val)),
    min(val, na.rm = na.rm),
    max(val, na.rm = na.rm),
    median(val, na.rm = na.rm),
    mean(val, na.rm = na.rm),
    sd(val, na.rm = na.rm)
  )
  names(res) <- c("Number", "min", "max", "median", "mean", "sd")
  res
}

# ------------------------------------------------------------------------------
#' plot of a 2D section of a mono-variate model
# ------------------------------------------------------------------------------
#' @param simu
#' @param model
plot_simulation <- function(simu, model, x = 1:dim(simu)[1], y = 1:dim(simu)[1], idx_x = 1, idx_y = 2,
                            sub = NULL,
                            flag.hist = TRUE, flag.map = TRUE, flag.vario = TRUE) {
  # histogram
  if (flag.hist) {
    hist(simu, main = "")
    title(
      main = model$description(), font.main = 4,
      sub = sub, font.sub = 3, col.sub = "red")
  }
  # map
  if (flag.map) {
    image(x = x, y = y, z = simu, asp = 1, col = hcl.colors(100, "terrain"), axes = FALSE)
    contour(x, y, simu, levels = c(-1, 0, 1),
            add = TRUE, col = "brown")
    axis(1, at = seq(min(x), max(x), length.out = 5))
    axis(2, at = seq(min(y), max(y), length.out = 5))
    box()
    title(main = model$description(), sub = sub, font.main = 4, font.sub = 3, col.sub = "red")
  }
  # variograms
  if (flag.vario) {
    dir_col  = c("skyblue", "green", "orange", "red")
    dir_list = matrix(c(1,0, 1,1, 0,1, -1, 1), nrow = 4, ncol = 2, byrow = TRUE)
    dx       = c(x[2] - x[1], y[2] - y[1])
    dir_norm = apply(X = dir_list, MARGIN = 1, FUN = function(u) sqrt(sum((u*dx)^2)))
    nlag     = 30
    plot(NULL, NULL, xlim = c(0, nlag*max(dir_norm)), ylim = 1.1*c(0, model$variance()),
         xlab = "Distance", ylab = "Variogram")
    abline(v = c(0), col = "gray", lty = 3)
    abline(h = c(0, model$variance()), col = "gray", lty = 3)
    dir_lab = NULL
    for (idir in 1:dim(dir_list)[1]) {
      dir_lab = c(dir_lab, sprintf("[%d,%d]", dir_list[idir,1], dir_list[idir,2]))
      # experimental variogram
      gam = compute_vario_on_grid(val = simu, nlag = nlag, dir = dir_list[idir,], dx = dx)
      lines(gam$hh, gam$gg, col = dir_col[idir], lty = 2)
      points(gam$hh, gam$gg, col = dir_col[idir], pch = 19, cex = 0.2)
      # variogram model
      dir_norm = sqrt(sum((dir_list[idir,]*dx)^2))
      h  = (0:100)/100*nlag*dir_norm
      u0 = rep(0.0, model$getNDim())
      u0[idx_x] = dir_list[idir, 1]*dx[1]/dir_norm
      u0[idx_y] = dir_list[idir, 2]*dx[2]/dir_norm
      H = outer(X = h, Y = u0, FUN = "*")
      lines(h, as.vector(model$variance())-model$covariance(H), col = dir_col[idir], lty = 1)
    }
    legend("bottomright", legend = dir_lab, col = dir_col, lty = 1)
    title(main = model$description(), sub = sub, font.main = 4, font.sub = 3, col.sub = "red")
  }
  invisible(NULL)
}

```

### Covariance models

```{r my_functions_cov, echo=FALSE, eval=TRUE}
# ------------------------------------------------------------------------------
MyType_ = c(
  "EXPONENTIAL",
  "GAUSSIAN",
  "MATERN",
  "CAUCHY",
  "CAUCHY_GEN",
  "GNEITING_G",
  "GNEITING_M",
  "GNEITING_C")
MyClass_ = c(
  "CorAniso",
  "CorMatern",
  "CorGaussianMixture",
  "CorGneiting"
  )
# ------------------------------------------------------------------------------

# ------------------------------------------------------------------------------
# MyCorAniso implemented in CorAniso: ACov
# ------------------------------------------------------------------------------

# The covariance function is defined by an isotropic kernel and a geometrical anisotropy
MyCorAniso <- function(type = MyType_[3], sigma = 1.0, params = NULL,
                       anisotropy = MyAniso(), verbose = FALSE) {
  nvar = 1
  ndim = anisotropy$getNDim()
  kernel = MyKernel(type = type, d = ndim, params = params)
  fn_description = function() {
    desc = kernel$description()
    if (sigma != 1.0) {
      desc = bquote(paste(.(desc), "; ", sigma == .(round(sigma, 2))))
    }
    desc = bquote(paste(.(desc), "; ", .(anisotropy$description())))
   desc
  }
  # simulation
  fn_simulate <- function(n, seed = NULL, cov0 = NULL) {
    # simulating the spectrum
    if (!is.null(seed)) {
      set.seed(seed)
    }
    if (is.null(cov0)) {
      sp_ = list(
      lambda = matrix(sqrt(-2 * sigma * log(runif(n = n)) / n), nrow = n, ncol = nvar),
      omega  = matrix(kernel$simulate(n) %*% anisotropy$getMatrixDirect(), nrow = n, ncol = ndim),
      phi    = runif(n = n, min = 0, max = 2 * pi)
      )
      } else {
      # the spectrum is simulated using the auxiliary covariance cov0
      sp_ = cov0$simulate(n = n, seed = seed)
      # the importance weights are computed and the spectrum modified
      aux_sp = cov0$spectrum(U = sp_$omega)
      tar_sp = kernel$spectrum(anisotropy$normS(U = sp_$omega))*anisotropy$jacobian()
      sp_$lambda = matrix(sp_$lambda * sqrt(sigma/cov0$variance()*tar_sp/aux_sp),
                          nrow = n, ncol = nvar)
      }
    # -------------------- defining the methods
    fn_compute = function(dbout, verbose = FALSE, prefix = type, suffix = "simu") {
      X   = expand_dbgrid(dbout)
      val = compute_simulation(sp = sp_, X = X, verbose = verbose)
      isim = 1
      for (ivar in 1:nvar) {
        nm_var  = sprintf("%s.V%d.S%d", prefix, ivar, isim)
        dbout[nm_var] = val[ivar,]
      }
      NULL
    }
    list(
      spectrum = function() {sp_},
      compute  = fn_compute
    )
  }
  # -----------------------------------------
  if (verbose) {
    print(paste0("Creating a ", type, " CovAniso model:"))
    print(paste0(" - dimension of the space :", ndim))
    print(paste0(" - number of variables    :", nvar))
    print(paste0(" - variance               :", sigma))
    if (!is.null(params)) {
      for (ipar in 1:length(params)) {
      print(sprintf(" - param[%d] = %f", ipar, params[ipar]))
      }
    }
  }
  # -----------------------------------------
  list(
    type = function(){type},
    class = function(){"CovAniso"},
    description = fn_description,
    getNDim = function(){ndim},
    getNVar = function(){nvar},
    variance = function(ivar = 1, jvar = 1){sigma},
    covariance  = function(H, ivar = 1, jvar = 1) {sigma*kernel$covariance(anisotropy$normH(H))},
    spectrum = function(U, ivar = 1, jvar = 1) {sigma*kernel$spectrum(anisotropy$normS(U))*anisotropy$jacobian()},
    simulate = fn_simulate,
    # CovAniso
    getAnisotropy = function(){anisotropy},
    getParameters = function() {kernel$getParameters()},
    getKernel = function(){kernel}
  )
}


# ------------------------------------------------------------------------------
# MyCorMatern implemented in CorMatern: ACov
# ------------------------------------------------------------------------------
# The covariance function is defined by a geometrical anisotropy, and the parameters of
# the marginal Materns covariances (nu et kappa)
MyCorMatern <- function(type = MyType_[3],
                        nu = 1/2, kappa = rep(1.0, length(nu)), sigma = rep(1, length(nu)),
                        anisotropy = MyAniso(), verbose = FALSE) {
  stopifnot(type == MyType_[3])
  ndim = anisotropy$getNDim() # dimension of the euclidean space (R^d)
  nvar = length(nu) # number of variables
  # definition of the multivariate parameters
  stopifnot((length(kappa) == nvar) & all(kappa > 0.0))
  stopifnot((length(nu) == nvar) & all(nu > 0.0))
  stopifnot((length(sigma) == nvar) & all(sigma > 0.0))
  if (nvar == 1) {
    Nu_ = matrix(nu, 1, 1)
    Kappa_ = matrix(kappa, 1, 1)
    C0_ = matrix(sigma, 1, 1)
  }
  if (nvar > 1) {
    Nu_ = outer(X = 1:nvar, Y = 1:nvar, FUN = function(i, j) (nu[i] + nu[j])/2)
    Kappa_ = outer(X = 1:nvar, Y = 1:nvar, FUN = function(i, j) sqrt((kappa[i]^2 + kappa[j]^2)/2))
    C0_ = outer(X = sigma, Y = sigma, FUN = function(v1, v2){sqrt(v1*v2)}) *
      gamma(Nu_)/Kappa_^(2*Nu_) / sqrt((gamma(nu)/kappa^(2*nu)) %*% t(gamma(nu)/kappa^(2*nu)))
  }
  # C0_ is a positive definite matrix?
  stopifnot(is.matrix(C0_) & isSymmetric(C0_) & all(eigen(C0_)$values >= -1e-12) )
  # reference covariance function
  covRef_ = MyCorAniso(type = MyType_[3], sigma = 1.0, params = min(nu),
    anisotropy = MyAniso(
      scales = anisotropy$getScales()/max(kappa),
      rotation = anisotropy$getRotation())
    )

  #----------------------------------------------
  # methods
  #----------------------------------------------
    fn_description = function() {
    desc = sprintf("%s [%d, %d]", type, ndim, nvar)
    bquote(paste(.(desc), "; ", .(anisotropy$description())))
    }

  #----------------------------------------------
  matern_kernel <- function(h, nu) {
    stopifnot(nu > 0)
    h = abs(h)
    sel = (h > 0)
    cov = rep(1.0, length(h))
    cov[sel] <- h[sel]^nu * besselK(x = h[sel], nu = nu) / 2^(nu-1) / gamma(nu)
    cov
  }
  matern_spectral_density <- function(u, nu, d = 2) {
    stopifnot(nu > 0)
    gamma(nu + d/2)/gamma(nu)/(pi^(d/2))/(1 + u^2)^(nu + d/2)
  }

  fn_covariance = function(H, ivar = 1, jvar = 1) {
    C0_[ivar,jvar]*matern_kernel(anisotropy$normH(H)*Kappa_[ivar,jvar], nu = Nu_[ivar,jvar])
  }

  fn_spectrum = function(U, ivar = 1, jvar = 1) {
    C0_[ivar,jvar]*matern_spectral_density(
      u = anisotropy$normS(U)/Kappa_[ivar,jvar],
      nu = Nu_[ivar,jvar],
      d = ndim)*anisotropy$jacobian()/Kappa_[ivar,jvar]^ndim
  }
  fn_spectrumMat <- function(u) {
    H = matrix(NaN, nrow = nvar, ncol = nvar)
    for (ivar in 1:nvar) {
      for (jvar in 1:nvar) {
        H[ivar, jvar] = fn_spectrum(U = t(u), ivar = ivar, jvar = jvar)
      }
    }
    H
  }
  #----------------------------------------------
  fn_simulate <- function(n, seed = NULL, cov0 = NULL) {
    sp0 = covRef_$simulate(n = n, seed = seed)$spectrum()
    # *square* root of a positive definite matrix
    sqrt_mat <- function(A) {
      if (is.matrix(A)&dim(A)[1] > 1) {
        VD = eigen(A)
        VD$vectors %*% diag(sqrt(pmax(VD$values, 0))) %*% t(VD$vectors)
      } else {
        sqrt(A)
      }
    }
    # weighting the spectral components for the p variables
    Lambda = outer(X = as.numeric(sp0$lambda), Y = rep(sqrt(nvar), nvar), FUN = "*")
    sp0_spectrum = covRef_$spectrum(U = sp0$omega)
    for (i in 1:n) {
      H = fn_spectrumMat(u =sp0$omega[i,])/sp0_spectrum[i]
      H = sqrt_mat(H)
      Lambda[i,] <- Lambda[i,] * H[, sample(x = 1:p, size = 1)]
    }
    sp_ = list(lambda = Lambda, omega = sp0$omega, phi = sp0$phi)
    # -------------------- defining the methods
    fn_compute = function(dbout, verbose = FALSE, prefix = type, suffix = "simu") {
      X   = expand_dbgrid(dbout)
      val = compute_simulation(sp = sp_, X = X, verbose = verbose)
      isim = 1
      for (ivar in 1:nvar) {
        nm_var  = sprintf("%s.V%d.S%d", prefix, ivar, isim)
        dbout[nm_var] = val[ivar,]
      }
      NULL
    }
    list(
      spectrum = function() {sp_},
      compute  = fn_compute
    )
  }
  # -----------------------------------------
  if (verbose) {
    print(paste0("Creating a ", type, " CovMatern model:"))
    print(paste0(" - dimension of the space :", ndim))
    print(paste0(" - number of variables    :", nvar))
    for (ivar in 1:nvar) {
        print(sprintf(" - variance[%d] = %.2f", ivar, C0_[ivar, ivar]))
        print(sprintf(" - nu[%d]       = %.2f", ivar, Nu_[ivar, ivar]))
        print(sprintf(" - kappa[%d] = %.2f", ivar, Kappa_[ivar, ivar]))
      }
    }

  # -----------------------------------------
  list(
    type = function(){type},
    class = function(){"CovMatern"},
    description = fn_description,
    getNDim = function(){ndim},
    getNVar = function(){nvar},
    variance    = function(ivar = 1, jvar = 1){C0_[ivar, jvar]},
    covariance  = fn_covariance,
    spectrum = fn_spectrum,
    simulate = fn_simulate,
    # CovMatern
    getCovRef = function(){covRef_},
    getAnisotropy = function(){anisotropy},
    getNus = function() {Nu_},
    getKappas = function(){Kappa_}
  )
}

# ------------------------------------------------------------------------------
# MyCorGaussianMixture implemented in CorGaussianMixture: ACov
# ------------------------------------------------------------------------------
MyCorGaussianMixture <- function(type = MyType_[3], nu = 1/2, kappa = rep(1.0, length(nu)),
                              sigma = rep(1.0, length(nu)),
                              anisotropy = MyAniso(), verbose = FALSE
                              ) {
  stopifnot(type %in% MyType_[1:4]) # EXPONENTIAL/GAUSSIAN/MATERN/CAUCHY
  # EXPONENTIAL and GAUSSIAN are mono-variate
  if (type == MyType_[1]) { # EXPONENTIAL
    nu = 1/2
    kappa = kappa[1]
  }
  if (type == MyType_[2]) { # GAUSSIAN
    nu    = 1.0
    kappa = kappa[1]
  }
  ndim  = anisotropy$getNDim()
  nvar   = length(nu)
  stopifnot(length(kappa) == nvar)
  stopifnot(length(sigma) == nvar)
  # attributes
  corRef_ = MyCorAniso(type = type, sigma = 1, params = nu)
  Nu_    = outer(X = nu, Y = nu, FUN = function(nu_i, nu_j) {(nu_i + nu_j)/2})
  C0_    = outer(X = sigma, Y = sigma, FUN = function(v1, v2){sqrt(v1*v2)})
  if (type %in% MyType_[c(2)]) { # Gaussian (nvar == 1)
    Kappa_ = matrix(kappa, nrow = nvar, ncol = nvar)
  }
  if (type %in% MyType_[c(1,3)]) { # Matern
    Kappa_ = outer(X = kappa, Y = kappa, FUN = function(ri, rj) {sqrt((ri^2 + rj^2)/2)})
    C0_    = C0_ * gamma(Nu_) / sqrt(gamma(nu) %*% t(gamma(nu))) *
      ((kappa^nu %*% t(kappa^nu)) / Kappa_^(2*Nu_))
  }
  if (type == MyType_[4]) { # Cauchy
    Kappa_ = outer(X = kappa, Y = kappa, FUN = function(ri, rj) {1/sqrt((1/ri^2 + 1/rj^2)/2)})
    C0_    = C0_ * gamma(Nu_) / sqrt(gamma(nu) %*% t(gamma(nu))) /
      ((kappa^nu %*% t(kappa^nu)) / Kappa_^(2*Nu_))
  }
  # C0_ is a positive definite matrix?
  stopifnot(is.matrix(C0_) & isSymmetric(C0_) & all(eigen(C0_)$values >= -1e-12) )
  #----------------------------------------------
  # methods
  #----------------------------------------------
    fn_description = function() {
    desc = sprintf("%s[%d, %d]", type, ndim, nvar)
    bquote(paste(.(desc), "; ", .(anisotropy$description())))
    }
  fn_parameters <- function(case = c("nu", "kappa", "var")[3], ivar = NULL, jvar = NULL) {
    if (is.null(ivar)|is.null(jvar)) {
      if (case == "nu") {
        Nu_
      } else if (case == "kappa") {
        Kappa_
      } else if (case == "var") {
        C0_
      }
    } else {
      stopifnot((ivar %in% 1:nvar)&(jvar %in% 1:nvar))
      if (case == "nu") {
        Nu_[ivar, jvar]
      } else if (case == "kappa") {
        Kappa_[ivar, jvar]
      } else if (case == "var") {
        C0_[ivar, jvar]
      }
    }
  }
  fn_variance <- function(ivar = NULL, jvar = NULL) {
    fn_parameters(case = "var", ivar, jvar)
  }
  fn_reduce <- function(ivar = 1, jvar = ivar) {
    MyCorAniso(type = type, sigma = C0_[ivar, jvar], params = Nu_[ivar, jvar],
        anisotropy = MyAniso(
                scales = anisotropy$getScales() / Kappa_[ivar,jvar],
                rotation = anisotropy$getRotation())
        )
  }
  fn_covariance <- function(H, ivar = 1, jvar = 1) {
      stopifnot((is.matrix(H))&(dim(H)[2] == ndim))
      stopifnot((ivar %in% 1:nvar)&(jvar %in% 1:nvar))
      fn_reduce(ivar, jvar)$covariance(H)
    }
  fn_simulate <- function(n, seed = NULL, nu0 = NULL, kappa0 = NULL) {
    if (is.null(nu0)) {nu0 = min(nu)}
    if (is.null(kappa0)) {kappa0 = max(kappa)}
    if (!is.null(seed)) {set.seed(seed)}
    ratio = matrix(1.0, nrow = n, ncol = nvar)
    if (type %in% MyType_[c(2)]) { # Gaussian
      xi_    = rep(1.0, n)
      for (ivar in 1:nvar) {
        ratio[,ivar] = sigma[ivar] * rep(1.0, n)
      }
    }
    if (type %in% MyType_[c(1,3)]) { # Exponential or Matern
      xi_    = 1/rgamma(n = n, shape = nu0, rate = (kappa0/2)^2)
      daux   = density_IG(xi_, alpha = nu0, beta = (kappa0/2)^2)
      for (ivar in 1:nvar) {
        ratio[,ivar] = sigma[ivar]*density_IG(xi_, alpha = nu[ivar], beta = (kappa[ivar]/2)^2) / daux
      }
    } else if (type == MyType_[4]) { # Cauchy
      xi_  = rgamma(n = n, shape = nu0, rate = (1/kappa0^2))
      log_daux  = dgamma(xi_, shape = nu0, rate = 1/kappa0^2, log = TRUE)
      for (ivar in 1:nvar) {
        ratio[,ivar] = sigma[ivar]*exp(dgamma(xi_, shape = nu[ivar], rate = 1/kappa[ivar]^2, log = TRUE) - log_daux)
      }
    } else if (type == MyType_[2]) { # Gaussian
      xi_  = kappa[ivar]^2
    }

  omega0_ = matrix(rnorm(n*ndim), nrow = n, ncol = ndim) * sqrt(2*xi_)
  sp_ = list(
    lambda = sqrt(-2*log(runif(n))/n*ratio),
    omega  = (omega0_)%*%anisotropy$getMatrixDirect(),
    phi    = runif(n = n, min = 0, max = 2*pi),
    xi     = xi_,
    omega0 = omega0_
  )
    # -------------------- defining the methods
    fn_compute = function(dbout, verbose = FALSE, prefix = type, suffix = "simu") {
      X   = expand_dbgrid(dbout)
      val = compute_simulation(sp = sp_, X = X, verbose = verbose)
      isim = 1
      for (ivar in 1:nvar) {
        nm_var  = sprintf("%s.V%d.%Sd", prefix, ivar, isim)
        dbout[nm_var] = val[ivar,]
      }
      NULL
    }
    list(
      spectrum = function() {sp_},
      compute  = fn_compute
    )
  }
  fn_spectrum = function(U, ivar = 1, jvar = 1) {
    "Not yet implemented (TODO)"
  }
  # -----------------------------------------
  if (verbose) {
    print(paste0("Creating a ", type, " GaussianMixture model:"))
    print(paste0(" - dimension of the space :", ndim))
    print(paste0(" - number of variables    :", nvar))
    for (ivar in 1:nvar) {
        print(sprintf(" - variance[%d] = %.2f", ivar, C0_[ivar, ivar]))
        print(sprintf(" - nu[%d]       = %.2f", ivar, Nu_[ivar, ivar]))
        print(sprintf(" - kappa[%d] = %.2f", ivar, Kappa_[ivar, ivar]))
      }
    }
  # -----------------------------------------
  list(
    type = function(){type},
    class = function(){"GaussianMixture"},
    description = fn_description,
    getNDim = function(){ndim},
    getNVar = function(){nvar},
    variance    = function(ivar = 1, jvar = 1){C0_[ivar, jvar]},
    covariance  = fn_covariance,
    spectrum = fn_spectrum,
    simulate = fn_simulate,
    reduce      = fn_reduce,
    # Gaussian mixture
    getAnisotropy = function(){anisotropy},
    getNus = function() {Nu_},
    getKappas = function(){Kappa_},
    getParameters = fn_parameters
  )
}

# ------------------------------------------------------------------------------
# MyCovGneiting implemented in CorGneiting: ACov
# ------------------------------------------------------------------------------
# Gneiting model for $d %\in% c(1,2,3)$
# Gneiting model for $type %\in% c(6,7,8)$  Spatial covariance is Gaussian, Matern, or Cauchy
MyCorGneiting <- function(type = MyType_[7], nu = 1/2, kappa = rep(1.0, length(nu)), sigma = rep(1.0, length(nu)),
                            anisotropy = MyAniso(),
                            alpha = 1.0, beta = 1.0, timeRange = 1.0,
                            verbose = TRUE
                            ) {
  # checking parameters
  stopifnot(type %in% MyType_[6:8])
  if (type %in% MyType_[6]) typeS = MyType_[2] # Gaussian space trace
  if (type %in% MyType_[7]) typeS = MyType_[3] # Matern space trace
  if (type %in% MyType_[8]) typeS = MyType_[4] # Cauchy space trace
  # defining the spatial trace
  covS = MyCorGaussianMixture(type = typeS, nu = nu, kappa = kappa, sigma = sigma,
                           anisotropy = anisotropy,
                           verbose = verbose)
  ndim = covS$getNDim() # number of spatial dimensions
  stopifnot(ndim %in% 1:3)
  nvar = covS$getNVar()
  stopifnot(nvar > 0)

  # defining the generalized Cauchy temporal covariance
  stopifnot((alpha > 0.0)&(alpha <= 2.0))
  stopifnot((beta > 0.0)&(beta <= 1.0))
  stopifnot((timeRange > 0.0))
  covT = MyCorAniso(type = MyType_[5],
                     anisotropy = MyAniso(scales = timeRange),
                     sigma = 1.0, params = c(alpha, beta*ndim/2),
                     verbose = verbose)

  # ----------------------------------------------------------
  fn_type = function(){type}
  # -------------- description
  fn_description = function() {
    desc = sprintf("%s[%d, %d]", type, ndim, nvar)
    # desc = bquote(paste(.(desc), "; ", .(anisotropy$description())))
    # desc = bquote(paste(.(desc), "; ", a_t == .(timeRange)))
    desc = bquote(paste(.(desc), "; ", alpha == .(alpha), "; ", beta == .(beta)))
    if (nvar == 1) {
      desc = bquote(paste(.(desc), "; ", nu  == .(round(nu,1)), "; ", kappa == .(round(kappa,1))))
      }
    desc
    }
  # -------------- space-time covariance function
  fn_covariance <- function(H, ivar = 1, jvar = 1) {
    stopifnot(is.matrix(H)&(ncol(H) == ndim+1))
    h = matrix(H[,1:ndim],  nrow = nrow(H), ncol = ndim)
    u = matrix(H[, ndim+1], nrow = nrow(H), ncol = 1)
    # temporal covariance
    ct = covT$covariance(u, ivar = 1, jvar = 1)
    # spatial covariance
    ct * covS$covariance(H = h * ct^(1/ndim), ivar = ivar, jvar = jvar)
  }

  #  -------------- simulation
  fn_simulate <- function(n, seed = NULL, nu0 = NULL, kappa0 = NULL) {
    if(!is.null(seed)) {set.seed(seed)}
    # simulating the spatial components
    spS_ = covS$simulate(n = n, seed = NULL, nu0 = nu0, kappa0 = kappa0)$spectrum()

    # simulating the temporal frequency
    lambda = (apply(X = as.matrix(spS_$omega0),
                   MARGIN = 1,
                   FUN = function(u){sum(u^2)})/(4*spS_$xi))^(1/beta)
    sim_S = rep(NaN, length(lambda))

    flag.double = TRUE
    for (i in seq_along(lambda)) {
      sim_S[i] = simulate_stable_unilateral_exptilted(n = 1, alpha = beta, lambda = lambda[i],
                                                   flag.double = flag.double)$X
    }
    sim_T = simulate_stable_bilateral(n = n, alpha = alpha)

    # combining spatial components and time frequencies
    omega = matrix(NaN, nrow = n, ncol = ndim + 1)
    omega[, 1:ndim] = spS_$omega0[, 1:ndim] %*% anisotropy$getMatrixDirect()
    omega[, 1+ndim] = (sim_S*lambda)^(1/alpha) * sim_T / timeRange
    # storing the simulated spectrum
    sp_ = list(
      lambda = spS_$lambda,
      omega  = omega,
      phi    = spS_$phi
    )
    # -------------------- defining the methods
    fn_compute = function(dbout, verbose = FALSE, prefix = type, suffix = "simu") {
      X   = expand_dbgrid(dbout)
      val = compute_simulation(sp = sp_, X = X, verbose = verbose)
      isim = 1
      for (ivar in 1:nvar) {
        nm_var  = sprintf("%s.V%d.S%d", prefix, ivar, isim)
        dbout[nm_var] = val[ivar,]
      }
      NULL
    }
    list(
      spectrum = function() {sp_},
      compute  = fn_compute
    )
  }
  # ----------------------------------------------------------
  if (verbose) {
    print(paste0("Creating a space-time ", fn_type(), " Gneiting model:"))
    print(paste0(" - dimension of the model :", ndim+1))
    print(paste0(" - dimension of the space :", ndim))
    print(paste0(" - number of variables    :", nvar))
    print(paste0(" - alpha                  :", alpha))
    print(paste0(" - beta                   :", beta))
    print(paste0(" - a_t                    :", timeRange))
  }
  # ----------------------------------------------------------
    list(
    type = fn_type,
    class = function(){"CovGneiting"},
    description = fn_description,
    getNDim = function(){ndim+1},
    getNVar = function(){nvar},
    variance = function(ivar = 1, jvar = 1){covS$variance(ivar = ivar, jvar = jvar)},
    covariance = fn_covariance,
    spectrum = function(U, ivar=1, jvar=1) {
      print("Not yet implemented (TODO)")
      NULL},
    simulate = fn_simulate,
    # CovGneiting
    covS = function() {covS},
    getAnisotropy = function(){covS$getAnisotropy()},
    getNus = function() {covS$getNus()},
    getKappas = function(){covS$getKappas()},
    getParameters = function(case = c("nu", "kappa", "var")[3], ivar = NULL, jvar = NULL) {
      covS$getParameters(case = case, ivar = ivar, jvar = jvar)
    },
    covT = function(){covT},
    alpha = function(){alpha},
    beta  = function(){beta},
    timeRange = function(){timeRange}
  )
}


# ------------------------------------------------------------------------------
# MySimu Spectral is implemented in the simuSpectral function
# ------------------------------------------------------------------------------
MySimuSpectral <- function(grid, cov, nsim = 1, ns = 1000, seed = NULL, verbose = FALSE, cov0 = NULL,
                           prefix = "MySimu") {
  ierr = 0
  ndim = cov$getNDim()
  stopifnot(grid$getNDim() == ndim)
  nvar = cov$getNVar()
  if (!is.null(seed)) {set.seed(seed)}
  if (verbose) {
    print(sprintf("Simulation of model %s:", cov$type()))
    print(sprintf(" Number of dimensions  = %d", ndim))
    print(sprintf(" Number of variables   = %d", nvar))
    print(sprintf(" Number of simulations = %d", nsim))
    print(sprintf(" Number of frequencies = %d", ns))
  }
  for (isim in 1:nsim) {
    if (verbose) {
      print(sprintf("Computing simulation #%d(/%d)", isim, nsim))
    }
    simu = cov$simulate(n = ns, seed = NULL)
    ierr = simu$compute(dbout = grid, verbose = verbose, prefix = prefix, suffix = sprintf("S%d", isim))
  }
  ierr
}

```

# Test of the models

```{r covariance_test, echo = FALSE, eval = TRUE}
# global parameters
flag.gstlearn = TRUE
flag.verbose  = FALSE
flag.plot     = TRUE

# geometrical anisotropy of the spatial trace
ndim   = 2 # ndim %in% 1:3
ani_scales = c(3/4, 3/2, 3/4)[1:ndim]
ani_angles = c(30, 0, 0)[1:ndim]
ani0 = MyAniso(scales = ani_scales, rotation = MyRotation(angles = ani_angles))

# tests to be performed
list_test = list()
# CorAniso
list_test[[1+length(list_test)]] <- list(idx_class = 1, idx_type = 1, nvar = 1)
list_test[[1+length(list_test)]] <- list(idx_class = 1, idx_type = 2, nvar = 1)
list_test[[1+length(list_test)]] <- list(idx_class = 1, idx_type = 3, nvar = 1)
list_test[[1+length(list_test)]] <- list(idx_class = 1, idx_type = 4, nvar = 1)
# CorMatern (mono/bi variate)
list_test[[1+length(list_test)]] <- list(idx_class = 2, idx_type = 3, nvar = 1)
list_test[[1+length(list_test)]] <- list(idx_class = 2, idx_type = 3, nvar = 2)
# CorGaussianMixture (Gaussian - mono)
list_test[[1+length(list_test)]] <- list(idx_class = 3, idx_type = 2, nvar = 1)
# CorGaussianMixture (Matern - mono/bi)
list_test[[1+length(list_test)]] <- list(idx_class = 3, idx_type = 3, nvar = 1)
list_test[[1+length(list_test)]] <- list(idx_class = 3, idx_type = 3, nvar = 2)
# CorGaussianMixture (Cauchy - mono/bi)
list_test[[1+length(list_test)]] <- list(idx_class = 3, idx_type = 4, nvar = 1)
list_test[[1+length(list_test)]] <- list(idx_class = 3, idx_type = 4, nvar = 2)
# CorGneiting (Gaussian - mono)
list_test[[1+length(list_test)]] <- list(idx_class = 4, idx_type = 6, nvar = 1)
# CorGneiting (Matern - mono/bi)
list_test[[1+length(list_test)]] <- list(idx_class = 4, idx_type = 7, nvar = 1)
list_test[[1+length(list_test)]] <- list(idx_class = 4, idx_type = 7, nvar = 2)
# CorGneiting (Cauchy - mono/bi)
list_test[[1+length(list_test)]] <- list(idx_class = 4, idx_type = 8, nvar = 1)
list_test[[1+length(list_test)]] <- list(idx_class = 4, idx_type = 8, nvar = 2)

# parameters for the plot of the covariance
dd = seq(from = 0.0, to = 5, length.out = 1000)
dir_angles = c(0, 45, 90, 135)
dir_list = matrix(c(1,0, 1, 1, 0,1, -1, 1), nrow = 4, ncol = 2, byrow = TRUE)
ndir = nrow(dir_list)
col_dir = c("skyblue", "green", "orange", "red")[1:ndir]
lab_inc = NULL
lab_s_dir = NULL
lab_st_dir = NULL

# defining the grids in space R^3 and time
x0_st = rep(0.0, 4)
nx_st = rep(200, 4)
dx_st = rep(0.1, 4)
xx = list()
for (idim in 1:4) {
  xx[[1+length(xx)]] = x0_st[idim] + (0:(nx_st[idim]-1))*dx_st[idim]
}

# parameters for the simulations
ns   = 500    # number of spectral components
nsim = 2      # number of simulations
nlag = 30     # number of lags of the experimental variograms on grid
seed = 258425
set.seed(seed)

# slice in R^2 \times R (1:xOy, 2:xOt, 3:yOt)
idx_slice = 2
stopifnot(idx_slice %in% 1:3)

# ==============================================================================
for (id in seq_len(length(list_test))) {
# ==============================================================================
  # defining the model
  gs_class = MyClass_[list_test[[id]]$idx_class]
  type     = MyType_[list_test[[id]]$idx_type]
  nvar     = list_test[[id]]$nvar
  # spatial trace (mono/multi variate)
  nus    = c(3/4, 1/2)[1:nvar]
  kappas = c(2, 1)[1:nvar]
  # time trace
  alpha  = 2.0
  beta   = 1.0
  timeRange    = 1.0

  print(sprintf(" Begining of test #%d:",  id))
  cat("\n")
  print(sprintf(" - Class %s", gs_class))
  cat("\n")
  print(sprintf(" - Model %s", type))
  cat("\n")
  print(sprintf(" - Number of variables %d", nvar))
  cat("\n")

  # checking the consistency
  stopifnot(
    ((gs_class == "CorAniso")&(type %in% MyType_[1:4])&(nvar == 1))|
    ((gs_class == "CorMatern")&(type %in% MyType_[3]))|
    ((gs_class == "CorGaussianMixture")&(type %in% MyType_[1:4]))|
    ((gs_class == "CorGneiting")&(type %in% MyType_[6:8]))
      )

  # ----------------------------------------------------------------------------
  print(sprintf("Creation of the model"))
  # ----------------------------------------------------------------------------
  # creation of the model
  gs_ctxt = CovContext(ndim = ndim, nvar = nvar)
  gs_type = ECov_fromKey(type)
  if (gs_class == "CorAniso") {
    my_cov = MyCorAniso(type = type, sigma = 1.0, params = nus, anisotropy = ani0, verbose = TRUE)
    gs_cov = CorAniso_createAnisotropic(ctxt = gs_ctxt, type = gs_type,
      ranges = ani_scales,
      param = nus,
      angles = ani_angles,
      flagRange = FALSE)
  } else if (gs_class == "CorMatern") {
    my_cov = MyCorMatern(type = type, nu = nus, kappa = kappas, anisotropy = ani0)
    gs_cov = CorMatern_create(ctxt = gs_ctxt, type = gs_type,
      params = nus,
      kappas = kappas,
      ranges = ani_scales,
      angles = ani_angles,
      flagRange = FALSE)

  } else if (gs_class == "CorGaussianMixture") {
    my_cov = MyCorGaussianMixture(type = type, nu = nus, kappa = kappas, anisotropy = ani0)
    gs_cov = CorGaussianMixture_create(ctxt = gs_ctxt, type = gs_type,
      params = nus,
      kappas = kappas,
      ranges = ani_scales,
      angles = ani_angles,
      flagRange = FALSE)
  } else if (gs_class == "CorGneiting") {
    my_cov = MyCorGneiting(type = type, nu = nus, kappa = kappas, anisotropy = ani0,
      alpha = alpha, beta = beta, timeRange = timeRange,
      verbose = TRUE)
    gs_cov = CorGneiting_create(ctxt = gs_ctxt, type = gs_type,
      params = nus,
      kappas = kappas,
      alpha  = alpha,
      beta   = beta,
      timeRange = timeRange,
      ranges = ani_scales,
      angles = ani_angles,
      separability = 1.0,
      flagRange = FALSE
      )
  }
  prefix = my_cov$type()

  # ----------------------------------------------------------------------------
  print(sprintf("Testing the covariance"))
  cat("\n")
  # ----------------------------------------------------------------------------
  s_u0  = dir_list
  s_u0  = s_u0 / apply(X = s_u0, MARGIN = 1, FUN = function(u)sqrt(sum(u^2)))
  st_u0 = matrix(c(1,0,1, 1,1,1, 0,1,1, 0,0,1), nrow = 4, ncol = 3, byrow = TRUE)
  st_u0 = st_u0 / apply(X = st_u0, MARGIN = 1, FUN = function(u)sqrt(sum(u^2)))
  lab_inc    = NULL
  lab_s_dir  = NULL
  lab_st_dir = NULL
  for (idir in 1:ndir) {
    # inc
    lab_inc = c(lab_inc,
                sprintf("[%d,%d]", dir_list[idir,1], dir_list[idir,2]))
    # space
    lab_s_dir = c(lab_s_dir,
                  sprintf("[%.2f,%.2f]", s_u0[idir,1], s_u0[idir,2]))
    # space x time
    lab_st_dir = c(lab_st_dir,
                   sprintf("[%.2f,%.2f,%.2f]", st_u0[idir,1], st_u0[idir,2], st_u0[idir,3]))
    }

    for (ivar in 1:nvar) {
      for (jvar in ivar:nvar) {
        my_cov_var = reduce_model(my_cov, ivar, jvar)

        # --------------------------------------------------------------------------
        # spatial trace...
        # --------------------------------------------------------------------------
        sub = sprintf("%s (Var-%d, Var-%d) in space", gs_class, ivar, jvar)
        if (flag.plot) {
          plot(NULL, NULL, xlim = range(dd), ylim = 1.1*c(0, my_cov_var$variance()),
               xlab = "Distance", ylab = "Covariance",
               main = my_cov_var$description(),
               sub = sub , col.sub = "red")
              abline(h = c(0, my_cov_var$variance()), col = "gray", lty = 3)
              abline(v = c(0), col = "gray", lty = 3)
        }

        for (idir in 1:length(dir_angles)) {
          # computing the covariance
          u0 = as.numeric(MyRotation(angles = c(dir_angles[idir],0))$applyDirect(u = c(1,0)))
          if (my_cov$type() %in% MyType_[6:8]) {u0 = c(u0, 0.0)}
          val_R  = my_cov$covariance(H = dd %*% t(u0), ivar = ivar, jvar = jvar)
          val_gs = NULL
          if (flag.gstlearn) {
            val_gs = gs_cov$sample(h = dd, codir = u0, ivar = ivar - 1, jvar = jvar - 1)
          }
          if (flag.plot) {
          lines(dd, val_R, col = col_dir[idir], lty = 1)
          if (flag.gstlearn) {
            lines(dd, val_gs, col = "black", lty = 2)
            }
          }
          if(all(abs(val_gs - val_R) < 1.e-10)) {res_test = "OK"
          } else { res_test = "KO"}
           print(sprintf("Test in dir #%d for %s: %s", idir, sub, res_test))
           cat("\n")
        }
        if (flag.plot) {
          legend("topright",
                 legend = c(lab_s_dir, "R", "gst"),
                 col = c(col_dir, "gray", "black"),
                 lty = c(rep(1, ndir), 1, 2))
        }

        # --------------------------------------------------------------------------
        # space-time if GNEITING
        # --------------------------------------------------------------------------
        if (my_cov$type() %in% MyType_[6:8]) {
          sub = sprintf("%s (Var-%d, Var-%d) in space x time", gs_class, ivar, jvar)
          if (flag.plot) {
            plot(NULL, NULL, xlim = range(dd), ylim = 1.1*c(0, my_cov_var$variance()),
                 xlab = "Distance", ylab = "Covariance",
                 main = my_cov_var$description(),
                 sub = sub, col.sub = "red")
                abline(h = c(0, my_cov_var$variance()), col = "gray", lty = 3)
                abline(v = c(0), col = "gray", lty = 3)
          }
          for (idir in 1:length(dir_angles)) {
            u0 = st_u0[idir,]
            val_R = my_cov$covariance(H = dd %*% t(u0), ivar = ivar, jvar = jvar)
            val_gs = NULL
            if (flag.gstlearn) {
              val_gs = gs_cov$sample(h = dd, codir = u0, ivar = ivar - 1, jvar = jvar - 1)
            }
            if (flag.plot) {
              lines(dd, val_R, col = col_dir[idir], lty = 1)
              if (flag.gstlearn) {
                lines(dd, val_gs, col = "black", lty = 2)
              }
            }
          if(all(abs(val_gs - val_R) < 1.e-10)) {
           res_test = "OK"
          } else { res_test = "KO"}
           print(sprintf("Test in dir #%d for %s: %s", idir, sub, res_test))
           cat("\n")
          }
          if (flag.plot) {
          legend("topright",
                 legend = c(lab_st_dir, "R", "gst"),
                 col = c(col_dir, "gray", "black"),
                 lty = c(rep(1, ndir), 1, 2))
          }

        } # time space covariance
      } # ivar
    } # jvar

  # ----------------------------------------------------------------------------
  print(sprintf("Defining the 2d slice"))
  cat("\n")
  # ----------------------------------------------------------------------------
  if (my_cov$type() %in% MyType_[1:4]){ # spatial covariance
    slice = "xOy"
    x0 = x0_st[1:ndim]
    nx = nx_st[1:ndim]
    dx = dx_st[1:ndim]
    idx_x = 1
    idx_y = 2
    }
  if (my_cov$type() %in% MyType_[6:8]){ # space-time covariance (Gneiting's model)
    slice = c("xOy", "xOt", "yOt")[idx_slice]
    x0 = x0_st[c(1:ndim,4)]
    dx = dx_st[c(1:ndim,4)]
    if (slice == "xOy") {
      nx = c(nx_st[1:ndim], 1)
      idx_x = 1
      idx_y = 2
      }
    if (slice == "xOt") {
      nx = c(nx_st[1], 1, nx_st[4])
      idx_x = 1
      idx_y = 3
      }
    if (slice == "yOt") {
      nx = c(1, nx_st[2], nx_st[4])
      idx_x = 2
      idx_y = 3
      }
  }
  print(sprintf("Computing simulation in slice %s", slice))
  cat("\n")

  grid = DbGrid_create(x0 = x0, nx = nx, dx = dx)


  # ----------------------------------------------------------------------------
  print(sprintf("Performing a single simulation"))
  cat("\n")

  # ----------------------------------------------------------------------------
  err  = grid$deleteColumns(names = paste(prefix, "*", sep = "."))
  if (flag.gstlearn) {
    # using the interface
    modelGeneric = ModelGeneric(gs_ctxt)
    err = modelGeneric$setCov(gs_cov)
    err = simuSpectral(dbin = NULL, dbout = grid, model = modelGeneric, nbsimu = 1, ns = ns,
                       verbose = flag.verbose, seed = seed,
                       namconv = NamingConvention(prefix))
  } else {
    # using the R functions
    sim  = my_cov$simulate(n = ns, seed = seed)
    val  = sim$compute(dbout = grid, verbose = flag.verbose, prefix = prefix)
  }

  # ------------------
  # QC of the simulation (statistics and variograms)
  # ------------------

  tab  = NULL
  tab_var = NULL
  isim = 1
  for (ivar in 1:nvar) {
    nm_var  = sprintf("%s.V%d.S%d", prefix, ivar, isim)
    tab_var = c(tab_var, nm_var )
    val = grid[nm_var]
    # compute statistics
    tab = rbind(tab, t(compute_statistics(val)))
    # plots
    if (flag.plot) {
      my_cov_var = reduce_model(my_cov, ivar = ivar, jvar = ivar, verbose = TRUE)
      plot_simulation(
        simu  = matrix(data = val, nrow = nx[idx_x], ncol = nx[idx_y]),
        model = my_cov_var, x = xx[[idx_x]], y = xx[[idx_y]],
        sub = sprintf("%s in section %s", gs_class, slice), idx_x = idx_x, idx_y = idx_y,
        flag.hist = TRUE, flag.map = TRUE, flag.vario = TRUE
        )
    }
  }
  rownames(tab) <- tab_var
  print(knitr::kable(tab, digits = 4,
               caption = paste0("Mono-variate statistics for the ", my_cov$type(), " covariance")))
  cat("\n")
  # ----------------------------------------------------------------------------
  print(sprintf("Average variogram on various simulations"))
  cat("\n")
  # ----------------------------------------------------------------------------
  err  = grid$deleteColumns(names = paste(prefix, "*", sep = "."))
  if (flag.gstlearn) {
    # using the *gstlearn* interface
    modelGeneric = ModelGeneric(gs_ctxt)
    err = modelGeneric$setCov(gs_cov)
    err = simuSpectral(dbin = NULL, dbout = grid, model = modelGeneric, nbsimu = nsim, ns = ns,
                       verbose = flag.verbose,
                       namconv = NamingConvention(prefix))
  } else {
    # using the *R* interface
    err = MySimuSpectral(grid = grid, cov = my_cov, nsim = nsim, ns = ns,
                         verbose = flag.verbose,
                         prefix = prefix)

  }

  # ------------------
  # QC of simulations
  # ------------------

  # computing the variograms and statistics on the simulations
  ngam  = nvar*(nvar+1)/2
  gg = array(data = 0.0, dim = c(nlag, ndir, nsim, ngam))
  hh = array(data = 0.0, dim = c(nlag, ndir, nsim, ngam))
  tab = NULL
  var_name = NULL
  for (isim in 1:nsim) {
    print(sprintf(">>> Simulation #%d", isim))
    cat("\n")
    # processing the simulation
    ig = 0
    for (ivar in 1:nvar) {
      nm_vari  = sprintf("%s.V%d.S%d", prefix, ivar, isim)
      sim_vari = matrix(grid[nm_vari], nx[idx_x], nx[idx_y])
      # statistics
      var_name = c(var_name, nm_vari)
      tab = rbind(tab, t(compute_statistics(grid[nm_vari])))

      for (jvar in ivar:nvar) {
        ig = ig+1
        nm_varj = sprintf("%s.V%d.S%d", prefix, jvar, isim)
        sim_varj = matrix(grid[nm_varj], nx[idx_x], nx[idx_y])
        # variogram
        for (idir in 1:nrow(dir_list)) {
        gam = compute_vario_on_grid(val = sim_vari, val2 = sim_varj,
                                    nlag = nlag, dir = dir_list[idir,], dx = dx[c(idx_x, idx_y)])
        gg[,idir,isim, ig] = gam$gg
        hh[,idir,isim, ig] = gam$hh
        } # idir
      } # jvar
    } # ivar
  } # isim

  # statistics
  rownames(tab) <- var_name
  print(knitr::kable(tab, digits = 4,
               caption = paste0("Mono-variate statistics for the ", my_cov$type(), " covariance")))
  cat("\n")

  # plot of the variograms
  gg_moy = apply(X = gg, MARGIN = c(1,2,4), FUN = mean)
  hh_moy = apply(X = hh, MARGIN = c(1,2,4), FUN = mean)
  dd = seq(from = 0, to = max(hh_moy), length.out = 100)

  ig = 0
  for (ivar in 1:nvar) {
    for(jvar in ivar:nvar) {
      sub = sprintf("%s (ivar=%d, jvar=%d) in section %s", my_cov$class(), ivar, jvar, slice)
      my_cov_var = reduce_model(my_cov, ivar = ivar, jvar = jvar, verbose = FALSE)
      ig = ig+1
      if (flag.plot) {
        plot(NULL, NULL, xlim = range(dd), ylim = 1.1*c(0,max(gg)),
         xlab = "Distance", ylab = "Variogram",
         main = my_cov_var$description(),
         sub = sub, col.sub = "red")
        abline(h = c(0,1), col = "gray", lty = 3)
        abline(v = c(0), col = "gray", lty = 3)
      }
      for (idir in 1:nrow(dir_list)) {
        # direction
        u0 = rep(0.0, my_cov$getNDim())
        v0 = dir_list[idir,]/sqrt(sum(dir_list[idir,]^2))
        u0[idx_x] = v0[1]
        u0[idx_y] = v0[2]
        # experimental variogram
        hh_experimental = hh_moy[,idir,ig]
        gg_experimental = gg_moy[,idir,ig]
        # model at experimental distance
        if (flag.gstlearn) {
          gg_model =
            gs_cov$sample(h = 0.0, codir = u0, ivar = ivar - 1, jvar = jvar - 1) -
            gs_cov$sample(h = hh_experimental, codir = u0, ivar = ivar - 1, jvar = jvar - 1)
        } else {
          gg_model =
          my_cov$covariance(H = t(rep(0, my_cov$getNDim())), ivar = ivar, jvar = jvar) -
              my_cov$covariance(H = hh_experimental %*% t(u0), ivar = ivar, jvar = jvar)
        }

        rmse_var = sqrt(mean((gg_experimental - gg_model)^2))
        print(sprintf("RMSE on vario %s in dir#%d = %.4f", sub, idir, rmse_var))
        cat("\n")

        if (flag.plot) {
          # experimental variogram
          lines(hh_moy[,idir,ig], gg_moy[,idir,ig], col = col_dir[idir], lty = 2)
          points(hh_moy[,idir,ig], gg_moy[,idir,ig], col = col_dir[idir], pch = 19, cex = 0.2)
          # model
          if (flag.gstlearn) {
          lines(dd, gs_cov$sample(h = 0.0, codir = u0, ivar = ivar - 1, jvar = jvar - 1) -
                  gs_cov$sample(h = dd, codir = u0, ivar = ivar - 1, jvar = jvar - 1),
                  col = col_dir[idir], lty = 1)
          } else {
          lines(dd, my_cov$covariance(H = t(rep(0, my_cov$getNDim())), ivar = ivar, jvar = jvar) -
                my_cov$covariance(H = dd %*% t(u0), ivar = ivar, jvar = jvar),
                  col = col_dir[idir], lty = 1)
          }
        }
      } # idir
      if (flag.plot) {
        legend("bottomright",
               legend = c(lab_inc, "vario.", "model"),
               col = c(col_dir, "gray", "gray"),
               lty = c(rep(1, ndir), 2, 1))
      }
    } #jvar
  } #ivar
  # ----------------------------------------------------------------------------
  # END of test
  # ----------------------------------------------------------------------------
  print("-------------------------------------------")
  print(sprintf(" End of test #%d",  id))
  cat("\n")
  print("-------------------------------------------")

} # loop over the tests

```

# References

* Allard, D., Benoit, L., and Obakrim, S. (2025). Modeling and simulating spatio-temporal, multivariate and nonstationary Gaussian Random Fields: a Gaussian mixtures perspective.
