---
title: "Tutorial PPM"
author: "D. Renard, N. Desassis, X. Freulon"
date: "December 6th 2022"
output:
  pdf_document: default
  html_document: default
  editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE, echo=TRUE}
knitr::opts_chunk$set(echo = TRUE)
rm(list=ls())
library(gstlearn)
library(ggplot2)
library(ggpubr)
library(Matrix)
```

# General Introduction

This notebook presents sevral possibilities of the SPDE procedure

## Creating the Data Bases

Some global constants

```{r}
ndat = 1000
rangev = 0.2
sill = 1.
nugget = 0.1
law_set_random_seed(123)
```

Creating the Data Base for conditioning information

```{r}
dat = Db_create()
dat["x"] = VectorHelper_simulateUniform(ndat)
dat["y"] = VectorHelper_simulateUniform(ndat)
dat$setLocators(c("x","y"),ELoc_X())
dat
```

Creating the Data Base for the output grid

```{r}
grid = DbGrid_create(nx=c(50,50),dx=c(0.02,0.02))
```

Creating the Meshing (Turbo) based on extended grid

```{r}
gridExt = DbGrid_create(nx=c(75,75),dx=c(0.02,0.02),x0=c(-0.25,-0.25))
mesh = MeshETurbo(gridExt)
mesh
```

Creating the Model

```{r}
model = Model_createFromParam(type=ECov_BESSEL_K(),param=1,range=rangev,sill=sill)
model$addCovFromParam(ECov_NUGGET(),sill=nugget)
model
```

# SPDE processing

## Non-conditional simulation

Building the SPDE environment for Non-conditional environments

```{r}
spdeS = SPDE()
spdeS$init(model=model, field=grid, calc=ESPDECalcMode_SIMUNONCOND(), mesh=mesh)
spdeS$compute()
```

Apply the SPDE system in order to obtain a non-conditional simulation on the input and output Data Bases

```{r}
iuid = spdeS$query(dat)
iuid = spdeS$query(grid)
```

Representing the resulting non-conditional simulation on the Grid

```{r}
p = plot.grid(grid, "spde.simu")
p = plot.decoration(p, title="Non Conditional Simulation")
p
```

## Estimation by Kriging

Preparing the SPDE environment

```{r}
spdeK = SPDE()
spdeK$init(model=model, field=grid, data=dat, calc=ESPDECalcMode_KRIGING(), mesh=mesh)
spdeK$compute()
```

Apply the SPDE system in order to obtain an estimation on the output Data Base

```{r}
iuid = spdeK$query(grid, namconv=NamingConvention("spde",FALSE))
```

Representing the resulting estimation on the Grid

```{r}
p = plot.grid(grid, "spde.kriging")
p = plot.decoration(p, title="Estimation")
p
```

## Producing the internal elements of the SPDE environment

Extracting the internal elements from 'spdeK' in order to perform calculations by hand (if necessary).

```{r}
Qtr = csToTriplet(PrecisionOpCs(mesh, model)$getQ(), flag_from_1=TRUE)
Atr = csToTriplet(ProjMatrix(dat,mesh)$getAproj(), flag_from_1=TRUE)

Q = sparseMatrix(i=Qtr$rows,j=Qtr$cols,x=Qtr$values)
Aproj = sparseMatrix(i=Atr$rows, j=Atr$cols, x=Atr$values,
                     dims=c(Atr$nrows,Atr$ncols))
```

Posterior calculations

```{r}
size = dim(Q)[1]
cholQ = Cholesky(Q)
u = VectorHelper_simulateGaussian(size)
```

```{r}
Qop = Q + 1/nugget * t(Aproj) %*% Aproj
```

