---
  title: "Test of the centering"
  author: "D. Renard, N. Desassis, X. Freulon"
  date: "March 8, 2023"
  output:
    pdf_document: default
    html_document: default
    editor_options: 
    chunk_output_type: console
---
<!-- SUMMARY: Performs centering of a Point data set on a regular grid (possibly rotated)  -->

<!-- CATEGORY: Data_Base_management -->

# Initialisation

```{r setup, include=FALSE, echo=TRUE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(fig.align="center")
rm(list=ls())
library(gstlearn)
library(ggplot2)
library(ggpubr)
library(ggnewscale)
library(parallel)
set.seed(1235)
```

## Function implementing the centering of the data set

```{r Internal_functions, echo=FALSE, eval=TRUE}
# Centering the data on the target grid
manual_centering <- function(data, target) {
  err = 0.0
  if (target$isGrid()) {
    dx = target$getDXs()
    x0 = target$getX0s()
    nx = target$getNXs()
    theta = target$getAngles()
    
    xy_nm = data$getNamesByLocator(ELoc_X())
    # initial coordinates
    xy = matrix(NaN, nrow = data$getSampleNumber(), ncol = length(xy_nm))
    for (i in seq_along(xy_nm)) {
      xy[, i] = data$getColumns(xy_nm[i])
    }
    # local coordinates
    uv = XY_to_UV(target, xy)
    
    # centering
    for (i in seq_along(xy_nm)) {
      uv[,i] = dx[i] * floor((uv[,i] + dx[i]/2)/dx[i])
    }
    
    # global coordinates
    xy_disc = UV_to_XY(target, uv)
  
    # storing
    for (i in seq_along(xy_nm)) {
      data$setColumn(tab = xy_disc[,i], name = paste0(xy_nm[i], "_disc"))
    }
    data$setLocators(paste0(xy_nm, "_disc"), ELoc_X())
  
    } else {err = 1}
  err  
}

# Some auxiliary functions for the rotation
translate <- function (X, dx) {
  stopifnot(dim(X)[2] == length(dx))
  X_0 = outer(X = rep(1, dim(X)[1]), Y = dx, FUN = '*')  
  X + X_0
}

UV_to_XY <- function(target, UV) {
  stopifnot(target$getNDim() == dim(UV)[2])
  rot = Rotation(); rot$setAngles(target$getAngles())
  r   = t(matrix(rot$getMatrixDirectVec(), nrow = target$getNDim(), ncol = target$getNDim()))
  translate(UV %*% r, dx = target$getX0s())
}

XY_to_UV <- function(target, XY) {
  stopifnot(target$getNDim() == dim(XY)[2])
  rot = Rotation(); rot$setAngles(target$getAngles())
  r   = (matrix(rot$getMatrixDirectVec(), 
                nrow = target$getNDim(), ncol = target$getNDim()))
  translate(XY, dx = -target$getX0s()) %*% r 
}

bounding_uv <- function(target, delta = 0) {
  dx = target$getDXs()
  nx = target$getNXs()
  duv = dx * (nx - 1)
  matrix(c(
    0 - delta, 0 - delta, 
    duv[1] + delta, 0 - delta, 
    duv[1] + delta, duv[2] + delta, 
    0 - delta, duv[2] + delta,
    0 - delta, 0 - delta
    ), nrow = 5, ncol = 2, byrow = TRUE)
}

bounding_xy <- function(target, delta = 0) {
  bb_uv = bounding_uv(target, delta)
  rot = Rotation()
  rot$setAngles(target$getAngles())
  bb_uv = bb_uv %*% t(matrix(rot$getMatrixDirectVec(), nrow = 2, ncol = 2))
  translate(bb_uv, target$getX0s())
}
```

# Simulation of the data sets

Two data sets are simulated:
- the first data set is constituted os samples randomly located within a dilated bounding box around a rotated grid
- the second data set is a sampling of the nodes of the rotated grid nodes.

The first one is used to check the validity of the centering while the second one prooves that centering does not do anything when the data are alreay centered.

```{r Dataset_construction, echo=TRUE, eval=TRUE}
nech = 25
dx = c(0.05, 0.025)
nx = c(11, 23)
x0 = c(0.5, 2)
angles = c(20, 0)
delta = 0.1 # margin around the grid bounding box

# Creating the rotated grid
target = DbGrid_create(nx = nx, dx = dx, x0 = x0, angles = angles)

# In real coordinates
bb = bounding_xy(target = target, delta = delta) 
XY = as.matrix(cbind(target["x1"], target["x2"]))

# Test of the identity
UV = XY_to_UV(target, XY)
xy = UV_to_XY(target, UV)

# Simulation of uniform scattered points in the local coordinates 
# inside the bounding box
bb_uv   = bounding_uv(target, delta)
bb_xy   = bounding_xy(target, delta)
lim_uv  = bb_uv[c(1,3),]
sim_uv  = lim_uv[1,] + 
  matrix(runif(nech * 2), nrow=nech, ncol=2) %*% diag(lim_uv[2,] - lim_uv[1,])
sim_xy  = UV_to_XY(target, sim_uv)

data_ini1 = Db_createFromSamples(nech)
data_ini1["x1"] = sim_xy[,1]
data_ini1["x2"] = sim_xy[,2]
data_ini1$setLocators(c("x1", "x2"), ELoc_X())
data_ini1["Y"] = rnorm(nech)
data_ini1$setLocators("Y", ELoc_Z())

# Second input data set (extracted from the target grid)
data_ini2 = Db_createSamplingDb(target, number=nech, names=c("x1","x2"))
data_ini2["Y"] = rnorm(data_ini2$getSampleNumber())
data_ini2$setLocators("Y", ELoc_Z())
```

Display of the first data set

```{r Plotting_data1_initial}
ggDefaultGeographic() + 
  plot.XY(bb[,1], bb[,2], flagLine=TRUE, color='blue') + 
  plot.point(target) +
  plot.point(data_ini1, color = "grey", size=4) +
  plot.decoration(title="First Data Set")
```

Display of the second data set

```{r Plotting_data2_initial}
ggDefaultGeographic() + 
  plot.point(target) +
  plot.point(data_ini2, col = "grey", size=4) + 
  plot.decoration(title = "Second Data Set")
```

# Test of the centering

Two centering are systematically tested:

- the first one uses the manual procedure
- the second one uses the procedure which is available in gstlearn.

The graphic convention is the following one:
- the nodes of the target grid are always represented (black dots)
- the initial data set are displayed with large size grey symbols
- the centered data are displayed with small red circles.
- the bounding box is represented as a solid blue line (when relevant)

## Scattered data

Performing the manual centering

```{r Manual_centering_data1, echo=TRUE, eval=TRUE}
data_man1 = data_ini1$clone()
err = manual_centering(data_man1, target)
```

Displaying the centered results

```{r Plotting_manual_centering_data1}
ggDefaultGeographic() + 
  plot.XY(bb[,1], bb[,2], flagLine=TRUE, color='blue') + 
  plot.point(target) +
  plot.point(data_ini1, col = "grey", size=4) +
  plot.point(data_man1, col='red', size=2) + 
  plot.decoration(title="First Data Set (centered manually)")
```

Performing the centering using gstlearn: this is done in place (coordinates have been modified).

```{r Gstlearn_centering_data1}
data_auto1 = data_ini1$clone()
err = DbHelper_centerPointToGrid(data_auto1, target)
```

Displaying the centered results

```{r Plotting_gstlearn_centering_data1}
ggDefaultGeographic() + 
  plot.XY(bb[,1], bb[,2], flagLine=TRUE, color='blue') + 
  plot.point(target) +
  plot.point(data_ini1, col = "grey", size=4) +
  plot.point(data_auto1, col='red', size=2) + 
  plot.decoration(title="First Data Set (centered by gstlearn)")
```

## Data extracted from the target grid

Performing the manual centering 

```{r Manual_centering_data2}
data_man2 = data_ini2$clone()
err = manual_centering(data_man2, target)
```

Displaying the centered results

```{r Plotting_manual_centering_data2}
ggDefaultGeographic() + 
  plot.point(target) +
  plot.point(data_ini2, col = "grey", size=4) +
  plot.point(data_man2, col='red', size=2) + 
  plot.decoration(title="Second Data Set (centered manually)")
```

Performing the centering using gstlearn: this is done in place (coordinates have been modified).

```{r Gstlearn_centering_data2}
data_auto2 = data_ini2$clone()
err = DbHelper_centerPointToGrid(data_auto2, target)
```

Displaying the centered results

```{r Plotting_gstlearn_centering_data2}
ggDefaultGeographic() + 
  plot.point(target) +
  plot.point(data_ini2, col = "grey", size=4) +
  plot.point(data_auto2, col='red', size=2) + 
  plot.decoration(title="Second Data Set (centered by gstlearn)")
```

