---
title: "Spectral simulation of random functions on $R^2$"
author: "gstlearn Core Team"
date: "2024-06-11"
output: pdf_document
---

# Initialization

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
rm(list = ls())
library(gstlearn)
library(ggplot2)
library(ggpubr)

# global parameters
flag.verbose  = FALSE
opers = EStatOption_fromKeys(c("NUM", "MINI", "MAXI", "MEAN", "STDV"))

```

### Somme auxiliary functions

```{r fn_stat, echo=FALSE, eval=TRUE}
matToGst <- function(mat) {
  nr = dim(mat)[1]
  nc = dim(mat)[2]
  stopifnot(is.matrix(mat))
  res = MatrixDense(nr, nc)
  for (i in 1:nc) {
      res$setColumn(icol = i-1, mat[,i])
    }
  res
}
covToGst <- function(mat) {
  nr = dim(mat)[1]
  nc = dim(mat)[2]
  stopifnot(is.matrix(mat)&(nr == nc))
  res = MatrixSymmetric(nr)
  for (i in 1:nr) {
    for (j in i:nr) {
      res$setValue(icol = i-1, irow = j-1, mat[i,j])
    }}
  res
}

create_random_covmat = function(nvar) {
  uu         = rnorm(nvar)
  s2         = rnorm(nvar)^2
  outer(
    X = 1:nvar,
    Y = 1:nvar,
    FUN = function(i,j){sqrt(s2[i]*s2[j])*exp(-abs(uu[i]-uu[j]))})
}

#' -----------------------------------------------------------------------------
#' @param val a vector with the value of the first variable
#' @param val2 a vectro with the value of the second variable (optional for cross variagram)
#' @param nlag a integer giving the the number of lags
#' @param dx   a value defining the grid mesh
#' @value return a list defining the experimental variogram (nc, hh, gg)
compute_vario_on_1d_grid<-function(val, val2 = NULL, nlag = 50, dx = 1.0) {
  nx = length(val)
  stopifnot(is.null(val2)|(length(val2) == nx))
  stopifnot((nlag > 0)&(nlag < nx))
  stopifnot(dx > 0)
  gg = rep(NaN, nlag)
  hh = rep(NaN, nlag)
  nc = rep(NaN, nlag)
  lag_norm = dx
  for (ilag in 1:nlag) {
      idx_deb = 1:(nx-ilag)
      idx_end = idx_deb + ilag
      if (is.null(val2)) {
        tmp = 1/2 * as.numeric((val[idx_deb] - val[idx_deb + ilag])^2)
      } else {
        tmp = 1/2 * as.numeric((val[idx_deb] - val[idx_deb + ilag])*
                               (val2[idx_deb] - val2[idx_deb + ilag]))
      }
      nc[ilag] = length(tmp)
      if (nc[ilag] > 0) {
          hh[ilag] = lag_norm * ilag
          gg[ilag] = sum(tmp)/nc[ilag]
      }
    }
  list(
    nc = nc,
    hh = hh,
    gg = gg
  )
}

#' -----------------------------------------------------------------------------
#' @param val a matrix with the value of the first variable
#' @param val2 a matrix with the value of the second variable (optional for cross variagram)
#' @param nlag a integer giving the the number of lags
#' @param lag  a vector of grid increments defining the lag
#' @param dx   a vector defining the grid mesh
#' @value return a list defining the experimental variogram (nc, hh, gg)
compute_vario_on_2d_grid <- function(val, val2 = NULL, nlag = 10, lag = c(1,0), dx = c(1,1)) {
  stopifnot(is.matrix(val))
  stopifnot(is.null(val2)|(is.matrix(val2)&(all(dim(val) == dim(val2)))))
  stopifnot((nlag > 0)&(nlag*max(abs(lag)) < min(dim(val))))
  stopifnot((length(lag) == 2)&(length(dx) == 2))
  nx = dim(val)
  gg = rep(NaN, nlag)
  hh = rep(NaN, nlag)
  nc = rep(NaN, nlag)
  lag_norm = sqrt(sum((lag*dx)^2))
  for (ilag in 1:nlag) {
      idx_deb = max(1,1-ilag*lag[1]):min(nx[1], nx[1]-ilag*lag[1])
      jdx_deb = max(1,1-ilag*lag[2]):min(nx[2], nx[2]-ilag*lag[2])
      idx_end = idx_deb + ilag*lag[1]
      jdx_end = jdx_deb + ilag*lag[2]
      if (is.null(val2)) {
        tmp = 1/2 * as.numeric((val[idx_deb, jdx_deb] - val[idx_end, jdx_end])^2)
      } else {
        tmp = 1/2 * as.numeric((val[idx_deb, jdx_deb] - val[idx_end, jdx_end])*
                               (val2[idx_deb, jdx_deb] - val2[idx_end, jdx_end])
                              )
      }
      nc[ilag] = length(tmp)
      if (nc[ilag] > 0) {
          hh[ilag] = lag_norm * ilag
          gg[ilag] = sum(tmp)/nc[ilag]
      }
    }
  list(
    nc = nc,
    hh = hh,
    gg = gg
  )
}

```


```{r fn_draw_model, echo=FALSE, eval=TRUE}
draw_model <- function(model, prefix, flag.st = FALSE) {
  ndim = model$getNDim()
  stopifnot(ndim %in% 1:3)
  nvar = model$getNVar()
  C0 = matrix(NaN, nvar, nvar)
  u0 = c(rep(0, ndim-1), 1.0)
  for (ivar in 1:nvar)  {
    for (jvar in 1:nvar) {
      C0[ivar, jvar] = model$sample(h = 0.0, codir = u0, ivar-1, jvar-1)
    }
  }
  angles_in_deg = c(0, 45, 90, 135)
  angles_col = c("red", "orange", "darkgreen", "skyblue")
  var_col = palette.colors(n = nvar*(nvar+1)/2, palette = palette.pals()[1])
  h = seq(from = 0, to = 10, length.out = 100)
  if (!flag.st) {
  stopifnot(ndim %in% c(1,2))
  ndir = c(1,4)[ndim]
  for (ivar in 1:nvar) {
    for (jvar in ivar:nvar) {
      titre = sprintf("%s[%d, %d]", prefix, ivar, jvar)
      plot(NULL, NULL, xlim = range(h), ylim = c(0, C0[ivar, jvar]),
      xlab = "Distances", ylab = "Variogram", main = titre)
      abline(h = c(0, C0[ivar, jvar]), col = "gray", lty = 3)
      abline(v = c(0), col = "gray", lty = 3)
      leg = NULL
      for (idir in 1:ndir) {
        if (ndim == 1) {
          u0 = 1.0
          leg = c(leg, "model")
        } else {
          theta = angles_in_deg[idir] * pi / 180
          u0    = c(cos(theta), sin(theta))
          leg = c(leg, sprintf("theta = %.0f°", angles_in_deg[idir]))
        }
        lines(h, C0[ivar, jvar] - model$sample(h = h, codir = u0, ivar-1, jvar-1),
                , col = angles_col[idir], lty = 1)
      } # loop over the directions
      legend("bottomright",
             legend = leg, col = angles_col[1:ndir], lty = 1, lw = 1)
  }} # loops over the variables

  } else {
    # space-time model
    ndir = 1 + c(1,4)[ndim-1] + 4
    for (idir in 1:ndir) {
      if (idir == 1) {
        # time trace
        lab_dir = "time trace"
        lab_x   = "Time distances"
        u0      = c(rep(0.0, ndim - 1), 1.0)
      } else if (idir %in% 2:(1+c(1,4)[ndim-1])) {
        # space trace
        kdir    = idir - 1
        lab_dir = sprintf("space trace in dir%d#", kdir)
        lab_x   = "Distances"
        if (ndim == 2) {
          u0      = c(1.0, 0.0)
        }
        if (ndim == 3) {
          theta = angles_in_deg[kdir] * pi / 180
          u0    = c(cos(theta), sin(theta), 0.0)
        }
      } else {
        # random direction
        u0 = rnorm(ndim)
        u0 = u0 /sqrt(sum(u0^2))
        if (ndim == 2){
          lab_dir = sprintf("u0 = [%.2f, %.2f]", u0[1], u0[2])
        }
        if (ndim == 3) {
          lab_dir = sprintf("u0 = [%.2f, %.2f, %.2f]", u0[1], u0[2], u0[3])
        }
        lab_x   = "Space-time distances"
      }
      titre = sprintf("%s[R%d]: %s", prefix, ndim-1, lab_dir)
      # the graphic
      plot(NULL, NULL, xlim = range(h), ylim = 1.1*c(0, max(as.numeric(C0))),
      xlab = lab_x, ylab = "Variogram", main = titre)
      abline(h = c(0, as.numeric(C0)), col = "gray", lty = 3)
      abline(v = c(0), col = "gray", lty = 3)
      var_lab = NULL
      idx = 0
      for (ivar in 1:nvar) {
      for (jvar in ivar:nvar) {
        idx = idx + 1
        var_lab = c(var_lab, sprintf("{v%d, v%d}", ivar, jvar))
        lines(h, C0[ivar, jvar] - model$sample(h = h, codir = u0, ivar-1, jvar-1),
          , col = var_col[idx], lty = 1)
      }}
      legend ("bottomright", legend = var_lab, col = var_col, lty = 1)
    } # loop on the space-time directions
  } # space-time model
}
```


```{r fn_simu_model, echo=FALSE, eval=TRUE}
# ------------------------------------------------------------------------------
simu_model <- function(model, prefix, nsimu = 1, ns = 1000, flag.st = TRUE,
                       flag.QC = TRUE,
                       flag.histo = TRUE, flag.map = TRUE, flag.vario = TRUE,
                       verbose = TRUE) {
  ndim = model$getNDim()
  nvar = model$getNVar()
  opers = EStatOption_fromKeys(c("NUM", "MINI", "MAXI", "MEAN", "STDV"))
  # defining the name of the simulated variables
  simu_nm = NULL
  for (is in 1:nsimu) {
  for (ivar in 1:nvar) {
    simu_nm = c(simu_nm, sprintf("%s.V%d.S%d", prefix, ivar, is))
    }
  }
  # defining the sections
  sec_list = list()
  # Spatial covariance on R^1 or R^2
  if (!flag.st == TRUE) {
    stopifnot(ndim %in% 1:2)
    if (ndim == 1) {
      sec_list[[1+length(sec_list)]] <- list(
        grid = DbGrid_create(nx = 1000, dx = 0.5),
        section = "Ot",
        coordinates = c("t")
      )
    }
    if (ndim == 2) {
      sec_list[[1+length(sec_list)]] <- list(
        grid = DbGrid_create(nx = c(100, 100), dx = c(0.5, 0.5)),
        section = "xOy",
        coordinates = c("x", "y")
      )
    }
    } else {
    # Space-time covariance on R^1xTime or R^2xTime
    stopifnot(ndim %in% 2:3)
    if (ndim == 2) {
      sec_list[[1+length(sec_list)]] <- list(
        grid = DbGrid_create(nx = c(100, 100), dx = c(0.5, 0.5)),
        section = "xOt",
        coordinates = c("x", "t")
      )
    } else {
      for (isec in 1:3) {
        nx = c(100, 100, 100)
        nx[isec] = 1
      sec_list[[1+length(sec_list)]] <- list(
        grid = DbGrid_create(nx = nx, dx = c(0.5, 0.5, 0.5)),
        section = c("yOt", "xOt", "xOy")[isec],
        coordinates = c("x", "y", "t")[-isec]
      )
      }
    }
  }
  nsec = length(sec_list)

  # --------------
  # simulations
  # --------------
  for (is in 1:nsimu) {
    if (verbose) {print(sprintf(">>> simulation %d/%d", is, nsimu))}
    sp  = model$simulateOnRN(ns = ns)
    # computing the simulations on 1D or 2D sections
    for (sec in 1:nsec) {
      if (verbose) {
        print(sprintf(" - computing values on section %s", sec_list[[sec]]$section))
        }
      val = sp$computeToMatrix(sec_list[[sec]]$grid)$toTL()
      for (ivar in 1:nvar) {
        sec_list[[sec]]$grid[simu_nm[ivar+(is-1)*nvar]] = val[,ivar]
        } # loop over the variables
      } # loop over the sections
    } # loop over the simulations

  # --------------
  # QC
  # --------------
  if (flag.QC) {
  res = list()
    for (sec in 1:nsec) {
        if (verbose) {
          print(sprintf(">>> computing statistics on section %s", sec_list[[sec]]$section))
          }
      res[[1+length(res)]] <- QC_section(
        model = model,
        grid  = sec_list[[sec]]$grid,
        prefix = prefix,
        nsimu = nsimu, isec = sec,
        flag.histo = flag.histo, flag.map = flag.map, flag.vario = flag.vario,
        verbose = verbose)
    } # loop on sections
  res
  } else {
    sec_list
  }
}

# QC of the simulations on the sections
QC_section <- function(model, grid, prefix, nsimu, isec = 1,
                       flag.histo = TRUE, flag.map = TRUE, flag.vario = TRUE,
                       verbose = TRUE) {
  opers = EStatOption_fromKeys(c("NUM", "MINI", "MAXI", "MEAN", "STDV"))
  ndim  = model$getNDim()
  stopifnot((grid$isGrid())&(ndim == grid$getNDim()))
  nx = grid$getNXs()
  dx = grid$getDXs()
  sec_lab  = c("yOt", "xOt", "xOy")
  sec_coor = c("x", "y", "t")
  nvar  = model$getNVar()
  C0 = matrix(NaN, nvar, nvar)
  for (ivar in 1:nvar) {
    for (jvar in 1:nvar) {
      C0[ivar, jvar] = model$sample(h = 0, codir = rep(0,ndim), ivar - 1, jvar - 1)
    }
  }
  # --------------------------------------------------------------------------
  # histogram and raster for each variable of the first simulation
  # --------------------------------------------------------------------------
  for (ivar in 1:nvar) {
    s_nm  = sprintf("%s.V%d.S%d", prefix, ivar, 1)
    titre = sprintf("%s[%d] in section %s", prefix, ivar, sec_lab[isec])
    # histogram
    if (flag.histo) {
    p = plot.init() +
      plot.hist(db = grid, name = s_nm, bins = 25, col = "skyblue") +
      plot.decoration(title = titre)
    plot.end(p)
    } # flag.histo
    # base map
    if (flag.map) {
    if (ndim > 1) {
      p = plot.init(asp = 1)
      if (ndim == 2) {
        p = p + plot.raster(db = grid, name = s_nm)
      }
      if (ndim == 3) {
        dbg = DbGrid_create(nx = nx[-isec], dx = dx[-isec])
        dbg[s_nm] = grid[s_nm]
        p = p + plot.raster(db = dbg, name = s_nm)
      }
      p = p + plot.decoration(title = titre,
              xlab = sec_coor[-isec][1],
              ylab = sec_coor[-isec][2]
              )
      plot.end(p)
    }
    } # flag.map
  } # loop over the variables

  # --------------------------------------------------------------------------
  # experimental variograms vs. model
  # --------------------------------------------------------------------------
  if (flag.vario) {
  nlag  = 50
  probs = c(0.25, 0.5, 0.75)
  probs_col = c("blue", "orange", "green")
  if (ndim == 1) {
  lag_list = matrix(c(1), nrow = 1, ncol = 1)
  } else {
  lag_list = matrix(c(1, 0, 1, 1, 0, 1, -1, 1), nrow = 2, ncol = 4)
  }
  ndir = dim(lag_list)[2]
  # --------------------------------------------------------------------------
  for (ivar in 1:nvar) {
  for (jvar in ivar:nvar) {
    for (idir in 1:ndir) {
      # computing the variograms...
      gg_simu = array(NaN, dim = c(nsimu, nlag))
      hh_simu = array(NaN, dim = c(nlag))
      for (is in 1:nsimu) {
        v1_nm = sprintf("%s.V%d.S%d", prefix, ivar, is)
        v2_nm = sprintf("%s.V%d.S%d", prefix, jvar, is)
        v1    = grid$getColumns(names = v1_nm)
        v2    = grid$getColumns(names = v2_nm)
        if (length(dx) == 1) {
          varexp = compute_vario_on_1d_grid(
            val  = v1,
            val2 = v2,
            nlag = nlag, dx = dx)
        } else if (length(dx) == 2) {
          varexp = compute_vario_on_2d_grid(
              val  = matrix(v1, nrow = nx[1], ncol = nx[2]),
              val2 = matrix(v2, nrow = nx[1], ncol = nx[2]),
              nlag = nlag, lag = lag_list[,idir], dx = dx)
        } else if (length(dx) == 3) {
          varexp = compute_vario_on_2d_grid(
              val  = matrix(v1, nrow = nx[-isec][1], ncol = nx[-isec][2]),
              val2 = matrix(v2, nrow = nx[-isec][1], ncol = nx[-isec][2]),
              nlag = nlag, lag = lag_list[,idir], dx = dx[-isec])
        }
        gg_simu[is, ] = varexp$gg
        if (is == 1) {hh_simu = varexp$hh}
      } # loop over the simulations

      # plot for the ivar, jvar, idir
      titre = sprintf("%s[%d, %d]: idir = %d in section %s",
                      prefix, ivar, jvar, idir, sec_lab[isec])
      plot(NULL, NULL, xlim = c(0, max(hh_simu)), ylim = 1.1*c(0, C0[ivar, jvar]),
         xlab = "Distance", ylab = "Variogram", main = titre)
      abline(h = c(0, C0[ivar, jvar]), col = "gray", lw = 1, lty = 3)
      abline(v = c(0), col = "gray", lw = 1, lty = 3)
      # experimental variogram
      for (is in 1:nsimu) {
        lines(hh_simu, gg_simu[is, ], col = "gray", lw = 1, lty = 2)
      }
      for (n in seq_along(probs)) {
        lines(hh_simu,
              apply(X = gg_simu, MARGIN = 2,
              FUN = quantile, probs = probs[n], na.rm = TRUE),
          col = probs_col[n], lw = 1, lty = 2)
        }
      lines(hh_simu, apply(X = gg_simu, MARGIN = 2,
        FUN = mean, na.rm = TRUE),
        col = "red", lw = 1, lty = 1 )
      # model
      if (ndim == 1) {u0 = 1.0}
      if (ndim == 2) {u0 = lag_list[,idir]*dx}
      if (ndim == 3) {
        u0 = rep(0.0, ndim)
        u0[-isec] = lag_list[,idir]
        u0 = u0 * dx
        }
      u0 = u0/sqrt(sum(u0^2))
      h = seq(from = 0.0, to = max(hh_simu), length.out = 200)
      lines(h, C0[ivar, jvar] - mod$sample(h = h, codir = u0, ivar-1, jvar-1),
              col = "black", lw = 2, lty = 1)
      # legend
      legend("bottomright",
           legend = c("mean", paste0("Q", round(100*probs,0)), "empirical", "model"),
           col = c("red", probs_col, "gray", "black"),
           lw = c(1, rep(1, length(probs_col)), 1, 2),
           lty = c(1, rep(2, length(probs_col)), 1, 1)
           )
  }}} # loops on the variables and directions
  } # flag.vario
  # --------------------------------------------------------------------------
  # statistics and results
  # --------------------------------------------------------------------------
  list(prefix  = prefix, section = sec_lab[isec],
       grid = grid,
       tab = dbStatisticsMono(db = grid, names = paste(prefix, "*", sep = "."),
                           opers = opers)$toTL()
  )
}

```

```{r simu_parameters, eval=TRUE}
# simulation parameters
nb = 5000
nsimu = 20
```

# 1d models

Simulation of the the generalized Cauchy covariance on $\mathbb{R}^1$ using *CorAniso*:
$$
C(h; \alpha, \beta, a_t) = (1 + \|\frac{h}{a_t}\|^\alpha)^{-\beta}
$$
with

* $\alpha \in (0, 2]$
* $\beta > 0$
* $a_t > 0$

```{r 1d_model, eval=TRUE}
ctxt = CovContext(nvar = 1, space = SpaceRN_create(1))
mod  = CorAniso_create(ctxt = ctxt, type = ECov_fromKey("CAUCHY_GEN"),
                      params = c(2.0, 1.0),
                      ranges = 2.0,
                      flagRange = FALSE)

# QC of the model
err = draw_model(model = mod, prefix = "CAUCHY_GEN")
res = simu_model(model = mod, prefix = "CAUCHY_GEN",
                 nsimu = nsimu, ns = nb,
                 flag.st = FALSE, verbose = TRUE)

# statistics
knitr::kable(res[[1]]$tab, digits = 3,
               caption = sprintf("Statistics for %s in section %s",
                           res[[1]]$prefix, res[[1]]$section))

```

# Univariate models in R^2

```{r 2d_simulation, eval=TRUE}
ctxt = CovContext(nvar = 1, space = SpaceRN_create(2))
mod  = CorAniso_create(ctxt = ctxt, type = ECov_fromKey("MATERN"),
                      params = c(3/4),
                      ranges = c(1,2), angles = c(30,0),
                      flagRange = FALSE)

# QC of the model
err = draw_model(model = mod, prefix = "MATERN")
res = simu_model(model = mod, prefix = "MATERN",
                 nsimu = nsimu, ns = nb,
                 flag.st = FALSE, verbose = TRUE)

# statistics
knitr::kable(res[[1]]$tab, digits = 3,
               caption = sprintf("Statistics for %s in section %s",
                           res[[1]]$prefix, res[[1]]$section))

```

# Multivariate models in R^2

```{r 2d_multi_simulation, eval=TRUE}
type = "MATERN"
ctxt = CovContext(nvar = 2, space = SpaceRN_create(2))
mod  = CorGaussianMixture_create(ctxt = ctxt, type = ECov_fromKey(type),
                      params = c(3/4, 1.0), kappas =  c(1,1),
                      ranges = c(2,4), angles = c(30,0),
                      flagRange = FALSE)

# QC of the model
err = draw_model(model = mod, prefix = type)
res = simu_model(model = mod, prefix = type, nsimu = nsimu, ns = nb)
lab = sprintf("Statistics for %s in section %s", res[[1]]$prefix, res[[1]]$section)
knitr::kable(res[[1]]$tab, digits = 3, caption = lab)

```

# Factorized model in 2d

```{r 2d_factorized, eval=TRUE}
type = "Fac2d_exp_gauss"
ctxt1d = CovContext(nvar = 1, space = SpaceRN_create(1))
ctxt2d = CovContext(nvar = 1, space = SpaceRN_create(2))

mod  = CorFactorized_create(ctxt = ctxt2d)
# first factor: exponential covariance along Ox
err = mod$addFactor(
  CorAniso_create(ctxt = ctxt1d, type = ECov_fromKey("EXPONENTIAL"),
                          params = NULL, ranges = 5.0, flagRange = FALSE),
        proj = matToGst(matrix(c(1,0), 2, 1))
)
# Second factor: gaussian covariance along Oy
err = mod$addFactor(
  CorAniso_create(ctxt = ctxt1d, type = ECov_fromKey("GAUSSIAN"),
                          params = NULL, ranges = 5.0, flagRange = FALSE),
        proj = matToGst(matrix(c(0,1), 2, 1))
)
# QC of the model
err = draw_model(model = mod, prefix = type)
res = simu_model(model = mod, prefix = type, nsimu = nsimu, ns = nb)
lab = sprintf("Statistics for %s in section %s", res[[1]]$prefix, res[[1]]$section)
knitr::kable(res[[1]]$tab, digits = 3, caption = lab)

```

# Intrinsic correlation in 2d

```{r 2d_intrisic, eval=TRUE}
type = "CorrIntrisic_exp"
ctxt2d = CovContext(nvar = 1, space = SpaceRN_create(2))
sigma = covToGst(matrix(c(1, sqrt(1*3)*0.7, sqrt(1*3)*0.7, 3), 2, 2))
mod  = CovBase(
  cor = CorAniso_create(ctxt = ctxt2d, type = ECov_fromKey("EXPONENTIAL"),
          params = NULL, ranges = c(2.0, 4), angles = c(30,0), flagRange = FALSE),
  sills = sigma)

# QC of the model
err = draw_model(model = mod, prefix = type)
res = simu_model(model = mod, prefix = type, nsimu = nsimu, ns = nb)
lab = sprintf("Statistics for %s in section %s", res[[1]]$prefix, res[[1]]$section)
knitr::kable(res[[1]]$tab, digits = 3, caption = lab)

```

# A linear combination of 2d factors

```{r 2d_lc_2d_factors, eval=TRUE}
ctxt2d = CovContext(nvar = 1, space = SpaceRN_create(2))

mod  = CovList(ctxt = ctxt2d)
# first factor: exponential covariance
fac1 = CorAniso_create(ctxt = ctxt2d, type = ECov_fromKey("EXPONENTIAL"),
                          params = NULL, ranges = c(5.0, 5.0), flagRange = FALSE)
err = mod$addCov(cov = CovBase(cor = fac1, sills = covToGst(as.matrix(1/2))))

# Second factor: gaussian covariance along Oy
fac2 = CorAniso_create(ctxt = ctxt2d, type = ECov_fromKey("GAUSSIAN"),
                          params = NULL, ranges = c(2.0, 5), flagRange = FALSE)
err = mod$addCov(cov = CovBase(cor = fac2, sills = covToGst(as.matrix(2.0))))

# QC of the model
err = draw_model(model = mod, prefix = type)
res = simu_model(model = mod, prefix = type, nsimu = nsimu, ns = nb)
lab = sprintf("Statistics for %s in section %s", res[[1]]$prefix, res[[1]]$section)
knitr::kable(res[[1]]$tab, digits = 3, caption = lab)

```

# A CovBase with a factorized factor

```{r 2d_covbase_1d_factors, eval=TRUE}
ctxt2d = CovContext(nvar = 1, space = SpaceRN_create(2))
ctxt1d = CovContext(nvar = 1, space = SpaceRN_create(1))

# first factor: 1d exponential covariance along Ox
fac1  = CorFactorized_create(ctxt = ctxt2d)
err   = fac1$addFactor(
  CorAniso_create(ctxt = ctxt1d, type = ECov_fromKey("EXPONENTIAL"),
                          params = NULL, ranges = 2.5, flagRange = FALSE),
        proj = matToGst(matrix(c(1,0), 2, 1))
)
C1  = covToGst(matrix(1.5, 1, 1))

mod = CovBase(fac1, C1)

# QC of the model
type = "CovBase_F1D"
err = draw_model(model = mod, prefix = type)
res = simu_model(model = mod, prefix = type, nsimu = nsimu, ns = nb)
lab = sprintf("Statistics for %s in section %s", res[[1]]$prefix, res[[1]]$section)
knitr::kable(res[[1]]$tab, digits = 3, caption = lab)

```

# A linear combination of 1d factors

```{r 2d_lc_1d_factors, eval=TRUE}
ctxt2d = CovContext(nvar = 1, space = SpaceRN_create(2))
ctxt1d = CovContext(nvar = 1, space = SpaceRN_create(1))

mod  = CovList(ctxt = ctxt2d)

# first factor: 1d exponential covariance along Ox
fac1  = CorFactorized_create(ctxt = ctxt2d)
err   = fac1$addFactor(
  CorAniso_create(ctxt = ctxt1d, type = ECov_fromKey("EXPONENTIAL"),
                          params = NULL, ranges = 2.5, flagRange = FALSE),
        proj = matToGst(matrix(c(1,0), 2, 1))
)
C1  = covToGst(matrix(1.0, 1, 1))
err = mod$addCov(CovBase(fac1, C1))
# err = mod$addCov(fac1)

# second factor: 1d gaussian covariance along Oy
fac2  = CorFactorized_create(ctxt = ctxt2d)
# first factor: exponential covariance along Ox
err = fac2$addFactor(
  CorAniso_create(ctxt = ctxt1d, type = ECov_fromKey("GAUSSIAN"),
                          params = NULL, ranges = 5.0, flagRange = FALSE),
        proj = matToGst(matrix(c(0,1), 2, 1))
)
C2  = covToGst(matrix(2.0, 1, 1))
err = mod$addCov(CovBase(fac2, C2))
# err = mod$addCov(fac2)

# QC of the model
type = "Fac2D_2F_1D"
err = draw_model(model = mod, prefix = type)
res = simu_model(model = mod, prefix = type, nsimu = nsimu, ns = nb)
lab = sprintf("Statistics for %s in section %s", res[[1]]$prefix, res[[1]]$section)
knitr::kable(res[[1]]$tab, digits = 3, caption = lab)

```

# A LMC with two variables and 3 stuctures

```{r 2d_LMC_2v, eval=TRUE}
type = "LMC_2v"
ctxt2d2v = CovContext(nvar = 2, space = SpaceRN_create(2))
ctxt2d1v = CovContext(nvar = 1, space = SpaceRN_create(2))

mod  = CovList(ctxt = ctxt2d2v)
# first structure: exponential covariance, only the first variable
rho1 = CorAniso_create(ctxt = ctxt2d1v, type = ECov_fromKey("EXPONENTIAL"),
                          params = NULL, ranges = c(2.5, 2.5), flagRange = FALSE)
C1   = covToGst(matrix(c(1/2, 0, 0, 0), 2, 2))
err = mod$addCov(cov = CovBase(cor = rho1, sills = C1))

# Second factor: gaussian covariance, intrinsic on the two variables
rho2 = CorAniso_create(ctxt = ctxt2d1v, type = ECov_fromKey("GAUSSIAN"),
                          params = NULL, ranges = c(2.0, 2.0), flagRange = FALSE)
C2   = covToGst(matrix(c(1, 0.7, 0.7, 1), 2, 2))
err = mod$addCov(cov = CovBase(cor = rho2, sills = C2))

# Third factor: bi-matern covariance
rho3  = CorGaussianMixture_create(ctxt = ctxt2d2v, type = ECov_fromKey("CAUCHY"),
                      params = c(3/4, 1.0), kappas =  c(1,2),
                      ranges = c(2,4), angles = c(0,0),
                      flagRange = FALSE)
C3   = covToGst(diag(2))
err = mod$addCov(cov = CovBase(cor = rho3, sills = C3))

# QC of the model
err = draw_model(model = mod, prefix = type)
res = simu_model(model = mod, prefix = type, nsimu = nsimu, ns = nb)
lab = sprintf("Statistics for %s in section %s", res[[1]]$prefix, res[[1]]$section)
knitr::kable(res[[1]]$tab, digits = 3, caption = lab)

```

# Space-time model

## 1d univariate

```{r space-time_1d1v, eval=TRUE}
type = c("GNEITING_G", "GNEITING_M", "GNEITING_C")[2]
ndim = 1
nvar = 1
# parameters for the time trace
alpha = 2.0
beta  = 1.0
# parameters for the space trace
nu   = c(1/2, 3/4, 1)[1:nvar]
rr   = c(1, 2, 1)[1:nvar]
# space-time anisotropy
timeRange = 2.0
ranges = c(5,10)[1:ndim]
angles = c(30, 0)[1:ndim]
sep = 1.0

ctxt = CovContext(nvar = nvar, space = SpaceRN_create(ndim = ndim))
mod = CorGneiting_create(ctxt = ctxt, type = ECov_fromKey(type),
                       alpha = alpha, beta = beta, timeRange = timeRange,
                       params = nu, kappas = rr,
                       ranges = ranges, angles = angles,
                       flagRange = FALSE,
                       separability = sep)

# QC of the model
defineDefaultSpaceRnT(ndim = ndim)
err = draw_model(model = mod, prefix = type, flag.st = TRUE)
res = simu_model(model = mod, prefix = type, nsimu = nsimu, ns = nb, flag.st = TRUE)
lab = sprintf("Statistics for %s in section %s", res[[1]]$prefix, res[[1]]$section)
knitr::kable(res[[1]]$tab, digits = 3, caption = lab)

```

## 1d bivariate

```{r space-time_1d2v, eval=TRUE}
type = c("GNEITING_G", "GNEITING_M", "GNEITING_C")[2]
ndim = 1
nvar = 2
# parameters for the time trace
alpha = 2.0
beta  = 1.0
# parameters for the space trace
nu   = c(1/2, 3/4, 1)[1:nvar]
rr   = c(1, 2, 1)[1:nvar]
# space-time anisotropy
timeRange = 2.0
ranges = c(5,10)[1:ndim]
angles = c(30, 0)[1:ndim]
sep = 1.0

ctxt = CovContext(nvar = nvar, space = SpaceRN_create(ndim = ndim))
mod = CorGneiting_create(ctxt = ctxt, type = ECov_fromKey(type),
                       alpha = alpha, beta = beta, timeRange = timeRange,
                       params = nu, kappas = rr,
                       ranges = ranges, angles = angles,
                       flagRange = FALSE,
                       separability = sep)

# QC of the model
defineDefaultSpaceRnT(ndim = ndim)
err = draw_model(model = mod, prefix = type, flag.st = TRUE)
res = simu_model(model = mod, prefix = type, nsimu = nsimu, ns = nb, flag.st = TRUE)
lab = sprintf("Statistics for %s in section %s", res[[1]]$prefix, res[[1]]$section)
knitr::kable(res[[1]]$tab, digits = 3, caption = lab)

```

## 2d univariate

```{r space-time_2d1v, eval=TRUE}
type = c("GNEITING_G", "GNEITING_M", "GNEITING_C")[2]
ndim = 2
nvar = 1
# parameters for the time trace
alpha = 2.0
beta  = 1.0
# parameters for the space trace
nu   = c(1/2, 3/4, 1)[1:nvar]
rr   = c(1, 2, 1)[1:nvar]
# space-time anisotropy
timeRange = 2.0
ranges = c(5,10)[1:ndim]
angles = c(30, 0)[1:ndim]
sep = 1.0

ctxt = CovContext(nvar = nvar, space = SpaceRN_create(ndim = ndim))
mod = CorGneiting_create(ctxt = ctxt, type = ECov_fromKey(type),
                       alpha = alpha, beta = beta, timeRange = timeRange,
                       params = nu, kappas = rr,
                       ranges = ranges, angles = angles,
                       flagRange = FALSE,
                       separability = sep)

# QC of the model
defineDefaultSpaceRnT(ndim = ndim)
err = draw_model(model = mod, prefix = type, flag.st = TRUE)
res = simu_model(model = mod, prefix = type, nsimu = nsimu, ns = nb, flag.st = TRUE)
lab = sprintf("Statistics for %s in section %s", res[[1]]$prefix, res[[1]]$section)
knitr::kable(res[[1]]$tab, digits = 3, caption = lab)

```

## 2d bivariate

```{r space-time_2d2v, eval=TRUE}
type = c("GNEITING_G", "GNEITING_M", "GNEITING_C")[2]
ndim = 2
nvar = 2
# parameters for the time trace
alpha = 2.0
beta  = 1.0
# parameters for the space trace
nu   = c(1/2, 3/4, 1)[1:nvar]
rr   = c(1, 2, 1)[1:nvar]
# space-time anisotropy
timeRange = 2.0
ranges = c(5,10)[1:ndim]
angles = c(30, 0)[1:ndim]
sep = 1.0

ctxt = CovContext(nvar = nvar, space = SpaceRN_create(ndim = ndim))
mod = CorGneiting_create(ctxt = ctxt, type = ECov_fromKey(type),
                       alpha = alpha, beta = beta, timeRange = timeRange,
                       params = nu, kappas = rr,
                       ranges = ranges, angles = angles,
                       flagRange = FALSE,
                       separability = sep)

# QC of the model
defineDefaultSpaceRnT(ndim = ndim)
err = draw_model(model = mod, prefix = type, flag.st = TRUE)
res = simu_model(model = mod, prefix = type, nsimu = nsimu, ns = nb, flag.st = TRUE)
lab = sprintf("Statistics for %s in section %s", res[[1]]$prefix, res[[1]]$section)
knitr::kable(res[[1]]$tab, digits = 3, caption = lab)

```

## 1d bivariate - factorized

```{r space-time_1d2v_fac, eval=TRUE}
type = c("GNEITING_G", "GNEITING_M", "GNEITING_C")[2]
ndim = 1
nvar = 2
# parameters for the time trace
alpha = 2.0
beta  = 1.0
# parameters for the space trace
nu   = c(1/2, 3/4, 1)[1:nvar]
rr   = c(1, 2, 1)[1:nvar]
# space-time anisotropy
timeRange = 2.0
ranges = c(5,10)[1:ndim]
angles = c(30, 0)[1:ndim]
sep = 0.0

ctxt = CovContext(nvar = nvar, space = SpaceRN_create(ndim = ndim))
mod = CorGneiting_create(ctxt = ctxt, type = ECov_fromKey(type),
                       alpha = alpha, beta = beta, timeRange = timeRange,
                       params = nu, kappas = rr,
                       ranges = ranges, angles = angles,
                       flagRange = FALSE,
                       separability = sep)

# QC of the model
defineDefaultSpaceRnT(ndim = ndim)
err = draw_model(model = mod, prefix = type, flag.st = TRUE)
res = simu_model(model = mod, prefix = type, nsimu = nsimu, ns = nb, flag.st = TRUE)
lab = sprintf("Statistics for %s in section %s", res[[1]]$prefix, res[[1]]$section)
knitr::kable(res[[1]]$tab, digits = 3, caption = lab)

```

## 2d bivariate - factorized

```{r space-time_2d2v_fac, eval=TRUE}
type = c("GNEITING_G", "GNEITING_M", "GNEITING_C")[2]
ndim = 2
nvar = 2
# parameters for the time trace
alpha = 2.0
beta  = 1.0
# parameters for the space trace
nu   = c(1/2, 3/4, 1)[1:nvar]
rr   = c(1, 2, 1)[1:nvar]
# space-time anisotropy
timeRange = 2.0
ranges = c(5,10)[1:ndim]
angles = c(30, 0)[1:ndim]
sep = 0.0

ctxt = CovContext(nvar = nvar, space = SpaceRN_create(ndim = ndim))
mod = CorGneiting_create(ctxt = ctxt, type = ECov_fromKey(type),
                       alpha = alpha, beta = beta, timeRange = timeRange,
                       params = nu, kappas = rr,
                       ranges = ranges, angles = angles,
                       flagRange = FALSE,
                       separability = sep)

# QC of the model
defineDefaultSpaceRnT(ndim = ndim)
err = draw_model(model = mod, prefix = type, flag.st = TRUE)
res = simu_model(model = mod, prefix = type, nsimu = nsimu, ns = nb, flag.st = TRUE)
lab = sprintf("Statistics for %s in section %s", res[[1]]$prefix, res[[1]]$section)
knitr::kable(res[[1]]$tab, digits = 3, caption = lab)

```

## 2d bivariate - partial separability

```{r space-time_2d2v_par, eval=TRUE}
# TEST of separability
type = c("GNEITING_G", "GNEITING_M", "GNEITING_C")[2]
nsimu = 20
ndim = 2
nvar = 2
sep = 1/2
# parameters for the time trace
alpha = 2.0
beta  = 1.0
# parameters for the space trace
nu   = c(1/2, 3/4, 1)[1:nvar]
rr   = c(1, 2, 1)[1:nvar]
# space-time anisotropy
timeRange = 2.0
ranges = c(5,10)[1:ndim]
angles = c(30, 0)[1:ndim]

ctxt = CovContext(nvar = nvar, space = SpaceRN_create(ndim = ndim))
mod = CorGneiting_create(ctxt = ctxt, type = ECov_fromKey(type),
                       alpha = alpha, beta = beta, timeRange = timeRange,
                       params = nu, kappas = rr,
                       ranges = ranges, angles = angles,
                       flagRange = FALSE,
                       separability = sep)

# QC of the model
defineDefaultSpaceRnT(ndim = ndim)
err = draw_model(model = mod, prefix = type, flag.st = TRUE)
res = simu_model(model = mod, prefix = type, nsimu = nsimu, ns = nb, flag.st = TRUE,
                 flag.QC = TRUE,
                 flag.histo = TRUE, flag.map = TRUE, flag.vario = TRUE, verbose = TRUE)

lab = sprintf("Statistics for %s in section %s", res[[1]]$prefix, res[[1]]$section)
knitr::kable(res[[1]]$tab, digits = 3, caption = lab)

lab = sprintf("Statistics for %s in section %s", res[[2]]$prefix, res[[2]]$section)
knitr::kable(res[[2]]$tab, digits = 3, caption = lab)

lab = sprintf("Statistics for %s in section %s", res[[3]]$prefix, res[[3]]$section)
knitr::kable(res[[3]]$tab, digits = 3, caption = lab)


```
