---
title: "Application of PPMT to a Gaussian mixture"
author: "D. Renard, N. Desassis, X. Freulon"
date: "December 6th 2022"
output:
  pdf_document: default
  html_document: default
  editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE, echo=TRUE}
knitr::opts_chunk$set(echo = TRUE)
rm(list=ls())
library(gstlearn)
library(ggplot2)
library(ggpubr)
OptCst_defineByKey("ASP",0)
```

# General Introduction

This notebook presents an example for the PPMT algorithm.

# Simulation of a reference data set

The data set is a mixture of 4 Gaussian random vectors in $\mathbb{R}^3$.

```{r simu_data_set, echo=TRUE, eval=TRUE}
np  <- 10000
nd  <- 3
nit <- 200
ndir<- 100
flag.ns <- FALSE

# mixing probabilities
pg <- c(0.25, 0.5, 0.2, 0.05)
ng <- table(sample(1:length(pg), prob = pg, size = np, replace = TRUE))

# center of the Gaussians
mu  <- vanDerCorput(n = length(pg), nd = nd)
mu  <- matrix(mu$getValues(), nrow = mu$getNRows(), ncol = mu$getNCols())

# dispersion of the Gaussians
ls <- list()
ls[[1+length(ls)]] <- 0.05 * diag(c(0.25,0.25,0.1))
ls[[1+length(ls)]] <- 0.05 * diag(c(0.2,0.01,0.1))
ls[[1+length(ls)]] <- 0.05 * diag(c(0.1,0.1,0.05))
ls[[1+length(ls)]] <- 0.05 * diag(c(0.25,0.01,0.25))


X <- data.frame()
for (k in seq_along(pg)){
  U <- rep(1,ng[k]) %*% t(mu[k,]) + matrix(rnorm(ng[k]*nd), nrow = ng[k], ncol = nd) %*% t(chol(ls[[k]]))
  X <- rbind(X,U)
}

# creation of a database

data <- Db_createFromSamples(nech = np)
data["x1"] <- X[,1]
data["x2"] <- X[,2]
data["x3"] <- X[,3]

dum = data$statistics(c("x*"), opers=EStatOption_fromKeys(c("MINI","MAXI","MEAN","STDV")))

# histograms and correlation
plot.hist(data, name = "x1", title = "X1", col = "gray", fill = "orange")
plot.hist(data, name = "x2", title = "X2", col = "gray", fill = "skyblue")
plot.hist(data, name = "x3", title = "X3", col = "gray", fill = "green")
plot.correlation(data, name1 = "x1", name2 = "x2", flagDiag = TRUE, xlab = "x1", ylab = "x2", 
                 title = "Gaussian mixture (data)")
plot.correlation(data, name1 = "x1", name2 = "x3", flagDiag = TRUE, xlab = "x1", ylab = "x3", 
                 title = "Gaussian mixture (data)")
plot.correlation(data, name1 = "x2", name2 = "x3", flagDiag = TRUE, xlab = "x2", ylab = "x3", 
                 title = "Gaussian mixture (data)")

```

# Application of PPMT

## Normal score

```{r}
# normal score
if(flag.ns){
  data["x1.NS"] = VectorHelper_normalScore(data["x1"])
  data["x2.NS"] = VectorHelper_normalScore(data["x2"])
  data["x3.NS"] = VectorHelper_normalScore(data["x3"])
} else {
  data["x1.NS"] = data["x1"]
  data["x2.NS"] = data["x2"]
  data["x3.NS"] = data["x3"]
}
data$setName("x1.NS","Y1")
data$setName("x2.NS","Y2")
data$setName("x3.NS","Y3")
```

## PPMT

```{r}
nd  <- 100
nit <- 100

data$setName("Y1","Z1")
data$setName("Y2","Z2")
data$setName("Y3","Z3")

# normal score
data["Y1"] = VectorHelper_normalScore(data["Z1"])
data["Y2"] = VectorHelper_normalScore(data["Z2"])
data["Y3"] = VectorHelper_normalScore(data["Z3"])

# Launching PPMT and fitting it on the vector of Input data
ppmt = PPMT_create(ndir=nd, FALSE, EDirGen_fromKey("vdc"))
Xmat <- data$getColumnsAsMatrix(names = c("Y1", "Y2", "Y3"))
ppmt$fitFromMatrix(Xmat, niter = nit, FALSE)

# scores vs. iterations
plot(x = 1:nit, y = log(ppmt$getSerieScore(), base = 10), type = "l", col = "red", lty = 3,
     xlab = "Iterations", ylab = "Logarithm of index", main = "Wasserstein distance")

# adding the results to the data base
iuid = data$addColumns(Xmat$getValues(), radix = "U")
dum  = data$statistics(c("x*", "Y*","U-*"),opers=EStatOption_fromKeys(c("MINI","MAXI","MEAN","STDV")))

```

## Results

### Initial Gaussian values

```{r}
# Q-Q plots for initial Gaussian
probas = (1:99)/(1+99)
q2 = VectorHelper_qnormVec(probas)

# First components
q1 = VectorHelper_quantiles(data$getColumns("Y1"), probas)
plot(q1, q2, xlab = "Experimental quantiles", ylab = "Theoretical quantiles", main = "First Initial Variable")
abline(a = 0, b = 1, col = "red", lty = 3)

# Second components
q1 = VectorHelper_quantiles(data$getColumns("Y2"), probas)
plot(q1, q2, xlab = "Experimental quantiles", ylab = "Theoretical quantiles", main = "Second Initial Variable")
abline(a = 0, b = 1, col = "red", lty = 3)

# Third components
q1 = VectorHelper_quantiles(data$getColumns("Y3"), probas)
plot(q1, q2, xlab = "Experimental quantiles", ylab = "Theoretical quantiles", main = "Second Initial Variable")
abline(a = 0, b = 1, col = "red", lty = 3)

plot.correlation(data, name1 = "Y1", name2 = "Y2", flagDiag = TRUE, xlab = "Y1", ylab = "Y2", 
                 title = "Initial Gaussian variables (data)")

plot.correlation(data, name1 = "Y1", name2 = "Y3", flagDiag = TRUE, xlab = "Y1", ylab = "Y3", 
                 title = "Initial Gaussian variables (data)")

plot.correlation(data, name1 = "Y2", name2 = "Y3", flagDiag = TRUE, xlab = "Y1", ylab = "Y3", 
                 title = "Initial Gaussian variables (data)")


```


### Final Gaussian values

```{r}
# First components
q1 = VectorHelper_quantiles(data$getColumns("U-1"), probas)
plot(q1, q2, xlab = "Experimental quantiles", ylab = "Theoretical quantiles", main = "First Final Variable")
abline(a = 0, b = 1, col = "red", lty = 3)

# Second components
q1 = VectorHelper_quantiles(data$getColumns("U-2"), probas)
plot(q1, q2, xlab = "Experimental quantiles", ylab = "Theoretical quantiles", main = "Second Final Variable")
abline(a = 0, b = 1, col = "red", lty = 3)

# Third components
q1 = VectorHelper_quantiles(data$getColumns("U-3"), probas)
plot(q1, q2, xlab = "Experimental quantiles", ylab = "Theoretical quantiles", main = "Second Final Variable")
abline(a = 0, b = 1, col = "red", lty = 3)

plot.correlation(data, name1 = "U-1", name2 = "U-2", flagDiag = TRUE, xlab = "U-1", ylab = "U-2", 
                 title = "Final Gaussian variables (data)")

plot.correlation(data, name1 = "U-1", name2 = "U-3", flagDiag = TRUE, xlab = "U-1", ylab = "U-3", 
                 title = "Final Gaussian variables (data)")

plot.correlation(data, name1 = "U-2", name2 = "U-3", flagDiag = TRUE, xlab = "U-2", ylab = "U-3", 
                 title = "Final Gaussian variables (data)")
```

# Back transform

## Implementation of the transform

```{r}
library(FNN)

PPMT.anam <- function(this, X){
  X  <- as.matrix(X)
  np <- dim(X)[1]
  nd <- dim(X)[2]
  stopifnot(nd == dim(this$X)[2])
  
  res<- get.knnx(data = this$X, query = X, k = 1+nd)
  Y  <- matrix(NaN, nrow = np, ncol = nd)
  for (i in 1:np) {
    idx <- res$nn.index[i,]
    w   <- res$nn.dist[i,]
    if (max(w == 0)){
      w   <- as.numeric(w == 0)
    } else {
      w   <- 1/w^2
      w   <- w / sum(w)
    }
    Y[i,] <- t(w) %*% this$Z[idx,]
  }
  Y
}
```

## Simulations and application of the transform

```{r}
# mapping Y to Z
anam <- list(
  Z = as.matrix(cbind(data["Z1"], data["Z2"], data["Z3"])),
  X = as.matrix(cbind(data["U-1"], data["U-2"], data["U-3"]))
  )

# simulations
ntest <- 10000
Y <- matrix(rnorm(n = ntest*3), nrow = ntest, ncol = 3)
Z <- PPMT.anam(this = anam, X = Y)

simu <- Db_createFromSamples(nech = ntest)
simu["x1"] <- Z[,1]
simu["x2"] <- Z[,2]
simu["x3"] <- Z[,3]

# Results
dum = simu$statistics(c("x*"), opers=EStatOption_fromKeys(c("MINI","MAXI","MEAN","STDV")))

# histograms and correlation
plot.hist(simu, name = "x1", title = "X1", col = "gray", fill = "orange")
plot.hist(simu, name = "x2", title = "X2", col = "gray", fill = "skyblue")
plot.hist(simu, name = "x3", title = "X3", col = "gray", fill = "green")
plot.correlation(simu, name1 = "x1", name2 = "x2", flagDiag = TRUE, xlab = "x1", ylab = "x2", 
                 title = "Gaussian mixture (simulation)")
plot.correlation(simu, name1 = "x1", name2 = "x3", flagDiag = TRUE, xlab = "x1", ylab = "x3", 
                 title = "Gaussian mixture (simulation)")
plot.correlation(simu, name1 = "x2", name2 = "x3", flagDiag = TRUE, xlab = "x2", ylab = "x3", 
                 title = "Gaussian mixture (simulation)")

# Functions
fun_corr <- function(i1, i2){
plot(Z[,i1], Z[,i2], pch = 19, cex = 0.5, col = "red",
     xlab = paste0("Z", i1), ylab = paste0("Z", i2), asp = 1)
points(anam$Z[,i1], anam$Z[,i2], pch = 19, cex = 0.1, col = "black")
legend("topleft", legend = c("Initial", "Simulated"), col = c("black", "red"), pch = 19)
  NULL
}

fun_qq <- function(i1, nq = 100){
  p <- (1:(nq-1))/nq
  q_ini <- quantile(x = anam$Z[,i1], probs = p)
  q_sim <- quantile(x = Z[,i1], probs = p)
  
  plot(q_sim, q_ini, xlab = "Simulated values", ylab = "Initial values", asp = 1, 
       main = paste0("Q-Q plot - Z", i1), type = "b")
  abline(a = 0, b = 1, col = "red", lty = 3)
  NULL
}

# Q-Q plots between initial an final variables
fun_qq(1, nq = 100)
fun_qq(2, nq = 100)
fun_qq(3, nq = 100)

# correations
fun_corr(1,2)
fun_corr(1,3)
fun_corr(2,3)
```

# Discussion and further works

* other indices ?

* implementations of the back transform (hermite expension or nearest neighbors

* Do we need to apply a initial normal, sphering, and ppmt?

# References

* Rivoirard, J. (1994). Introduction to disjunctive kriging and non-linear geostatistics. Number 551.021 R626i. Clarendon Press.
