---
title: "Van Der Corput seried and quasi random directions"
author: "X. Freulon"
date: "2022-11-02"
output: pdf_document
---

<!-- SUMMARY: Generation of Random Directions in a high-dimension space: Van der Corput sequence or Quasi-random model  -->

<!-- CATEGORY: Basic_Calculations -->

# Initialisation

```{r setup,include=FALSE, fig.show='hide'}
knitr::opts_chunk$set(fig.width=5, fig.height=4, 
                      out.height="70%", out.width="70%",
                      fig.align = "center") 
rm(list=ls())
library("gstlearn")
```

# The Van de Corput sequence

```{r, child="../../references/Van-der-Corput.md"}
```

# Evaluation of the Van der Corput sequence

## A single dimension

```{r dim1, echo=TRUE, eval=TRUE}
U = vanDerCorput(n = 10000, nd=1)$getColumn(0)
summary(U)
hist(U, col = "orange", main = "VDC in one dimension")
```

## Two dimensions

```{r dim2, echo=TRUE, eval=TRUE}

UG <- vanDerCorput(n = 10000, nd = 2)
UG1 = UG$getColumn(0)
UG2 = UG$getColumn(1)

hist(UG1)
hist(UG2)

plot(UG1[1:1000], UG2[1:1000], 
     xlab = "U", ylab = "V", main = "The first 1000 points")
```

## Many dimensions and few points

```{r many1, echo=TRUE, eval=TRUE}
np <- 100
nd <- 50
Umat <- vanDerCorput(n = np, nd = nd)
U = matrix(Umat$getValues(),nrow=Umat$getNRows())
means <- apply(X = U, MARGIN = 2, FUN = mean)
stds  <- apply(X = U, MARGIN = 2, FUN = sd)
plot(NULL, NULL, xlim = c(1,50), ylim = c(0,1),
     xlab = "Number of dimensions",
     ylab = "Mean or Std.", main = paste0("Van Der Corput sequence - N = ", np))
abline(h = 0.5, col = "orange", lty = 2)
abline(h = sqrt(1/12), col = "skyblue", lty = 2)
lines(1:50, means, lty = 1, col = "orange")
lines(1:50, stds , lty = 1, col = "skyblue")
```

## Many dimensions and many points

```{r many2, echo=TRUE, eval=TRUE}
np <- 100000
nd <- 50
Umat <- vanDerCorput(n = np, nd = nd)
U = matrix(Umat$getValues(),nrow=Umat$getNRows())
means <- apply(X = U, MARGIN = 2, FUN = mean)
stds  <- apply(X = U, MARGIN = 2, FUN = sd)
plot(NULL, NULL, xlim = c(1,50), ylim = c(0,1),
     xlab = "Number of dimensions",
     ylab = "Mean or Std.", main = paste0("Van Der Corput sequence - N = ", np))
abline(h = 0.5, col = "orange", lty = 2)
abline(h = sqrt(1/12), col = "skyblue", lty = 2)
lines(1:50, means, lty = 1, col = "orange")
lines(1:50, stds , lty = 1, col = "skyblue")
```

# $n$-sphere and (quasi) random directions

```{r, child="../../references/Random_Directions.md"}
```


```{r, echo=FALSE}
plot_dir <- function(np, nd = 3, meth = "vdc", ix = 1, iy = 2, seed=13364)
{
  tit <- paste0(nd, "-dimensions (", meth, ", np = ", np, ")")
  plot(cs, sn, xlim = 1.1*c(-1,1), ylim = 1.1*c(-1,1), type = "l",
       lty = 1, col = "red",
       xlab = paste0("X",ix), ylab = paste0("X",iy), main = tit, asp = 1) 
  abline(h = 0, col = "gray", lty = 3)
  abline(v = 0, col = "gray", lty = 3)

  set.seed(seed)

  if (meth == "vdc"){
    Umat <- vanDerCorput(n = np, nd = nd)
    U = matrix(Umat$getValues(),nrow=Umat$getNRows())
  } else {
    U = matrix(runif(np*nd), nrow = np, ncol = nd)
  }
  Umat = MatrixRectangular_createFromVD(U, np, nd)
  xmat = GeometryHelper_getDirectionsInRn(Umat)
  x = matrix(xmat$getValues(), nrow=xmat$getNRows())
  points(x[,ix], x[,iy], pch = 20, cex = 0.5)
}
```

# Test in two-dimensions

```{r}
phi <- 2*pi*(0:1000)/1001
cs <- cos(phi)
sn <- sin(phi)
```

```{r}
# Directions in R^3
nd <- 3
for (meth in c("vdc", "prng")) {
  for (np in c(100, 1000, 10000)){
    for (ix in 1:(nd-1)) {
      for (iy in (ix+1):nd){
        plot_dir(np, nd = nd, meth = meth, ix = ix, iy = iy) 
      }
    }
  }
}
```

```{r}
# Directions in R^6
np <- 1000
nd <- 6
meth <- "vdc" 
ix <- 1
iy <- 4
plot_dir(np, nd = nd, meth = meth, ix = ix, iy = iy)  
```

# References

* Devroye, L. (1986) Non-Uniform Random Variate Generation. New York: Springer-Verlag.

* Harman, R., & Lacko, V. (2010). On decompositional algorithms for uniform sampling from n-spheres and n-balls. Journal of Multivariate Analysis, 101(10), 2297-2304.

* Muller, M. E. (1959). A note on a method for generating points uniformly on n-dimensional spheres. Communications of the ACM, 2(4), 19-20.

* Sibuya, M. (1962). A method for generating uniformly distributed points onN-dimensional spheres. Annals of the Institute of Statistical Mathematics, 14(1), 81-85.


