---
title: "Regression with spatially varying coefficients in gstlearn"
author: "Xavier Freulon, Didier Renard, and the gstlearn Team"
date: "2025-04-23"
output:
  html_document:
    df_print: paged
  pdf_document: default
editor_options:
  markdown:
    wrap: 72
  chunk_output_type: console
---

# Introduction

```{r initial,echo=FALSE, message=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(knitr)
library(gstlearn)
rm(list=ls())
```

Regression models are build to explain a response variable over a region
or domain. The regression coefficients are often assumed to be constant
over the region of interest. However, in some applications, these
coefficients may vary at the local level. The objective of this note is
to present how this approach can be implemented using *gstlearn*.

## Technical description

```{r SVC_Theory, echo=FALSE, results='asis'}

cat(loadDoc("SVC.md"), sep = "\n")

```

# Simulation of the reference case

```{r setup, include=FALSE, echo=TRUE}
library(ggplot2)
library(ggpubr)
library(readxl)
```

We are defining some global environment variables within the *gstlearn* package.

```{r environment}
set.seed(43243)
pal   = rev(RColorBrewer::brewer.pal(6, "RdYlBu"))
opers = EStatOption_fromKeys(c("NUM", "MINI", "MAXI", "MEAN", "STDV"))
err = OptCustom_define("NotOptimSimpleCase", 1)

# list of estimations to be performed
ESTIM_KEY = c("obs", "res", "drift", "coeff")
estim_list = list()
estim_list[[1+length(estim_list)]] <- list(type = 1, idx = -1, name = "Z")
estim_list[[1+length(estim_list)]] <- list(type = 2, idx = -1, name = "RES")
estim_list[[1+length(estim_list)]] <- list(type = 3, idx = -1, name = "DRIFT")
estim_list[[1+length(estim_list)]] <- list(type = 4, idx = 0,  name = "A")
estim_list[[1+length(estim_list)]] <- list(type = 4, idx = 1,  name = "B")

suffix_key = c("estim", "stdev")
suffix_dsc = c("Estimation", "Std.")

flag_est = TRUE
flag_std = TRUE
flag_varz = FALSE
verbose = TRUE
```

The reference data set used to evaluate the different methods is
simulated on the unit square $[0,1]^2$. The parameters are first
simulated on a fine grid, then the observation are extracted from the
grid. For each observation $\alpha \in \mathcal{D}$, we get the observed
value $Z(\alpha)$ and the drifts $(f^l(\alpha))_{l \in 0:L}$. Note that
the $f^0$ is the constant value.

```{r data_set, echo=TRUE, eval=TRUE}
nx = c(50, 50)
dx = c(1,1) / nx
grd = DbGrid_create(x0 = c(0,0), dx = dx, nx = nx, angles = c(0, 0))

# Model of the spatially varying effects
tau = 0.1
m = c(1.25, 6.5)
rho   = 0.7
sigma = c(2, 3)
c0    = diag(sigma^2)
c0[1,2] = prod(sigma)*rho
c0[2,1] = prod(sigma)*rho
model = Model_createFromParam(ECov_MATERN(), sills = c0, range = 0.5, param = 1)
err   = model$setMeans(m)

# simulation of the spatially varying effects
err   = simtub(dbin = NULL, dbout = grd, model = model, nbsimu = 1, nbtuba = 1000, 
               namconv = NamingConvention("Y"))
err = grd$setName(old_name = "Y.1", name = "A")
err = grd$setName(old_name = "Y.2", name = "B")

# simulation of the external drift Lognormal (m = 3, sigma = 0.25)
m_N  <- 3
sigma_N <- 0.25
model_N = Model_createFromParam(ECov_CUBIC(), sill = 1.0, ranges = c(0.75, 0.75))
err   = simtub(dbin = NULL, dbout = grd, model = model_N, nbsimu = 1, nbtuba = 1000, 
               namconv = NamingConvention("N"))
grd["N"] <- m_N * exp(sigma_N * grd["N"] - 1/2 * sigma_N^2)

# Computing the observed values
grd["DRIFT"] = grd["A"] + grd["B"]*grd["N"]
grd["RES"]   = tau * rnorm(grd$getNSample())
grd["Z"]     = grd["DRIFT"] + grd["RES"]

# names of the estimated variables
actual_var = c("Z", "RES", "DRIFT", "A", "B")

# a selection on the output grid (a disk)
grd["sel"]  <- (sqrt((grd["x1"] - 0.5)^2 + (grd["x2"] - 0.5)^2) < 0.5)

# defining the observations
np <- 100
data = Db_createSamplingDb(grd, number = np, names = c("x1", "x2", "N", "Z"))
err = data$addSelectionRandom(prop = 0.8, seed = 13167, name = "sel")

# setting the roles
# on the grid
err = grd$setLocators("Z", ELoc_Z(), cleanSameLocator = TRUE)
err = grd$setLocators("N", ELoc_F(), cleanSameLocator = TRUE)
err = grd$setLocators("sel", ELoc_SEL(), cleanSameLocator = TRUE)
# on the data
err = data$setLocators("N", ELoc_F(), cleanSameLocator = TRUE)
err = data$setLocators("Z", ELoc_Z(), cleanSameLocator = TRUE)
err = data$setLocators("sel", ELoc_SEL(), cleanSameLocator = TRUE)

# the number of points in the moving neighborhood
n_neigh = 20

```


```{r data_set_plots, echo=FALSE, eval=TRUE}
var_list = c("A", "B", "N", "Z")
var_dscr = c("First varying Coefficient", "Second varying Coefficient", 
             "External predictor", "Observed value")
stopifnot(length(var_list) == length(var_dscr))

plot_simulated_data <- function(ivar) {
  stopifnot(ivar %in% 1:4)
  var_range = round(range(grd[var_list[ivar]], na.rm = TRUE), digits = 1)
  plot.init(asp = 1) + 
    plot.raster(grd, name = var_list[ivar], 
                flagLegend = FALSE, palette = pal, limits=var_range) +
    # plot.decoration(title = paste(var_dscr[ivar], var_list[ivar], sep = ": "))
    plot.decoration(title = paste(var_list[ivar], sep = ": "))
}

# plotting the simulated variables: the ingredients A, B, and N, the observed values
ggarrange(
  plot_simulated_data(1),
  plot_simulated_data(2),
  plot_simulated_data(3),
  plot_simulated_data(4), nrow = 2, ncol = 2
)

# Statistics on the simulated values
knitr::kable(dbStatisticsMono(db = grd, names = var_list, opers = opers)$toTL(), 
             caption = "Statistics on the grid", digits = 2)

# Statistics on the observed data
knitr::kable(dbStatisticsMono(db = data, names = var_list[c(3,4)], opers = opers)$toTL(), 
             caption = "Statistics on the observed data", digits = 2)

# Z vs. N
p = plot.init() + 
  plot.correlation(grd, namex = var_list[4], namey = var_list[3], asPoint = FALSE, bins = 100) +
  plot.decoration(
    xlab = paste0(var_list[4], " (", var_dscr[4], ")"), 
    ylab = paste0(var_list[3], " (", var_dscr[3], ")"))
plot.end(p)

# defining the names of the data
nm_obs = data$getNamesByLocator(locatorType = ELoc_Z())
nm_fac = data$getNamesByLocator(locatorType = ELoc_F())
```


```{r fn_plots, echo=FALSE, eval=TRUE}
# functions for plots
plot_res <- function(iest) {
  type = ESTIM_KEY[estim_list[[iest]]$type]
  idx  = estim_list[[iest]]$idx
  if (type != ESTIM_KEY[4]) {
    var_key = paste(prefix_section, type, sep = ".")
  } else {
    var_key = paste(prefix_section, type, idx, sep = ".")
  }
  p_list = list()
  # plots
  for (i in seq_along(suffix_key) ) {
    tit = paste(estim_list[[iest]]$name, suffix_key[i], sep = " - ")
    vv = paste(var_key, nm_obs, suffix_key[i], sep = ".")
    range_vv = range(grd[vv], na.rm = TRUE)
    p_list[[1+length(p_list)]] <- plot.init(asp=1) +
      plot.raster(grd, name = vv, flagLegend = FALSE, legendName = suffix_dsc[i],
                  palette = pal, limits=range_vv)  +
      plot.symbol(data) + 
      plot.decoration(title = tit)
  }
  ggarrange(plotlist = p_list, nrow = 1, ncol = length(p_list))
}

plot_corr <- function(iest) {
  # correlation
    p_corr <- plot.init() + plot.correlation(grd, 
                     namex = paste(var_key, nm_obs, suffix_key[1], sep = "."), 
                     namey = estim_list[[iest]]$name, 
                     asPoint = FALSE, flagDiag = TRUE, diagColor = "red", bins = 100) +
    plot.decoration(
      xlab = paste0("Estimated value"), 
      ylab = paste0("Actual value"),
      title = estim_list[[iest]]$name
      )
    
    plot.end(p_corr)
}

```

# Kriging with external drift with a unique/moving neighbourhood

## Initialisation

```{r KED_init, echo =  TRUE, eval=TRUE}
flag.unique.ked = FALSE
flag.unique = flag.unique.ked
if (flag.unique) {
  neigh  = NeighUnique()
  prefix_section = "KED"
  neigh_txt = "UNIQUE"
} else {
  neigh  = NeighMoving_create(nmaxi = n_neigh, nmini = n_neigh)
  prefix_section = "KEDM"
  neigh_txt = "MOVING"
}
prefix_ked = prefix_section
title_section = paste("using", neigh_txt, "neigh.")

modED = model$createReduce(0)
err   = modED$addCovFromParam(ECov_NUGGET(), sill = tau)
err   = modED$setDriftIRF(order = 0, nfex = 1)

define_estim <- function(estim, model, verbose = FALSE) {
  # --------------------------------------------
  # estimation of the observed variable
  # --------------------------------------------
  if (estim$type == 1) {
    for (idx in 0:(model$getNCov()-1)) {
      err = model$setCovFiltered(icov = idx, filter = FALSE)
    }
    for (idx in 0:(model$getNDrift()-1)) {
      err = model$getDriftList()$setFiltered(i = idx, filter = FALSE)
    }
  }
  # --------------------------------------------
  # estimation of the residual
  # --------------------------------------------
  if (estim$type == 2) {
    for (idx in 0:(model$getNCov()-1)) {
      err = model$setCovFiltered(icov = idx, filter = FALSE)
    }
    for (idx in 0:(model$getNDrift()-1)) {
      err = model$getDriftList()$setFiltered(i = idx, filter = TRUE)
    }
  }
  # --------------------------------------------
  # estimation of the observed drift
  # --------------------------------------------
  if (estim$type == 3) {
    for (idx in 0:(model$getNCov()-1)) {
      err = model$setCovFiltered(icov = idx, filter = TRUE)
    }
    for (idx in 0:(model$getNDrift()-1)) {
      err = model$getDriftList()$setFiltered(i = idx, filter = FALSE)
    }
  }
  # --------------------------------------------
  # estimation of the coefficient
  # --------------------------------------------
  if (estim$type == 4) {
    for (idx in 0:(model$getNCov()-1)) {
      err = model$setCovFiltered(icov = idx, filter = TRUE)
    }
    for (idx in 0:(model$getNDrift()-1)) {
      err = model$getDriftList()$setFiltered(i = idx, filter = !(idx == estim$idx))
    }
  }
  # --------------------------------------------
  if (verbose) {
    model$display()
  }
  invisible()
}  
```

## Estimation

```{r KED_estim, echo=TRUE, eval=TRUE}
# cleaning the target data base
err = grd$deleteColumns(names = paste0(prefix_section, ".*.", c("estim", "stdev"), ".*"))

for (iest in seq_along(estim_list)) {
  if (verbose) {
    print(paste0(">>> Estimation of ", estim_list[[iest]]$name, " using KED"))
  }
  # reset the model
  err = define_estim(estim = estim_list[[iest]], model = modED, verbose = FALSE)
  if (estim_list[[iest]]$type != 4) {
    var_key = paste(prefix_section, ESTIM_KEY[estim_list[[iest]]$type], sep = ".")    
  } else {
    var_key = paste(prefix_section, ESTIM_KEY[estim_list[[iest]]$type], estim_list[[iest]]$idx, sep = ".")
  }
  
  # computing the kriging
  err = kriging(data, dbout = grd, model = modED, neigh = neigh, krigopt = KrigOpt(),
                flag_est = flag_est, flag_std = flag_std,
                namconv = NamingConvention(var_key))
  
  # rescaling the drift component by the deterministic predictor to get the effect estimate
  if ((estim_list[[iest]]$type == 4)&(estim_list[[iest]]$idx > 0)) {
    if (flag_est) {
      grd[paste(var_key, nm_obs, suffix_key[1], sep = ".")] = 
        grd[paste(var_key, nm_obs, suffix_key[1], sep = ".")] /
        grd[nm_fac[estim_list[[iest]]$idx]]
    }
    if (flag_std) {
      grd[paste(var_key, nm_obs, suffix_key[2], sep = ".")] = 
        grd[paste(var_key, nm_obs, suffix_key[2], sep = ".")] /
        abs(grd[nm_fac[estim_list[[iest]]$idx]])
    }
  }
}
```

## Check

```{r KED_QC, echo=FALSE, eval=TRUE}
# plots  of Z
iest = 1; plot_res(iest); plot_corr(iest)

# plots of Res
iest = 2; plot_res(iest); plot_corr(iest)

# plots of DRIFT
iest = 3; plot_res(iest); plot_corr(iest)

# plots of A
iest = 4; plot_res(iest); plot_corr(iest)

# plots of B
iest = 5; plot_res(iest); plot_corr(iest)

# statistics on the estimations
knitr::kable(dbStatisticsMono(grd, names = paste(prefix_section, "*", sep = "."),
                              opers = opers)$toTL(), 
             caption = "Statistics on drift coefficients (actual values and estimated)", digits = 2)

# statistics on the residual
for (iest in 1:length(estim_list)) {
  if (estim_list[[iest]]$idx < 0) {
    vv = paste(sep = ".", prefix_section, ESTIM_KEY[estim_list[[iest]]$type], 
               nm_obs, "estim")
  } else {
    vv = paste(sep = ".", prefix_section, ESTIM_KEY[estim_list[[iest]]$type], 
               estim_list[[iest]]$idx, 
               nm_obs, "estim")
  }
  grd[paste(sep = ".", prefix_section, "diff", estim_list[[iest]]$name)] =
    grd[estim_list[[iest]]$name] - grd[vv]
}

knitr::kable(dbStatisticsMono(db = grd, 
                              names = c(paste(prefix_section, "diff", "*", sep = "."), "Y"), 
                              opers = opers)$toTL(), 
             caption = paste0("Statistics on the residual (", neigh_txt, ")"), digits = 2)

err = grd$deleteColumns(names = paste(prefix_section, "diff", "*", sep="."))
```

# Kriging with spatially varying coefficients with a unique/moving neighbourhood

## Initialisation

```{r SVC_init, echo = TRUE, eval=TRUE}
flag.unique.svc = TRUE
flag.unique = flag.unique.svc
if (flag.unique) {
  neigh  = NeighUnique()
  prefix_section = "SVC"
  neigh_txt = "UNIQUE"
} else {
  neigh  = NeighMoving_create(nmaxi = n_neigh, nmini = n_neigh)
  prefix_section = "SVCM"
  neigh_txt = "MOVING"
}
prefix_svc = prefix_section
title_section = paste("using", neigh_txt, "neigh.")

nfac = length(nm_fac)
model_svc  = model$clone()
C0 = matrix(c(tau, 0, 0, 0), nrow = 2, ncol = 2)
err = model_svc$addCovFromParam(type = ECov_NUGGET(), sills = C0)
err = model_svc$setDriftIRF(order = 0, nfex = nfac)
```

## Estimation

```{r SVC_estim, echo=TRUE, eval=TRUE}
# cleaning the target data base
err = grd$clearLocators(locatorType = ELoc_Z())
err = grd$deleteColumns(names = paste0(prefix_section, ".*.", c("estim", "stdev"), ".*"))

err = kriging_SVC(dbin = data, dbout = grd, model = model_svc, neigh = neigh,
                flag_est = flag_est, flag_std = flag_std, flag_varz = flag_varz, 
                estim_list = estim_list,
                prefix = prefix_section, verbose = verbose
)
```

## Checking

```{r SVC_QC, echo=FALSE, eval=TRUE}
# plots  of Z
iest = 1; plot_res(iest); plot_corr(iest)

# plots of Res
iest = 2; plot_res(iest); plot_corr(iest)

# plots of DRIFT
iest = 3; plot_res(iest); plot_corr(iest)

# plots of A
iest = 4; plot_res(iest); plot_corr(iest)

# plots of B
iest = 5; plot_res(iest); plot_corr(iest)

# statistics on the estimations
knitr::kable(dbStatisticsMono(grd, names = paste(prefix_section, "*", sep = "."),
                              opers = opers)$toTL(), 
             caption = "Statistics on drift coefficients (actual values and estimated)", digits = 2)

# statistics on the residual
for (iest in 1:length(estim_list)) {
  if (estim_list[[iest]]$idx < 0) {
    vv = paste(sep = ".", prefix_section, ESTIM_KEY[estim_list[[iest]]$type], 
               nm_obs, "estim")
  } else {
    vv = paste(sep = ".", prefix_section, ESTIM_KEY[estim_list[[iest]]$type], 
               estim_list[[iest]]$idx, 
               nm_obs, "estim")
  }
  grd[paste(sep = ".", prefix_section, "diff", estim_list[[iest]]$name)] =
    grd[estim_list[[iest]]$name] - grd[vv]
}

knitr::kable(dbStatisticsMono(db = grd, 
                              names = c(paste(prefix_section, "diff", "*", sep = "."), "Y"), 
                              opers = opers)$toTL(), 
             caption = paste0("Statistics on the residual (", neigh_txt, ")"), digits = 2)

err = grd$deleteColumns(names = paste(prefix_section, "diff", "*", sep="."))

```

# KED vs. SVC

```{r SVC_vs_KED, echo=FALSE, eval=TRUE}
# Plot of the estimations
for (iest in 1:length(estim_list)) {
  type = ESTIM_KEY[estim_list[[iest]]$type]
  idx  = estim_list[[iest]]$idx
  if (type != ESTIM_KEY[4]) {
    var_key = paste(type, nm_obs, sep = ".")
  } else {
    var_key = paste(type, idx, nm_obs, sep = ".")
  }
  nm_svc = paste(prefix_svc, var_key, sep = ".")
  nm_ked  = paste(prefix_ked, var_key, sep = ".")
  
  # correlation estim with KED vs. estim with SVC
  p = plot.init() + 
    plot.correlation(grd, 
                     namex = paste(nm_svc, "estim", sep = "."),
                     namey = paste(nm_ked, "estim", sep = "."), 
                     asPoint = FALSE, flagDiag = TRUE, diagColor = "red", bins = 100) +
    plot.decoration(xlab = "Estimated value (SVC)", ylab = "Estimated value (KED)", 
                    title = paste0("KED vs. SVC", ": ", estim_list[[iest]]$name))
  plot.end(p)
  
  # comparison
  if (iest %in% c(1,4,5)) {
    # actual value
    vnm = actual_var[iest]
    range_var = range(grd[vnm], na.rm = TRUE)
    p0 = plot.init(asp=1) + 
      plot.raster(grd, name = vnm, flagLegend = FALSE, palette = pal, limits = range_var) +
      plot.decoration(title = vnm)
    # estimation using SVC
    vnm = paste(nm_svc, "estim", sep = ".")
    range_var = range(grd[vnm], na.rm = TRUE)
    p =   plot.init(asp=1) + 
      plot.raster(grd, name = vnm, flagLegend = FALSE, palette = pal, limits = range_var) +
      plot.decoration(title = vnm)
    # estimation using KED
    vnm = paste(nm_ked, "estim", sep = ".")
    range_var = range(grd[vnm], na.rm = TRUE)
    q = plot.init(asp=1) + 
      plot.raster(grd, name = vnm, flagLegend = FALSE, palette = pal, limits = range_var) +
      plot.decoration(title = vnm)
    plot.end(ggarrange(p0, p, q, nrow = 1, ncol = 3))
  }
}

# statistics
stat_qc <- function(iest) {
  stopifnot(iest %in% c(1,4,5))
  lvar = paste(c(ESTIM_KEY[-4], paste(ESTIM_KEY[4], 0:nfac, sep = ".")), nm_obs, sep = ".")
  nm_ked  = paste(prefix_ked,  lvar, sep = ".")[iest]
  nm_svc = paste(prefix_svc, lvar, sep = ".")[iest]
  nm_v = c(estim_list[[iest]]$name, paste(c(nm_ked, nm_svc), "estim", sep = "."))
  knitr::kable(dbStatisticsMono(db = grd, names = nm_v, , opers = opers)$toTL(), 
               caption = paste0("Statistics for ", actual_var[iest]), digits = 2)
}

stat_qc(1)
stat_qc(4)
stat_qc(5)
```

