---
title: "gstlearn"
author: "Didier RENARD,  Nicolas DESASSIS"
output: pdf_document
---

```{r setup,echo=FALSE}
library(knitr)
```

# Introduction

**gstlearn** is a package developed by the Geostatistical Team of Mines ParisTech.
It contains several well established geostatistical functions and methods and some prototypes based on ongoing research.

# Loading gstlearn

In each R session, you must load the **gstlearn** library

```{r Load-library, include=FALSE}
library(gstlearn)
```

# R used as a pocket calculator

```{r Calculator-1}
1+3
exp(log(sqrt(2^2)))
```

## Assignments

```{r Assignments, message=FALSE}
a = 1
b <- 3
a + b
d = a + b
d
d = 7
d
```

## Working directory

```{r Working-directory, eval=FALSE}
ls()
rm(d)
ls()
```

## Quit

```{r Quit,eval=FALSE}
q()
```

Save the workspace (y).

The content of your session is saved in the file **.Rdata** (with the **.Rhistory**) of your working directory. 

*Caution*: these files are hidden.

To continue, you must relaunch R (and load gstlearn again).

## Classes in R

```{r Classes}
b <- 3
class(b)
col = "blue"
class(col)
reponse=FALSE
class(reponse)
```

## Booleans

```{r Boolean-1}
a = 1
a < 2
comp = a < 2
comp
comp & reponse
comp | reponse
!comp
```


```{r Boolean-2}
a = 1
a == 1
a != 1
a > 1
a >= 1
```

## Vectors

```{r Vectors-1}
x = c(1,9,2,9,4,5)
x
class(x)
x[1]
x[c(1,3)]
x[-2]
2:5*3
```


```{r Vectors-2}
x = c(1,9,2,9,4,5)
x[2:5]
selection = x > 2
selection
x[selection]
```

Length of a vector

```{r}
length(x)
```

## Matrices

```{r Matrices}
M = matrix(1:20,nrow=4,ncol=5)
M[2,1]
M[2,1]=3
M[2,1]
M[,1]
M[2,]
dim(M)
```

## Data frame

```{r Data-Frames-1}
data = data.frame(M)
class(data)
data[,3]
names(data)
names(data) = c("A","B","C","D","E")
data
```


```{r Data-Frames-2}
data
data$C
data[data$A>1,]
```

## Data Frame from a File

Download the ASCII file called *Scotland_Temperatures.csv* and store it on your disk in the current working directory.



```{r Data-Frame-from-File, eval=FALSE}
dlfile = "https://soft.minesparis.psl.eu/gstlearn/data/Scotland/Scotland_Temperatures.csv"
filename = "Scotland_Temperatures.csv"
download.file(dlfile, filename)
temperatures = read.csv(filename,header=TRUE,na="MISS")
class(temperatures)
names(temperatures)
```

## Some functions

```{r Functions-1}
set.seed(123)
rnorm(3)
x=c(3,4,7,4,2)
x
sort(x)
unique(x)
```


```{r Functions-2}
x=c(3,4,7,4,2)
sum(x)
mean(x)
var(x)
sd(x)
```

## Help of Functions

Search for help on the function **image**.

```{r Help,eval=FALSE}
?image
```

To get some help on a class :

```{r Help-on-Class,eval=FALSE}
class?matrix
```

## Some functions of interest

Function **outer**

```{r}
A = c(2,4,10)
B = c(3,7)
outer(A,B,"-")
outer(A,B,"*")
```

## Loops

Dimension and initialize a vector: **numeric** 

Loop from 1 to 10: store loop index in a vector; then print the vector

```{r}
tab = numeric(10)
for (i in 1:10)
{
  tab[i] = i
}
print(tab)
```

## Create your own function

Generate a linear transform of the input argument

```{r}
my.func = function(x,a=2,b=4)
{
  y = a * x + b
  y
}
```

Play my function

```{r}
x = seq(0,20)
y = my.func(x)
plot(x,y,main="Plotting my function")
```

