---
title: "Variography"
author: "gstlearn Team"
fontsize: 8pt
output:
  beamer_presentation:
    colortheme: beaver
    fig_height: 5
    fig_width: 8
    theme: Boadilla
  ioslides_presentation: default
theme: Boadilla
colortheme: beaver
---

\frametitle{Preamble}

In this preamble, we load the **gstlearn** library and clean the workspace. 

```{r Preamble, include=FALSE}
library(gstlearn)
rm(list=ls())
knitr::opts_chunk$set(fig.width=3.5, fig.height=2.5, fig.align = "center") 
```


Then the necessary data set is downloaded and named **dat**: the target variable is **January_temp**

```{r Loading_Data}
fileNF = paste(Sys.getenv('GSTLEARN_DATA'),'Scotland',
  'Scotland_Temperatures.NF',sep="/")
dat = Db_createFromNF(fileNF)
```

---

\frametitle{Experimental Variograms}

\begin{itemize}

\item Data is a \alert{regionalized} variable

$$z_i = z(x_i)$$

\item The experimental variogram is a (dicrete) function:

$$\gamma(h)=\frac{1}{2N(h)}\sum_{i=1}^{N(h)}[z(x_i+h)-z(x_i)]^2$$

where $N(h)$ is the number of pairs of points distant by $h$

\end{itemize}

---

\frametitle {Variogram Cloud}


```{r Variogram_Cloud}
varioParamOmni = VarioParam_createOmniDirection(100)
grid.cloud = db_variogram_cloud(dat, varioParamOmni)
grid.cloud$display()
plot.grid(grid.cloud, name="Cloud.January*")
```

---

\frametitle {Experimental Variogram}

We calculate the omni-directional variogram of the temperatures.

```{r Experimental_Variogram}
varioParamOmni = VarioParam_createOmniDirection(40, 10)
varioexp = Vario(varioParamOmni, dat)
err = varioexp$compute()
```


Print the variogram contents
```{r Print_Experimental_Variogram,eval=FALSE}
varioexp
```

---

\frametitle {Experimental Variogram}

Plot the omni-directional variogram

```{r Plot_Experimental_Variogram}
plot.varmod(varioexp)
```

---

\frametitle {Experimental Variogram}

```{r Plot_Experimental_Variogram_continued}
plot.varmod(varioexp,draw_psize=TRUE, draw_plabels=TRUE, ratio_psize=3)
```

---

\frametitle {Automatic Model Fitting}

```{r Variogram_Model}
fitmod = Model()
err = fitmod$fit(varioexp)
plot.varmod(varioexp, fitmod)
```

---

\frametitle {Automatic Model Fitting}

```{r Variogram_Model_Print}
fitmod
```

---

\frametitle {List of Basic structures}

```{r Basic_Structures}
ECov_printAll()
```

---

\frametitle {Automatic Fitting (with given basic structures)}

```{r Model_Fitting_List1}
err = fitmod$fit(varioexp, types=ECov_fromValues(c(0,4,2)))
plot.varmod(varioexp, fitmod)
```

---

\frametitle {Automatic Fitting (with given basic structures)}

```{r Model_Fitting_List2}
fitmod
```

---

\frametitle{Model Fitting with Inequality constraints}

```{r Model_Fitting_Inequality_1}
constraints = Constraints()
err = constraints$addItemFromParamId(EConsElem_RANGE(),1,0,0,EConsType_UPPER(),20.)
err = constraints$addItemFromParamId(EConsElem_SILL(),1,0,0,EConsType_LOWER(),0.03)
err = fitmod$fit(varioexp, types=ECov_fromValues(c(0,4,2)), constraints, Option_VarioFit(TRUE))
plot.varmod(varioexp, fitmod)
```

---

\frametitle{Model Fitting with Inequality constraints}

```{r Model_Fitting_Inequality_2}
fitmod
```

---

\frametitle{Model Fitting with Equality constraints}

```{r Model_Fitting_Equality_1}
constraints = Constraints()
err = constraints$addItemFromParamId(EConsElem_RANGE(),1,0,0,EConsType_EQUAL(),1000.)
err = constraints$addItemFromParamId(EConsElem_SILL(),1,0,0,EConsType_EQUAL(),0.4)
err = fitmod$fit(varioexp, types=ECov_fromValues(c(0,4,2)), constraints, Option_VarioFit(TRUE))
plot.varmod(varioexp, fitmod)
```

---

\frametitle{Model Fitting with Equality constraints}

```{r Model_Fitting_Equality_2}
fitmod
```

---

\frametitle{Directional Variograms}

```{r Directional_Variograms}
varioParamMulti = VarioParam_createMultiple(4, 15, 15.)
vario.4dir = Vario(varioParamMulti, dat)
err = vario.4dir$compute()
plot.varmod(vario.4dir)
```

---

\frametitle{Directional Variograms}

```{r Directional_Variograms_Fitting}
model.4dir = Model()
err = model.4dir$fit(vario.4dir,types=ECov_fromValues(c(0,4,2)))
plot.varmod(vario.4dir, model.4dir)
```

---

\frametitle{Calculating Variogram Map}

```{r Variogram_Map}
grid.vmap = db_vmap_compute(dat, ECalcVario_VARIOGRAM())
plot.grid(grid.vmap)
```

---

\frametitle{Automatic Model Fitting from Variogram Map}

```{r Fitting_Variogram_Map_Fit}
modelVM = Model()
err = modelVM$fitFromVMap(grid.vmap, ECov_fromValues(c(0,4,2)))
modelVM
```

---

\frametitle{Automatic Model Fitting from Variogram Map}

```{r Fitting_Variogram_Map_Draw}
err = dbgrid_model(grid.vmap, modelVM)
plot.grid(grid.vmap)
```

---

\frametitle{Compare Directional Variograms and Variogram Map}


```{r Compare_Directional_Variograms_and_Variogram_Map}
plot.varmod(vario.4dir, modelVM)
```

