---
title: "Multivariate Geostatistics"
author: "D. Renard, N. Desassis, X. Freulon"
date: "3 novembre 2021"
output: beamer_presentation
---

```{r setup,include=FALSE, fig.show='hide'}
knitr::opts_chunk$set(fig.width=4, fig.height=4, 
                      out.width="50%",
                      fig.align = "center") 
```

\frametitle{Loading Data}

```{r Loading_Library, include=FALSE}
rm(list=ls())
library(gstlearn)
library(ggplot2)
library(ggpubr)
```

## Preamble

\footnotesize

```{r Preamble, message=FALSE, warning=FALSE, include=FALSE}
fileNF = paste(Sys.getenv('GSTLEARN_DATA'),'Scotland',
               'Scotland_Temperatures.NF',sep="/")
dat = Db_createFromNF(fileNF)
fileNF = paste(Sys.getenv('GSTLEARN_DATA'),'Scotland',
               'Scotland_Elevations.NF',sep="/")
target = DbGrid_createFromNF(fileNF)
ndim = 2
defineDefaultSpace(ESpaceType_RN(), ndim);

unique_neigh = NeighUnique_create()
dat$setName("*temp", "Temperature")
```

## Statistics

\footnotesize

```{r EDA1, echo=FALSE, message=FALSE, warning=FALSE}
tab = dbStatisticsMonoT(dat,c("Elevation", "Temperature"))
knitr::kable(tab$toTL(), caption = "Statistics on observations")
tab = dbStatisticsMonoT(target,c("Elevation"))
knitr::kable(tab$toTL(), caption = "Statistics on the grid")
```

## Observations

```{r maps-elevation, echo=FALSE}
plot.setDefault(1, dims=c(8,8))
plot(dat, name_color="Elevation")
```

## Map of the elevations on Grid

```{r maps-elevation, echo=FALSE}
plot(target, name_raster="Elevation")
```

## Temperature vs. Elevation

\footnotesize

```{r temp_vs_ele, echo=FALSE}

p = plot.correlation(dat, name1="Elevation", name2="Temperature", 
                     asPoint=TRUE, flagRegr=TRUE)
plot.decoration(p, title="Correlation between Temperature and Elevation")
```

## Temperature vs. Elevation (statistiques)

\footnotesize

```{r temp_vs_ele2, echo=TRUE}
tab <- dbStatisticsMultiT(dat, c("Temperature", "Elevation"),
                          oper=EStatOption_MEAN(), flagMono=FALSE)
knitr::kable(tab$toTL(), caption = "Statistics on the grid", digits=3)
```

## Variograms and Non-stationarity

\footnotesize

Directional variograms (N-S and E-W): no drift or with first order global trend

\center

```{r Experimental_Directional_Variograms, echo=FALSE}
varioparam = VarioParam_createMultiple(ndir=2, npas=30, dpas=10)
vario_raw2dir = Vario(varioparam, dat)
err = vario_raw2dir$compute()

model = Model_create()
err = model$setDriftIRF(1)
vario_res2dir = Vario(varioparam, dat)
err = vario_res2dir$compute(model=model)

p = plot(vario_raw2dir, linetype="solid")
p = plot(vario_res2dir, linetype="dashed", padd=p)
plot.decoration(p, title="Temperature (째C)") 
```

## Global Trend

\footnotesize

Find the coefficients of the global regression (first order polynomial)

```{r Regression}
err = regression(dat, name0="Temperature", mode=2, model=model, verbose=TRUE)
```

## Modelling Variogram of Raw Data with Anisotropy

\footnotesize

```{r Model_for_Raw_Variogram, echo=FALSE}
fitmod = Model()
err = fitmod$fit(vario_raw2dir,
                 types=ECov_fromKeys(c("NUGGET","EXPONENTIAL","CUBIC")))

p = plot.varmod(vario_raw2dir, fitmod)
plot.decoration(p,title="Temperature (째C)")
```


## Ordinary Kriging - Cross-Validation

\footnotesize

Perform Cross-validation (using Ordinary Kriging) and calculate the Mean Squared Error.

```{r Cross-validation_Ordinary_Kriging}
err = xvalid(dat, fitmod, unique_neigh,
             namconv=NamingConvention("OK",TRUE,TRUE,FALSE))
```


```{r Cross-validation_Ordinary_Kriging_stat, echo=FALSE}
tab = dbStatisticsMonoT(dat, c("OK.Temperature.*"))
knitr::kable(tab$toTL(), caption= "Cross-validation with Ordinary Kriging")
```

## Ordinary Kriging - Estimation

\footnotesize

```{r Estimation_Ordinary_Kriging}
err = kriging(dat, target, fitmod, unique_neigh,
              namconv=NamingConvention("OK"))
```

```{r Estimation_Ordinary_Kriging_plot, echo=FALSE}
p = plot(target, "OK*estim")
p = plot(dat, name_size="Temperature", reduction = 200, ,padd=p)
plot.decoration(p, title="Temperature - Ordinary Kriging")
```

## Ordinary Kriging - Statistics

\scriptsize

```{r Estimation_Ordinary_Kriging_stat, echo = FALSE}
opers=EStatOption_fromKeys(c("NUM","MINI","MAXI","MEAN","STDV"))

tab <- dbStatisticsMonoT(target, names = (c("OK.T*")), opers=opers)
knitr::kable(tab$toTL(), digits=3, 
             caption = "Statistics on the Ordinary Kriging")
```


## Fitting Variogram of Residuals

\footnotesize

```{r Fitting_Residuals}

fitmodUK = Model()
err = fitmodUK$fit(vario_res2dir,
                   types=ECov_fromKeys(c("SPHERICAL")),
                   optvar=Option_VarioFit(FALSE,FALSE))
```


```{r}
p = plot.varmod(vario_res2dir, fitmodUK)
plot.decoration(p, title="Temperature (째C)")
```

Note that the residuals seem isotropic, hence use isotropic option for Fitting.

## Universal Kriging - Cross-Validation

\footnotesize

Perform Cross-validation (using Universal Kriging) and calculate the Mean Squared Error.

```{r Cross-validation_Universal_ Kriging}
err = xvalid(dat, fitmodUK, unique_neigh,
             namconv=NamingConvention("UK",TRUE,TRUE,FALSE))
```

```{r Cross-validation_Universal_ Kriging_stat, echo=FALSE}
tab = dbStatisticsMonoT(dat, c("UK.Temperature.*"), opers=opers)
knitr::kable(tab$toTL(), caption= "Cross-validation with Ordinary Kriging",
             digits=3)
```


## Universal Kriging - Estimation

\scriptsize

```{r Universal_Kriging}
err = kriging(dat, target, fitmodUK, unique_neigh,
              namconv=NamingConvention("UK"))
```
 
```{r Universal_Kriging_stat, echo = FALSE}
p = plot(target, "UK*estim")
p = plot(dat, name_size="Temperature", padd=p)
plot.decoration(p, title="Temperature - Universal Kriging")
```

## Universal Kriging - Statistics

\scriptsize

```{r Estimation_Universal_Kriging_stat, echo = FALSE}
tab = dbStatisticsMonoT(target, c("UK.T*"), opers=opers)
knitr::kable(tab$toTL(), caption= "Statistics on the Universal Kriging",
             digits=3)
```


## Comparing Ordinary and Universal Krigings

\scriptsize

```{r Comparing_Ordinary_and_Universal_Krigings, results="hide"}
p = plot.correlation(target, "OK*estim", "UK*estim", flagDiag=TRUE, bins=100)
plot.decoration(p, xlab="Ordinary Kriging",ylab="Universal Kriging")
```


## Comparing Temperature and Elevation

\scriptsize

```{r Correlation_of_Temperature_and_Elevation}
plot.correlation(dat,"Elevation","Temperature",flagRegr=TRUE,asPoint=TRUE)
```

## Comparing Temperature and Elevation

\scriptsize


```{r selection_Temp_Ele, echo=TRUE}
dat$setLocators(c("Temperature", "Elevation"), ELoc_Z())
```

## Bivariate Modelling

\scriptsize

```{r Bivariate_Modelling, fig.align='center'}
varioexp2var = Vario_create(varioparam, dat)
err = varioexp2var$compute()
fitmod2var = Model()
err = fitmod2var$fit(varioexp2var,
                     types=ECov_fromKeys(c("NUGGET","EXPONENTIAL","CUBIC")))
```

```{r}
p = plot.varmod(varioexp2var, fitmod2var)
plot.decoration(p, title="Temperature (째C) and Elevation")
```

## Cokriging with elevation - Cross-Validation

\scriptsize

Most of the processes are more time-consuming in Unique Neighborhood.
We create a small neighborhood for demonstration.

```{r CK_neigh}
moving_neigh = NeighMoving_create(radius = 1000, nmaxi = 10)
```

Perform Cross-validation (Bivariate Model) and calculate the Mean Squared Error.

```{r CK_Cross-validation, echo=TRUE}
err = xvalid(dat, fitmod2var, moving_neigh,
             namconv=NamingConvention("COK",TRUE,TRUE,FALSE))
```


```{r CK_Cross-validation_stat, echo=FALSE}
tab = dbStatisticsMonoT(dat, c("COK.T*"), opers=opers)
knitr::kable(tab$toTL(), caption= "Cross-validation with Cokriging",
             digits=3)
```

## Cokriging with elevation - Estimate

\scriptsize

```{r Cokriging}
err = kriging(dat, target, fitmod2var, unique_neigh,
              namconv=NamingConvention("COK"))
```

```{r Cokriging_plot, echo=FALSE}
p = plot(target, "COK.T*estim")
p = plot(dat, name_size="Temperature", padd=p)
plot.decoration(p, title="Temperature - CoKriging (with Elevation)")
```

## Cokriging with elevation - Statistics

\scriptsize

```{r Estimation_cokriging_stat, echo = FALSE}
tab = dbStatisticsMonoT(target, c("COK.T*"), opers=opers)
knitr::kable(tab$toTL(), caption= "Statistics on Cokriging", digits=3)
```

## Comparing Kriging and CoKriging

\scriptsize

```{r Comparing_Kriging_and_CoKriging, results="hide", echo=FALSE}
p = plot.correlation(target, "OK.T*estim", "COK.T*estim", 
                     flagBiss=TRUE, bins=100)
plot.decoration(p, xlab="Ordinary Kriging",ylab="Ordinary CoKriging")

tab = dbStatisticsMonoT(target, c("COK.T*estim", "OK.T*estim"), opers=opers)
knitr::kable(tab$toTL(), 
             caption="Statistics on Kriging and CoKriging estimates",
             digits=3)
```

Note that CoKriging produces estimates which are mostly larger than Kriging estimates.

## Kriging the residuals

\scriptsize

$$
Z_2(s)=b + a Z_1(s) + R(s)
$$
```{r residual_init, echo=TRUE}
regr = regression(dat, "Temperature", "Elevation", flagCste=TRUE, verbose=TRUE)
b = regr$coeffs[1]
a = regr$coeffs[2]

err = dbRegression(dat, "Temperature", "Elevation",
                     namconv = NamingConvention("Regr",TRUE,TRUE,FALSE))
```


```{r residual_init_stat, echo=FALSE}
tab = dbStatisticsMonoT(dat, c("Regr*"), opers=opers)
knitr::kable(tab$toTL(), caption= "Statistics on the residual", digits=3)
```


## Kriging the residuals - Correlation

\scriptsize

```{r residual_correlation, echo=TRUE}
plot.correlation(dat,"Elevation","Regr*",flagRegr=TRUE,asPoint=TRUE)
```

## Kriging the residuals - Variogram of the residual

\scriptsize

```{r residual_variogram, echo=TRUE}
dat$setLocator("Regr*",ELoc_Z(), cleanSameLocator=TRUE)
varioexpR = Vario(varioparam, dat)
err = varioexpR$compute()
fitmodR = Model()
err = fitmodR$fit(varioexpR,
                  types=ECov_fromKeys(c("NUGGET","SPHERICAL","ORDER1_GC")))
```

```{r}
p = plot.varmod(varioexpR, fitmodR)
plot.decoration(p, title="Temperature Residual")
```

## Kriging the residuals - Variogram of the residual

\scriptsize

```{r residual_kriging, echo=TRUE}
err = kriging(dat, target, fitmodR, unique_neigh,
              namconv=NamingConvention("ROK"))
```

```{r residual_kriging_plot, echo=FALSE}
p = plot(target, "ROK.Regr*estim")
p = plot(dat, name_size="Temperature", padd=p)
plot.decoration(p, title="Temperature Residuals")
```

## Kriging the residuals - Computing the estimate

\scriptsize

$$
Z_2^{\star} = b + a Z_1(s) + R(s)^{OK}
$$

```{r residual_estimate, echo=TRUE}
ROK_estim = target["ROK.Regr*estim"] + b + a * target["Elevation"]
uid = target$addColumns(ROK_estim,"KR.Temperature.estim")
```

```{r residual_estimate_plot, echo=FALSE}
p = plot(target,"KR.T*estim")
p = plot(dat, name_size="Temperature", padd=p)
plot.decoration(p, title="Temperature with Residuals")
```

## Kriging the residuals - Correlation

Correlation between Ordinary Kriging and CoKriging
 
\scriptsize

```{r residual_estimate_corr, echo=FALSE}
p = plot.correlation(target, "OK.T*estim", "COK.T*estim", flagBiss=TRUE,
                     bins=100)
plot.decoration(p, xlab="Ordinary Kriging", ylab="Ordinary CoKriging")
```

## Kriging the residuals - Correlation

Correlation between Ordinary Kriging and Kriging with Residuals
 
\scriptsize

```{r residual_estimate_corr2, echo=FALSE}
p = plot.correlation(target, "OK.T*estim", "KR.T*estim", flagBiss=TRUE,
                     bins=100)
plot.decoration(p, xlab="Ordinary Kriging",ylab="Kriging with Residuals")
```

## Kriging the residuals - Correlation

\scriptsize

```{r residual_estimate_stat, echo=FALSE}
tab <- dbStatisticsMonoT(target, opers=opers,
                          names = (c("OK.T*estim", "COK.T*estim",
                                     "KR.T*estim")))
knitr::kable(tab$toTL(), digits=3,
             caption="Statistics on OK and Kriging of the resisual")
```

## Using Elevation Map as External Drift

\scriptsize

Preparing the data bases

```{r Preparation_for_KDE}
err = dat$setLocator("Temperature",ELoc_Z(),cleanSameLocator=TRUE)
err = dat$setLocator("Elevation",ELoc_F(),cleanSameLocator=TRUE)
```

## Linear Regression of Temperature knowing Elevation on Data

\scriptsize

We perform the Linear Regression of Temperature knowing Elevation (specified as External Drift)

```{r Regression_Temperature_Elevation}
regr = regression(dat, "Temperature", mode=1, flagCste=TRUE, verbose=TRUE)
```

## Experimental Variogram of Residuals (External Drift)

\scriptsize

```{r Fitting_Variogram_of_Residuals_with_ED}
varioKED = Vario(varioparam, dat)

model = Model_create()
err = model$setDriftIRF(order=0, nfex=1)
err = varioKED$compute(model=model)
```

```{r Fitting_Variogram_of_Residuals_with_ED_fig, echo=FALSE}
p = plot(vario_raw2dir)
p = plot(varioKED, padd=p)
p
```

## Model of Residuals (External Drift)

\scriptsize

```{r Model_of_Residuals_with_External_Drift}
modelKED = Model()
err = modelKED$fit(varioKED,
                   types=ECov_fromKeys(c("NUGGET","CUBIC","GAUSSIAN")))
modelKED$setDriftIRF(order = 0, nfex = 1)

plot.varmod(varioKED, modelKED)
```

## Kriging with External Drift - Cross-Validation

\scriptsize

Perform Cross-validation (External Drift) and calculate the Mean Squared Error.

```{r Cross-Validation_External_Drift}
err = xvalid(dat, modelKED, unique_neigh,
             namconv=NamingConvention("KED",TRUE,TRUE,FALSE))
```

```{r Cross-Validation_External_Drift_stat, echo=FALSE}
tab <- dbStatisticsMonoT(dat, opers=opers, names = (c("KED.T*")))
knitr::kable(tab$toTL(), digits=3,
             caption="Cross-validation with Kriging with External Drift")
```

## Kriging with External Drift - Estimate

\scriptsize

```{r Kriging_with_External_Drift}
err = kriging(dat, target, modelKED, unique_neigh,
              namconv=NamingConvention("KED"))
```

```{r Kriging_with_External_Drift_plot, echo=FALSE}
p = plot(target, "KED.T*estim")
p = plot(dat, name_size="Temperature", padd=p)
plot.decoration(p, title="Temperature with External Drift")
```

## Kriging with External Drift - Statistics

\scriptsize

```{r Estimation_ked_stat, echo = FALSE}
tab <- dbStatisticsMonoT(target, opers=opers, names = (c("KED.T*")))
knitr::kable(tab$toTL(), 
             caption = "Statistics on the Kriging with External Drift")
```

## Comparing OK with KED

\scriptsize

```{r Comparing_Ordinary_Kriging_with_Kriging_with_External_Drift, results="hide", echo=FALSE}
p = plot.correlation(target, "OK.T*estim", "KED.T*estim", flagBiss=TRUE,
                     bins=100)
plot.decoration(p, xlab="Ordinary Kriging", ylab="Kriging with External Drift")
```

Note that negative Estimates are present when using External Drift.

## Summary of Cross-validation scores

\scriptsize

Comparing the cross-validation Mean Squared Errors

```{r Comparing_Cross-validations}
tab <- dbStatisticsMonoT(dat, opers=opers, names = (c("*.Temperature.esterr")))
knitr::kable(tab$toTL(),
  caption = "Statistics of the cross-validation error for the temperature"
  )
```

## Statistics on the estimates

\scriptsize

Computing the statistics on the various estimates.

```{r Comparing_estimates_estim, echo = TRUE}
tab <- dbStatisticsMonoT(target, opers=opers, 
                         names = (c("*.Temperature.estim")))
knitr::kable(tab$toTL(), digits=3,
  caption = "Statistics of the estimates for the temperature"
  )
```

## Statistics on the estimates

\scriptsize

Computing the statistics on the various estimates.

```{r Comparing_estimates_stdev, echo = TRUE}

tab <- dbStatisticsMonoT(target, opers=opers, 
                         names = (c("*.Temperature.stdev")))
knitr::kable(tab$toTL(), digits=3,
  caption = "Statistics of the kriging Std. for the temperature"
  )
```
