---
title: "Conditional expectation"
author: "gstlearn Team"
date: "February 2023"
output:
  beamer_presentation:
    colortheme: beaver
    fig_height: 6
    theme: Boadilla
theme: Boadilla
colortheme: beaver
fontsize: 8pt
---

```{r setup,include=FALSE, fig.show='hide'}
knitr::opts_chunk$set(fig.width=4, fig.height=4, 
                      out.width="50%",
                      fig.align = "center") 
```

\frametitle{Preamble}

```{r loading_library, message=FALSE}
rm(list=ls())
library(gstlearn)
library(ggplot2)
library(ggpubr)
library(ggnewscale)

flagInternetAvailable=TRUE

```

```{r Preamble, message=FALSE}
## Data points
fileNF = "Scotland_Temperatures.NF"
if(flagInternetAvailable){
  download.file(paste0("https://soft.minesparis.psl.eu/gstlearn/data/Scotland/",fileNF), fileNF, quiet=TRUE)
}
dat = Db_createFromNF(fileNF)

## Target grid
fileNF = "Scotland_Elevations.NF"
if(flagInternetAvailable){
  download.file(paste0("https://soft.minesparis.psl.eu/gstlearn/data/Scotland/",fileNF), fileNF, quiet=TRUE)
}
grid = DbGrid_createFromNF(fileNF)
```

---

\frametitle{Histogram of the raw variable (temperature)}

```{r Histogram}
p = ggplot()
p = p + plot.hist(dat,"January*")
p = p + plot.decoration(title="Temperatures")
ggPrint(p)
```

---

\frametitle{Gaussian scores}

```{r Gaussian_Score}
anam = AnamHermite(30)
err = anam$fitFromLocator(dat)
err = anam$rawToGaussian(dat, "January_temp")
anam
```

---

\frametitle{Gaussian scores}

```{r Gaussian_Score_Plot}
p = ggplot()
p = p + plot.XY(dat["Y.January_temp"], dat["January_temp"], flagLine=FALSE, flagPoint=TRUE)
p = p + plot.decoration(xlab="Gaussian", ylab="Raw")
ggPrint(p)
```

---

\frametitle{Gaussian scores}

```{r Gaussian_Score_Histogram}
p = ggplot()
p = p + plot.hist(dat,"Y.January*")
p = p + plot.decoration(title="Temperatures (Gaussian scale)")
ggPrint(p)
```

---

\frametitle{Kriging of the Gaussian scores}
We calculate the experimental directional variogram of the gaussian scores and fit the Model (with the constraints that sill should be 1)

```{r Fitting_Model}
varioparam = VarioParam_createMultiple(ndir=2, npas=40, dpas=10)
vario_gauss2dir = Vario_create(varioparam, dat)
err = vario_gauss2dir$compute()

fitmodgauss = Model()
err = fitmodgauss$fit(vario_gauss2dir, 
                      types=ECov_fromKeys(c("NUGGET", "SPHERICAL","CUBIC")),
                      constraints = Constraints(1))
```

```{r}
ggplot() + plot.varmod(vario_gauss2dir, fitmodgauss)
```

---

\frametitle{Kriging of the Gaussian scores}

```{r Kriging_gaussian_scores } 
neighU = NeighUnique_create()

err = kriging(dat, grid, fitmodgauss, neighU)
```

\frametitle{Kriging of the Gaussian scores}
\framesubtitle{Estimation}

```{r}
p = ggDefaultGeographic()
p = p + plot(grid,"*estim")
p = p + plot(dat)
p = p + plot.decoration(title="Kriging of Gaussian scores")
ggPrint(p)
```

---

\frametitle{Kriging of the Gaussian scores}
\framesubtitle{Standard Deviation}

```{r Kriging_Gaussian_scores_stdev}
p = ggDefaultGeographic()
p = p + plot(grid,"*stdev")
p = p + plot(dat, flagCst=TRUE)
p = p + plot.decoration(title="St. Dev. of Gaussian scores")
ggPrint(p)
```

---

\frametitle{Conditional expectation}

Use the Turning Bands method with 1000 simulations

```{r}
selectivity = Selectivity_createByKeys(c("Z"), flag_est=TRUE, flag_std=TRUE)
err = ConditionalExpectation(grid, anam, selectivity, "K*.estim", "K*.stdev",
                             nbsimu=100,
                             namconv=NamingConvention("CE",FALSE,TRUE,FALSE))
```

---

\frametitle{Conditional expectation}

```{r Display_Conditional_Expectation}
p = ggDefaultGeographic()
p = p + plot(grid,"CE*estim")
p = p + plot(dat)
p = p + plot.decoration(title = "Conditional Expectation")
ggPrint(p)
```

---

\frametitle{Conditional Standard Deviation}

```{r Conditional_Standard_Deviation}
p = ggDefaultGeographic()
p = p + plot(grid, "CE*stdev")
p = p + plot(dat, flagCst=TRUE)
p = p + plot.decoration(title="Conditional Standard Deviation")
ggPrint(p)
```

---

\frametitle{Conditional probability}
\framesubtitle{Below cut-off 0}

```{r Conditional_Probability_above_0}
selectivity = Selectivity_createByKeys(c("PROP"), zcuts=c(0),
                                       flag_est=TRUE, flag_std=TRUE)
err = ConditionalExpectation(grid, anam, selectivity, 
                             "K*.estim", "K*.stdev",
                             namconv=NamingConvention("CE",FALSE,TRUE,FALSE))
```


```{r}
p = ggDefaultGeographic()
p = p + plot(grid,"CE.Proba*estim")
p = p + plot(dat)
p = p + plot.decoration(title = "Conditional Probability below 0")
ggPrint(p)
```

---

\frametitle{Conditional probability}
\framesubtitle{Above cut-off 1}

```{r Conditional_Probability_above_1}
selectivity = Selectivity_createByKeys(c("T"), zcuts=c(1),
                                       flag_est=TRUE, flag_std=TRUE)
err = ConditionalExpectation(grid, anam, selectivity, "K*.estim", "K*.stdev",
                             namconv=NamingConvention("CE",FALSE,TRUE,FALSE))

```

---

\frametitle{Conditional probability above 1}
\framesubtitle{Estimation}

```{r}
p = ggDefaultGeographic()
p = p + plot(grid,"CE.T*estim-1")
p = p + plot(dat)
p = p + plot.decoration(title = "Conditional Probability above 1")
ggPrint(p)
```

---

\frametitle{Conditional probability above 1}
\framesubtitle{Standard Deviation}

```{r Final_Histogram}
p = ggDefaultGeographic()
p = p + plot(grid, "CE.T*stdev-1")
p = p + plot(dat, flagCst=TRUE)
p = p + plot.decoration(title = "Conditional probability (Standard Deviation)")
ggPrint(p)
```


