---
title: "Tutorial on SPDE"
author: "gstlearn Team"
output:
  html_document:
    df_print: paged
    toc: true
  pdf_document:
    toc: true_variogram_compute
---

<!-- SUMMARY: Using Stochastic Partial Derivative Equation framework to perform Estimation (extension to the use of External Drift)  -->

<!-- CATGEORY: Courses -->

```{r setup,include=FALSE, fig.show='hide'}
knitr::opts_chunk$set(fig.width=4, fig.height=4, 
                      out.width="50%",
                      fig.align = "center") 
```

# Preamble

In this preamble, we load the **gstlearn** library, clean the workspace.

```{r loading_library, message=FALSE}
rm(list=ls())
library(gstlearn)
library(ggplot2)
library(ggpubr)
library(ggnewscale)

flagInternetAvailable=TRUE
```

# Getting the Data Bases from the repository.

The input Data Base (called **temperatures**) contains the target variable (**January_temp**). Note that this data base also contains the elevation variable (called **Elevation**) which is assigned the locator **f** (for external drift). We finally add a selection (called **sel**) which only consider the samples where the temperature is actually calculated.

```{r Loading_Data}
fileNF = "Scotland_Temperatures.NF"
if(flagInternetAvailable){
  download.file(paste0("https://soft.minesparis.psl.eu/gstlearn/data/Scotland/",fileNF), fileNF, quiet=TRUE)
}
temperatures = Db_createFromNF(fileNF)

temperatures$setLocator("Elevation", ELoc_F())
iuid = temperatures$addSelection(as.numeric(! is.na(temperatures["J*"])))
temperatures$display()
```

The output file is a grid (called grid). It contains an active selection (inshore) as well as the elevation over the field (called Elevation). Note that this variable is assigned the locator f (external drift).

```{r Loading_Grid}
fileNF = "Scotland_Elevations.NF"
if(flagInternetAvailable){
  download.file(paste0("https://soft.minesparis.psl.eu/gstlearn/data/Scotland/",fileNF), fileNF, quiet=TRUE)
}
grid = DbGrid_createFromNF(fileNF)
grid$display()
```

# Variography

Calculate the omni-directional variogram of the temperature for 18 lags of 25.

```{r Preamble,message=FALSE, fig.show='hide'}
vparam = VarioParam_createOmniDirection(npas=18, dpas=25)
vario = Vario(vparam,temperatures)
err = vario$compute()
vario$display()
```

```{r}
p = ggDefault()
p = p + plot.varmod(vario)
p = p + plot.decoration(title="Variogram of Temperature (Raw)")
ggPrint(p)
```

We calculate the variogram (using the same calculation parameters) based on the residuals after a trend has been removed. This trend is considered as a linear combinaison of the external drift information.

```{r}
vparam = VarioParam_createOmniDirection(npas=18, dpas=25)
vario = Vario(vparam,temperatures)
md = Model()
md$setDriftIRF(order=0, nfex=1)
vario$compute(model=md)
vario$display()
```

```{r}
p = ggDefault()
p = p + plot.varmod(vario)
p = p + plot.decoration(title="Variogram of Temperature (Residuals)")
ggPrint(p)
```

Fit the variogram of residuals in a model having drifts. Some constraints have been added during the fitting step.

```{r Fitting_Model_residuals}
model = md
structs = ECov_fromKeys(c("NUGGET", "BESSEL_K"))

const = Constraints()
err = const$addItemFromParamId(EConsElem_SILL(),icov=0,type=EConsType_UPPER(),value=0.1)
err = const$addItemFromParamId(EConsElem_PARAM(),icov=1, type = EConsType_EQUAL(),value = 1)
err = const$addItemFromParamId(EConsElem_RANGE(),icov=1, type = EConsType_LOWER(),value = 100)
err = const$addItemFromParamId(EConsElem_RANGE(),icov=1, type = EConsType_UPPER(),value = 350)

err = md$fit(vario,structs,constraints=const)
```

```{r}
p = ggDefault()
p = p + plot.varmod(vario, model)
p = p + plot.decoration(title="Variogram of Temperature (Residuals)")
ggPrint(p)
```

# Identify the Drift

Derive the parameter of a Global Trend (composed of the Elevation as a drift function) using the SPDE formalism.

```{r Trend_coefficients}
spde = SPDE(model,grid,temperatures,ESPDECalcMode_KRIGING())
coeffs = spde$getCoeffs()
cat("Trend coefficients:", coeffs)
```

Represent the scatter plot of the Temperature given the Elevation and add the Global Trend (caledlated beforehand)

```{r Creating_Grid}
p = ggplot()
p = p + plot.correlation(temperatures,namex="Elevation",namey="*temp", asPoint=TRUE)
p = p + geom_segment(aes(x = 0, y = coeffs[1], xend = 400, yend = coeffs[1] + 400 * coeffs[2]))
ggPrint(p)
```

# Estimation 

We perform the Estimation in the SPDE framework (considering the Elevation as the Global Trend)

```{r SPDE_Kriging}
err = krigingSPDE(temperatures, grid, model)
```

```{r Plot_SPDE_KRIGING}
p = ggDefaultGeographic()
p = p + plot.grid(grid)
p = p + plot.decoration(title="Temperature (using Elevation as global Trend)")
ggPrint(p)
```

We also perform the Estimation by Kriging (using Elevation as External Drift). This estimation is performed in Unique Neighborhood.

```{r Standard_Kriging}
neighU = NeighUnique_create()
err = kriging(temperatures, grid, model, neighU)
```

```{r Plot_Standard_Kriging}
p = ggDefaultGeographic()
p = p + plot.grid(grid)
p = p + plot.decoration(title="Temperature (with Elevation as External Drift)")
ggPrint(p)
```

Comparing both estimates

```{r}
p = ggplot()
p = p + plot.correlation(grid,namex="Kriging.January_temp.estim",
                         namey="KrigingSPDE.January_temp.estim", asPoint=TRUE)
p = p + plot.decoration(xlab="Kriging with External Drift", ylab="SPDE Kriging")
ggPrint(p)
```
